---
title: "fire_response_strategy_and_predictors"
author: "Sophie"
date: "22/11/2021"
output: html_document
editor_options:
  chunk_output_type: console
---


# Fire response strategy and fire frequency

Load libraries

```{r}
library(tidyverse)
library(patchwork)
```

Read in fire response strategy data

```{r}
resprouting_seeding <- readRDS("data/resprouting_seeding.rds")
```



## Mean fires from Poisson regression

Read in mean fire return interval data from GLMs

```{r}
mean_fires <- read_csv("outputs/mean_fires_df_with_poisson_glm.csv", guess_max = 10000)

# Check overdispersion/underdispersion
mean_fires |>
  filter(dispersion_p_value < .05) |>
  pull(dispersion_ratio) |>
  hist()
mean_fires |>
  filter(dispersion_p_value < .05, dispersion_ratio > 2) |>
  nrow() # 2243

mean_fires <- mean_fires %>% left_join(resprouting_seeding, by = c("taxon_name"))

mean_fires |> nrow() # 10871
mean_fires |>
  filter(num_pixels < 10) |>
  nrow() # 465
mean_fires |>
  filter(num_pixels < 10, !is.na(warnings) | !is.na(error) | !is.na(messages)) |>
  nrow() # 193
mean_fires |>
  filter(num_pixels > 10, !is.na(warnings) | !is.na(error) | !is.na(messages)) |>
  nrow() # 157
```

Filter out taxa with warning messages

```{r}
mean_fires_no_errors <-
  mean_fires %>%
  filter(is.na(warnings), is.na(error), is.na(messages)) %>%
  select(-warnings, -messages, -error)

mean_fires_no_errors |> nrow() # 10512
mean_fires |>
  filter(!is.na(warnings) | !is.na(error) | !is.na(messages)) |>
  nrow() # 359
mean_fires |>
  filter(num_unburnt_pixels == num_pixels) |>
  nrow() # 298
mean_fires |> nrow() # 10871
mean_fires |>
  filter(num_pixels < 10) |>
  nrow() # 465
mean_fires |>
  filter(num_pixels > 10) |>
  filter(num_unburnt_pixels == num_pixels) |>
  nrow() # 117
```

Clean data

```{r}
# Eliminate species with less than 10 pixels
mean_fires_clean <- mean_fires_no_errors |>
  dplyr::filter(num_pixels >= 10) |> # Removed 272
  # Convert to mean number of fires in 100 years, instead of mean number of fires in the sampling period
  mutate(
    mean_fires_100 = mean_fires / 22.17 * 100,
    mean_fires_100_upr = upr_conf_int / 22.17 * 100,
    mean_fires_100_lwr = lwr_conf_int / 22.17 * 100,
    conf_int = mean_fires_100_upr - mean_fires_100_lwr
  )

mean_fires_clean |>
  dplyr::filter(conf_int > 5) |>
  nrow() # 307

mean_fires_clean$conf_int |> hist()

mean_fires_clean |> nrow() # 10240

df_woody_herb_only <-
  mean_fires_clean %>%
  dplyr::filter(woody_or_herb %in% c("woody", "herb")) |>
  mutate(across(c("resprouting_binomial", "seeding_binomial", "woody_or_herb"), as.factor))

# write_csv(df_woody_herb_only, "outputs/final_df_predicted_fires_and_fire_response.csv")
# df_woody_herb_only <- read_csv("outputs/final_df_predicted_fires_and_fire_response.csv")
```


Histogram of mean fires

```{r}
hist(mean_fires_clean$mean_fires_100, breaks = 40)
```


TODO Investigate proportion of annuals

```{r}
# remotes::install_github("traitecoevo/austraits", dependencies = TRUE, upgrade = "ask")
library(austraits)
austraits <- load_austraits(version = "6.0.0", path = "data/")
```

```{r}
# Wenk_2023 dataset
life_history <- austraits$traits |>
  filter(trait_name == "life_history", dataset_id == "Wenk_2023") |>
  pivot_wider(values_from = value, names_from = trait_name) |>
  select(all_of(c("taxon_name", "life_history"))) |>
  group_by(taxon_name) |> # Fix when there are multiple rows per species # TODO Flag with Lizzy
  summarise(life_history = life_history[1])

life_history_summary <-
  life_history |>
  mutate(
    annual_perennial = case_when(
      str_detect(life_history, "annual") & !str_detect(life_history, "perennial") ~ "annual",
      str_detect(life_history, "perennial") & !str_detect(life_history, "annual") ~ "perennial",
      str_detect(life_history, "perennial") & str_detect(life_history, "annual") ~ "perennial" # Assign perennial if both annual and perennial values are listed
    )
  )

# TODO How do we decide annual vs perennial when both annual and perennial are listed?
# "I think I'd push toward perennial - because being a resprouter only relevant if you're perennial,
# but there is probably more nuance to this than I've thought about"
# TODO Could also check what resprouting/seeding classification the ones with both annual and perennial are
```

```{r}
df_woody_herb_only_life_history <- df_woody_herb_only |>
  left_join(life_history_summary)

# 26 species without data from Wenk_2023
missing_species <- df_woody_herb_only_life_history |>
  filter(is.na(life_history)) |>
  pull(taxon_name)
missing_species |> length() # 26

austraits$traits |>
  filter(trait_name == "life_history", taxon_name %in% missing_species) |>
  pivot_wider(values_from = value, names_from = trait_name) |>
  View() # Only 6 of these have data elsewhere in AusTraits
```


What are the proportions of resprouters and seeders in annuals and perennials?

```{r}
df_woody_herb_only_life_history |>
  filter(is.na(annual_perennial)) |>
  nrow() # Could not get life history for 33 species

df_woody_herb_only_life_history |>
  filter(woody_or_herb == "herb", !is.na(annual_perennial)) |>
  mutate(across(c("resprouting_binomial", "seeding_binomial"), ~ if_else(.x == TRUE, 1, 0))) |>
  group_by(annual_perennial) |>
  summarise(
    prop_resprout = sum(resprouting_binomial, na.rm = TRUE) / sum(!is.na(resprouting_binomial)),
    prop_seed = sum(seeding_binomial, na.rm = TRUE) / sum(!is.na(seeding_binomial))
  )

df_woody_herb_only_life_history |>
  filter(woody_or_herb == "herb", !is.na(resprouting_binomial)) |>
  mutate(across(c("resprouting_binomial", ), ~ if_else(.x == TRUE, 1, 0))) |>
  group_by(resprouting_binomial) |>
  summarise(
    prop_annual = sum(annual_perennial == "annual", na.rm = TRUE) / sum(!is.na(annual_perennial)),
    prop_perennial = sum(annual_perennial == "perennial", na.rm = TRUE) / sum(!is.na(annual_perennial))
  )

df_woody_herb_only_life_history |>
  filter(woody_or_herb == "herb", !is.na(seeding_binomial)) |>
  mutate(across(c("seeding_binomial"), ~ if_else(.x == TRUE, 1, 0))) |>
  group_by(seeding_binomial) |>
  summarise(
    prop_annual = sum(annual_perennial == "annual", na.rm = TRUE) / sum(!is.na(annual_perennial)),
    prop_perennial = sum(annual_perennial == "perennial", na.rm = TRUE) / sum(!is.na(annual_perennial))
  )
```


TODO Supplementary table 

```{r}

```


### Binomial logistic regression


#### Resprouting binomial

```{r}
# Subset normal logistic regression to taxa in phylogenetic logistic regression
# to check effect of phylogenetic regression
# tree <- ape::read.tree("data/v0.1-big-seed-plant-trees/ALLMB.tre")
# tree$tip.label <- gsub("_", " ", tree$tip.label)
# df_woody_herb_only |> nrow()
# df_woody_herb_only <- df_woody_herb_only[df_woody_herb_only$taxon_name %in% tree$tip.label, ]

glm_resprouting <- glm(
  resprouting_binomial ~
    log10(mean_fires_100) + I(log10(mean_fires_100)^2) + log10(mean_fires_100):woody_or_herb +
    I(log10(mean_fires_100)^2):woody_or_herb + woody_or_herb,
  data = df_woody_herb_only,
  family = binomial
)

summary(glm_resprouting)
sjPlot::tab_model(glm_resprouting)

ggeffects::ggpredict(glm_resprouting, c("mean_fires_100", "woody_or_herb")) %>%
  ggplot(aes(x, predicted, color = group)) +
  geom_line(linewidth = 1) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group), alpha = 0.2, linetype = 0) +
  scale_x_log10() +
  ylab("Resprouting probability") +
  xlab("Mean fires") +
  theme_classic()

modelbased::estimate_expectation(glm_resprouting) |>
  ggplot(aes(x = mean_fires_100, y = Predicted, color = woody_or_herb)) +
  geom_line(linewidth = 1) +
  geom_ribbon(aes(ymin = CI_low, ymax = CI_high, fill = woody_or_herb), alpha = 0.2, linetype = 0) +
  scale_x_log10() +
  ylab("Resprouting probability") +
  xlab("Mean fires") +
  theme_classic()
```


Original figure without confidence intervals

```{r}
# coef2 <- coef(glm_resprouting)
# x1 <- log10(df_woody_herb_only$mean_fires_100)
# x2 <- df_woody_herb_only$woody_or_herb
# x2 <- ifelse(x2 == "woody", 1, 0)
# pred_y <- coef2[1] + coef2[2] * x1 + coef2[3] * x1^2 + coef2[4] * x2 + coef2[5] * x1 * x2 + coef2[6] * x1^2 * x2
# pred_p <- exp(pred_y) / (1 + exp(pred_y))
# df1 <- data.frame(x1, x2, df_woody_herb_only$woody_or_herb, pred_y, pred_p)
# df1 <- df1 %>% rename(woody_or_herb = df_woody_herb_only.woody_or_herb)

# resprouting_plot <-
#   ggplot(data = df1, aes(x = 10^x1, y = pred_p, col = woody_or_herb)) +
#   geom_line(linewidth = 1.3) +
#   scale_x_log10(
#     expand = c(0, 0),
#     limits = c(0.01, 100),
#     labels = c("0.01", "0.1", "1", "10", "100")
#   ) +
#   scale_y_continuous(
#     limits = c(0, 0.8),
#     expand = c(0, 0)
#   ) +
#   ylab("Resprouting probability") +
#   xlab("Fire frequency (per century)") +
#   scale_colour_manual(
#     values = c("#5e3a25", "#8aa540"),
#     name = "",
#     breaks = c("woody", "herb"),
#     labels = c("Woody", "Herb")
#   ) +
#   ggtitle("a)") +
#   theme_classic() +
#   theme(
#     panel.border = element_rect(colour = "black", fill = NA),
#     legend.position = "none",
#     axis.text = element_text(size = 10),
#     plot.margin = unit(c(0.1, 1, 0.1, 0.1), "cm"),
#     plot.title.position = "plot",
#     plot.title = element_text(size = 22, margin = margin(0, 0, 20, 0)),
#     axis.title.y = element_text(size = 16, margin = margin(0, 15, 0, 0)),
#     axis.title.x = element_blank(),
#     panel.grid.major = element_line(colour = "#ebebeb", linewidth = 0.5),
#     axis.line = element_blank()
#   )

# resprouting_plot
```


Add confidence intervals

```{r}
# Extract model coefficients
coefficients <- coef(glm_resprouting)

# Create the design matrix for the new data
design_matrix <- model.matrix(~ log10(mean_fires_100) + I(log10(mean_fires_100)^2) + log10(mean_fires_100):woody_or_herb +
  I(log10(mean_fires_100)^2):woody_or_herb + woody_or_herb, data = df_woody_herb_only)


# Calculate predicted values on the logit scale
logit_values <- design_matrix %*% coefficients

# Convert logit values to probabilities
predicted_probabilities <- 1 / (1 + exp(-logit_values))

# Calculate standard errors
vcov_matrix <- vcov(glm_resprouting)
standard_errors_logit <- sqrt(diag(design_matrix %*% vcov_matrix %*% t(design_matrix)))

# Calculate confidence intervals on the logit scale
confidence_level <- 0.95
z_value <- qnorm((1 + confidence_level) / 2)
CI_low_logit <- logit_values - z_value * standard_errors_logit
CI_high_logit <- logit_values + z_value * standard_errors_logit

# Convert confidence intervals to probabilities
CI_low <- 1 / (1 + exp(-CI_low_logit))
CI_high <- 1 / (1 + exp(-CI_high_logit))

# Add predicted probabilities and confidence intervals to the df_woody_herb_only dataframe
df_woody_herb_only$predicted_resprouting <- predicted_probabilities
df_woody_herb_only$CI_low <- CI_low
df_woody_herb_only$CI_high <- CI_high

# Plot the predicted probabilities with confidence intervals
(resprouting_plot <- ggplot(
  df_woody_herb_only,
  aes(x = mean_fires_100, y = predicted_resprouting, color = woody_or_herb)
) +
  geom_line(linewidth = 1.3) +
  geom_ribbon(aes(ymin = CI_low, ymax = CI_high, fill = woody_or_herb), alpha = 0.2, linetype = 0) +
  scale_x_log10(
    expand = c(0, 0),
    limits = c(0.01, 100),
    labels = c("0.01", "0.1", "1", "10", "100")
  ) +
  scale_y_continuous(
    limits = c(0, 0.8),
    expand = c(0, 0)
  ) +
  ylab("Resprouting probability") +
  xlab("Fire frequency (per century)") +
  scale_colour_manual(
    values = c("#5e3a25", "#8aa540"),
    name = "",
    breaks = c("woody", "herb"),
    labels = c("Woody", "Herb")
  ) +
  scale_fill_manual(
    values = c("#b19586", "#9fb846"),
    name = "",
    breaks = c("woody", "herb"),
    labels = c("Woody", "Herb")
  ) +
  ggtitle("a)") +
  theme_classic() +
  theme(
    panel.border = element_rect(colour = "black", fill = NA),
    legend.position = "none",
    axis.text = element_text(size = 10),
    plot.margin = unit(c(0.1, 1, 0.1, 0.1), "cm"),
    plot.title.position = "plot",
    plot.title = element_text(size = 22, margin = margin(0, 0, 20, 0)),
    axis.title.y = element_text(size = 16, margin = margin(0, 15, 0, 0)),
    axis.title.x = element_blank(),
    panel.grid.major = element_line(colour = "#ebebeb", linewidth = 0.5),
    axis.line = element_blank()
  ))
```


Density distributions

```{r}
medians <- df_woody_herb_only %>%
  drop_na(resprouting_binomial) %>%
  group_by(woody_or_herb, resprouting_binomial) %>%
  summarise(median = median(mean_fires_100)) # Medians are robust to log scale and also more suited to non-normal distributions

data_labels <- tibble(
  label = c("Woody", "Herb"),
  woody_or_herb = c("woody", "herb"),
  x = c(0.015, 0.015),
  y = c(1, 1)
)

resprouting_density <-
  df_woody_herb_only %>%
  drop_na(resprouting_binomial) %>%
  ggplot() +
  geom_density(
    aes(mean_fires_100, fill = resprouting_binomial),
    alpha = 0.6,
    linewidth = 0.8,
    col = NA
  ) +
  # geom_vline(
  #  data = medians,
  #  aes(xintercept = median, col = resprouting_binomial),
  #  alpha = 1,
  #  size = 1,
  #  linetype = "longdash"
  # ) +
  facet_wrap(~ factor(woody_or_herb, ordered = TRUE, levels = c("woody", "herb")), ncol = 1, scales = "free") +
  geom_text(
    data = data_labels,
    aes(x = x, y = y, label = label, colour = label),
    hjust = 0,
    size = 7,
    fontface = "bold"
  ) +
  scale_x_log10(
    expand = c(0, 0),
    limits = c(0.01, 100),
    labels = c("0.01", "0.1", "1", "10", "100")
  ) +
  scale_y_continuous(
    expand = c(0, 0),
    limits = c(0, 1.25),
    breaks = seq(0, 1.25, by = 0.625),
    labels = c("0", "0.625", "1.25")
  ) +
  scale_fill_manual(
    values = c("#eb8775", "#172e6d"),
    labels = c("No", "Yes")
  ) +
  scale_colour_manual(
    values = c("#8aa540", "#5e3a25")
  ) +
  labs(
    x = "Fire frequency (per century)",
    y = "Density",
    fill = "Resprouting/seeding"
  ) +
  theme_classic() +
  theme(
    axis.line = element_blank(),
    strip.background = element_blank(),
    strip.text = element_blank(),
    plot.margin = unit(c(0.5, 1, 0.1, 0.1), "cm"),
    panel.border = element_rect(colour = "black", fill = NA),
    panel.spacing = unit(0.5, "lines"),
    axis.text = element_text(size = 10),
    axis.title.y = element_text(size = 16, margin = margin(0, 15, 0, 0)),
    axis.title.x = element_text(size = 16, margin = margin(15, 0, 0, 0)),
    legend.position = "none"
  )
resprouting_density
```





#### Seeding binomial


```{r}
glm_seeding <- glm(
  seeding_binomial ~
    log10(mean_fires_100) + I(log10(mean_fires_100)^2) + log10(mean_fires_100):woody_or_herb +
    I(log10(mean_fires_100)^2):woody_or_herb + woody_or_herb,
  data = df_woody_herb_only,
  family = binomial
)

summary(glm_seeding)
sjPlot::tab_model(glm_seeding)

ggeffects::ggpredict(glm_seeding, c("mean_fires_100", "woody_or_herb")) %>%
  ggplot(aes(x, predicted, color = group)) +
  geom_line(linewidth = 1) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group), alpha = 0.2, linetype = 0) +
  scale_x_log10() +
  ylab("Seeding probability") +
  xlab("Mean fires") +
  theme_classic()

modelbased::estimate_expectation(glm_seeding) |>
  ggplot(aes(x = mean_fires_100, y = Predicted, color = woody_or_herb)) +
  geom_line(linewidth = 1) +
  geom_ribbon(aes(ymin = CI_low, ymax = CI_high, fill = woody_or_herb), alpha = 0.2, linetype = 0) +
  scale_x_log10() +
  ylab("Seeding probability") +
  xlab("Mean fires") +
  theme_classic()

# Herb data does not look like it has enough statistical power
```


Try with only herb data

```{r}
df_herb_only <-
  df_woody_herb_only %>%
  filter(woody_or_herb == "herb") |>
  mutate(across(c("resprouting_binomial", "seeding_binomial", "woody_or_herb"), as.factor))


glm_seeding_herb <- glm(
  seeding_binomial ~ log10(mean_fires_100) + I(log10(mean_fires_100)^2),
  data = df_herb_only,
  family = binomial
)

summary(glm_seeding_herb)
sjPlot::tab_model(glm_seeding_herb)

ggeffects::ggpredict(glm_seeding_herb, c("mean_fires_100")) %>%
  ggplot(aes(x, predicted)) +
  geom_line(linewidth = 1) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2, linetype = 0) +
  scale_x_log10() +
  ylab("Seeding probability") +
  xlab("Mean fires") +
  theme_classic()
```


Model without herb data

```{r}
df_woody_only <-
  df_woody_herb_only %>%
  filter(woody_or_herb == "woody") |>
  mutate(across(c("resprouting_binomial", "seeding_binomial", "woody_or_herb"), as.factor))

glm_seeding_woody <- glm(
  seeding_binomial ~ log10(mean_fires_100) + I(log10(mean_fires_100)^2),
  data = df_woody_only,
  family = binomial
)

summary(glm_seeding_woody)
sjPlot::tab_model(glm_seeding_woody)

ggeffects::ggpredict(glm_seeding_woody, c("mean_fires_100")) %>%
  ggplot(aes(x, predicted)) +
  geom_line(linewidth = 1) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2, linetype = 0) +
  scale_x_log10() +
  ylab("Seeding probability") +
  xlab("Mean fires") +
  theme_classic()
```


Plot fitted model

```{r}
# coef2 <- coef(glm_seeding_woody)
# x1 <- log10(df_woody_only$mean_fires_100)
# x2 <- df_woody_only$woody_or_herb
# x2 <- ifelse(x2 == "woody", 1, 0)
# pred_y <- coef2[1] + coef2[2] * x1 + coef2[3] * x1^2
# pred_p <- exp(pred_y) / (1 + exp(pred_y))
# df1 <- data.frame(x1, x2, df_woody_only$woody_or_herb, pred_y, pred_p)
# df1 <- df1 %>% rename(woody_or_herb = df_woody_only.woody_or_herb)

# seeding_plot <-
#   ggplot(data = df1, aes(x = 10^x1, y = pred_p, col = woody_or_herb)) +
#   geom_line(linewidth = 1.3) +
#   scale_x_log10(
#     expand = c(0, 0),
#     limits = c(0.01, 100),
#     labels = c("0.01", "0.1", "1", "10", "100")
#   ) +
#   scale_y_continuous(
#     limits = c(0, 1.0),
#     expand = c(0, 0)
#   ) +
#   ylab("Seeding probability") +
#   xlab("Fire frequency (per century)") +
#   scale_colour_manual(
#     values = c("#5e3a25"),
#     name = "",
#     breaks = c("woody"),
#     labels = c("Woody")
#   ) +
#   ggtitle("b)") +
#   theme_classic() +
#   theme(
#     panel.border = element_rect(colour = "black", fill = NA),
#     legend.position = "none",
#     axis.text = element_text(size = 10),
#     plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"),
#     plot.title.position = "plot",
#     plot.title = element_text(size = 22, margin = margin(0, 0, 20, 0)),
#     axis.title.y = element_text(size = 16, margin = margin(0, 15, 0, 0)),
#     axis.title.x = element_blank(),
#     panel.grid.major = element_line(colour = "#ebebeb", linewidth = 0.5),
#     axis.line = element_blank()
#   )

# seeding_plot
```


Plot with confidence intervals

```{r}
# Extract model coefficients
coefficients <- coef(glm_seeding_woody)

# Create the design matrix for the new data
design_matrix <- model.matrix(
  ~ log10(mean_fires_100) + I(log10(mean_fires_100)^2),
  data = df_woody_only
)


# Calculate predicted values on the logit scale
logit_values <- design_matrix %*% coefficients

# Convert logit values to probabilities
predicted_probabilities <- 1 / (1 + exp(-logit_values))

# Calculate standard errors
vcov_matrix <- vcov(glm_seeding_woody)
standard_errors_logit <- sqrt(diag(design_matrix %*% vcov_matrix %*% t(design_matrix)))

# Calculate confidence intervals on the logit scale
confidence_level <- 0.95
z_value <- qnorm((1 + confidence_level) / 2)
CI_low_logit <- logit_values - z_value * standard_errors_logit
CI_high_logit <- logit_values + z_value * standard_errors_logit

# Convert confidence intervals to probabilities
CI_low <- 1 / (1 + exp(-CI_low_logit))
CI_high <- 1 / (1 + exp(-CI_high_logit))

# Add predicted probabilities and confidence intervals to the df_woody_only dataframe
df_woody_only$predicted_resprouting <- predicted_probabilities
df_woody_only$CI_low <- CI_low
df_woody_only$CI_high <- CI_high

# Plot the predicted probabilities with confidence intervals
(seeding_plot <- ggplot(
  df_woody_only,
  aes(x = mean_fires_100, y = predicted_resprouting, color = woody_or_herb)
) +
  geom_line(linewidth = 1.3) +
  geom_ribbon(aes(ymin = CI_low, ymax = CI_high, fill = woody_or_herb), alpha = 0.2, linetype = 0) +
  scale_x_log10(
    expand = c(0, 0),
    limits = c(0.01, 100),
    labels = c("0.01", "0.1", "1", "10", "100")
  ) +
  scale_y_continuous(
    limits = c(0, 1.0),
    expand = c(0, 0)
  ) +
  ylab("Seeding probability") +
  xlab("Fire frequency (per century)") +
  scale_colour_manual(
    values = c("#5e3a25"),
    name = "",
    breaks = c("woody"),
    labels = c("Woody")
  ) +
  scale_fill_manual(
    values = c("#b19586"),
    name = "",
    breaks = c("woody"),
    labels = c("Woody")
  ) +
  ggtitle("b)") +
  theme_classic() +
  theme(
    panel.border = element_rect(colour = "black", fill = NA),
    legend.position = "none",
    axis.text = element_text(size = 10),
    plot.margin = unit(c(0.1, 1, 0.1, 0.1), "cm"),
    plot.title.position = "plot",
    plot.title = element_text(size = 22, margin = margin(0, 0, 20, 0)),
    axis.title.y = element_text(size = 16, margin = margin(0, 15, 0, 0)),
    axis.title.x = element_blank(),
    panel.grid.major = element_line(colour = "#ebebeb", linewidth = 0.5),
    axis.line = element_blank()
  ))
```

Density distributions

```{r}
medians <- df_woody_only %>%
  drop_na(seeding_binomial) %>%
  group_by(seeding_binomial) %>%
  summarise(median = median(mean_fires_100))

data_labels <- tibble(
  label = c("Woody"),
  woody_or_herb = c("woody"),
  x = c(0.015),
  y = c(1)
)

seeding_density <-
  df_woody_only %>%
  drop_na(seeding_binomial) %>%
  ggplot(aes(mean_fires_100, fill = seeding_binomial)) +
  geom_density(alpha = 0.6, size = 0.8, col = NA) +
  # geom_vline(
  #  data = medians,
  #  aes(xintercept = median, col = seeding_binomial),
  #  alpha = 1,
  #  size = 1,
  #  linetype = "longdash") +
  scale_x_log10(
    expand = c(0, 0),
    limits = c(0.01, 100),
    labels = c("0.01", "0.1", "1", "10", "100")
  ) +
  scale_y_continuous(
    expand = c(0, 0),
    limits = c(0, 1.25),
    breaks = seq(0, 1.25, by = 0.625)
  ) +
  scale_fill_manual(
    values = c("#eb8775", "#172e6d"),
    labels = c("No", "Yes")
  ) +
  scale_colour_manual(values = c("#eb8775", "#172e6d")) +
  guides(fill = guide_legend(ncol = 2)) +
  labs(
    x = "Fire frequency (per century)",
    y = "Density",
    fill = "Resprouting/seeding"
  ) +
  theme_classic() +
  theme(
    axis.line = element_blank(),
    plot.margin = unit(c(0.5, 0.1, 0.1, 0.1), "cm"),
    panel.border = element_rect(colour = "black", fill = NA),
    panel.spacing = unit(0.5, "lines"),
    axis.text = element_text(size = 10),
    axis.title.y = element_text(size = 16, margin = margin(0, 15, 0, 0)),
    axis.title.x = element_text(size = 16, margin = margin(15, 0, 0, 0)),
    legend.position = "bottom",
    legend.title = element_text(size = 16),
    legend.text = element_text(size = 14)
  )

# Convert the plot to a grob object
grob <- ggplot2::ggplotGrob(seeding_density)
guide_box <- cowplot::get_plot_component(grob, "guide-box", return_all = TRUE)
density_legend <- guide_box[[3]]

seeding_density <-
  df_woody_only %>%
  drop_na(seeding_binomial) %>%
  ggplot() +
  geom_density(
    aes(mean_fires_100, fill = seeding_binomial),
    alpha = 0.6, size = 0.8, col = NA
  ) +
  geom_text(
    data = data_labels,
    aes(x = x, y = y, label = label, colour = label),
    hjust = 0,
    size = 7,
    fontface = "bold"
  ) +
  # geom_vline(
  #  data = medians,
  #  aes(xintercept = median, col = seeding_binomial),
  #  alpha = 1,
  #  size = 1,
  #  linetype = "longdash") +
  scale_x_log10(
    expand = c(0, 0),
    limits = c(0.01, 100),
    labels = c("0.01", "0.1", "1", "10", "100")
  ) +
  scale_y_continuous(
    expand = c(0, 0),
    limits = c(0, 1.25),
    breaks = seq(0, 1.25, by = 0.625),
    labels = c("0", "0.625", "1.25")
  ) +
  scale_fill_manual(
    values = c("#eb8775", "#172e6d"),
    labels = c("No", "Yes")
  ) +
  scale_colour_manual(
    values = c("#5e3a25")
  ) +
  guides(fill = guide_legend(ncol = 2)) +
  labs(
    x = "Fire frequency (per century)",
    y = "Density",
    fill = "Resprouting/seeding"
  ) +
  theme_classic() +
  theme(
    axis.line = element_blank(),
    plot.margin = unit(c(0.5, 0.1, 0.1, 0.1), "cm"),
    panel.border = element_rect(colour = "black", fill = NA),
    panel.spacing = unit(0.5, "lines"),
    axis.text = element_text(size = 10),
    axis.title.y = element_text(size = 16, margin = margin(0, 15, 0, 0)),
    axis.title.x = element_text(size = 16, margin = margin(15, 0, 0, 0)),
    legend.position = "none",
    legend.title = element_text(size = 16),
    legend.text = element_text(size = 14)
  )

seeding_density
```


Plot resprouting and seeding probability graphs together

```{r}
design <- "12\n34\n35"

resprouting_plot + seeding_plot + resprouting_density + seeding_density + density_legend +
  plot_layout(design = design, heights = c(8, 3.75, 2.5, 3.5))

ggsave("figures/fire_response_probabilities_density_functions.png", width = 10, height = 8)
```




### Phylogenetic logistic regression

#### Resprouting binomial

Phylogenetic regression

```{r}
library(phylolm)
library(future)
library(ape)
plan(multisession)

tree <- read.tree("data/v0.1-big-seed-plant-trees/ALLMB.tre")
tree$tip.label <- gsub("_", " ", tree$tip.label)
df_woody_herb_only |> nrow()
df_woody_herb_only <- df_woody_herb_only[df_woody_herb_only$taxon_name %in% tree$tip.label, ]
df_woody_herb_only |> nrow() # Lost 2587 taxa
df_woody_herb_only$resprouting_binomial <- ifelse(df_woody_herb_only$resprouting_binomial == TRUE, 1, 0) # Make numeric
# as the model requires a binary (0 or 1) response

pruned_tree <- drop.tip(tree, setdiff(tree$tip.label, df_woody_herb_only$taxon_name))
pruned_tree <- as.phylo(pruned_tree)
```


With sorted data

```{r}
sorted_data <- df_woody_herb_only[order(match(df_woody_herb_only$taxon_name, pruned_tree$tip.label)), ]

sorted_data <-
  sorted_data %>%
  select(taxon_name, mean_fires_100, resprouting_binomial, woody_or_herb) %>%
  column_to_rownames("taxon_name") |>
  mutate(across(c("woody_or_herb", "resprouting_binomial"), as.factor))

pglm_resprouting <- phyloglm(
  resprouting_binomial ~
    log10(mean_fires_100) + I(log10(mean_fires_100)^2) + log10(mean_fires_100):woody_or_herb +
    I(log10(mean_fires_100)^2):woody_or_herb + woody_or_herb,
  method = "logistic_MPLE",
  data = sorted_data,
  phy = pruned_tree,
  btol = 100,
  boot = 0
)


summary(pglm_resprouting)
sjPlot::tab_model(pglm_resprouting)

modelbased::estimate_prediction(pglm_resprouting) |>
  ggplot(aes(x = mean_fires_100, y = Predicted, color = woody_or_herb)) +
  geom_line(linewidth = 1) +
  scale_x_log10() +
  ylab("Resprouting probability") +
  xlab("Mean fires") +
  theme_classic()
```



```{r}
# coef2 <- coef(pglm_resprouting)
# x1 <- log10(sorted_data$mean_fires_100)
# x2 <- sorted_data$woody_or_herb
# x2 <- ifelse(x2 == "woody", 1, 0)
# pred_y <- coef2[1] + coef2[2] * x1 + coef2[3] * x1^2 + coef2[4] * x2 + coef2[5] * x1 * x2 + coef2[6] * x1^2 * x2
# pred_p <- exp(pred_y) / (1 + exp(pred_y))
# df1 <- data.frame(x1, x2, sorted_data$woody_or_herb, pred_y, pred_p)
# df1 <- df1 %>% rename(woody_or_herb = sorted_data.woody_or_herb)

# resprouting_phylo_plot <-
#   ggplot(data = df1, aes(x = 10^x1, y = pred_p, col = woody_or_herb)) +
#   geom_line(linewidth = 1.3) +
#   scale_x_log10(
#     expand = c(0, 0),
#     limits = c(0.01, 100),
#     labels = c("0.01", "0.1", "1", "10", "100")
#   ) +
#   scale_y_continuous(
#     limits = c(0, 1),
#     expand = c(0, 0)
#   ) +
#   ylab("Resprouting probability") +
#   xlab("Fire frequency (per century)") +
#   scale_colour_manual(
#     values = c("#5e3a25", "#8aa540"),
#     name = "",
#     breaks = c("woody", "herb"),
#     labels = c("Woody", "Herb")
#   ) +
#   ggtitle("a)") +
#   theme_classic() +
#   theme(
#     panel.border = element_rect(colour = "black", fill = NA),
#     legend.position = "none",
#     axis.text = element_text(size = 10),
#     plot.margin = unit(c(0.1, 1, 0.1, 0.1), "cm"),
#     plot.title.position = "plot",
#     plot.title = element_text(size = 22, margin = margin(0, 0, 20, 0)),
#     axis.title.y = element_text(size = 16, margin = margin(0, 15, 0, 0)),
#     axis.title.x = element_text(size = 16, margin = margin(15, 0, 0, 0)),
#     panel.grid.major = element_line(colour = "#ebebeb", linewidth = 0.5),
#     axis.line = element_blank()
#   )
# resprouting_phylo_plot

resprouting_phylo_plot_2 <-
  ggplot(data = df1, aes(x = 10^x1, y = pred_p, col = woody_or_herb)) +
  geom_line(linewidth = 1.3) +
  scale_x_log10(
    expand = c(0, 0),
    limits = c(0.01, 100),
    labels = c("0.01", "0.1", "1", "10", "100")
  ) +
  scale_y_continuous(
    limits = c(0, 1),
    expand = c(0, 0)
  ) +
  ylab("Resprouting probability") +
  xlab("Fire frequency (per century)") +
  scale_colour_manual(
    values = c("#5e3a25", "#8aa540"),
    name = "",
    breaks = c("woody", "herb"),
    labels = c("Woody", "Herb")
  ) +
  ggtitle("a)") +
  theme_classic() +
  theme(
    panel.border = element_rect(colour = "black", fill = NA),
    legend.position = "none",
    axis.text = element_text(size = 10),
    plot.margin = unit(c(0.1, 1, 0.1, 0.1), "cm"),
    plot.title.position = "plot",
    plot.title = element_text(size = 22, margin = margin(0, 0, 20, 0)),
    axis.title.y = element_text(size = 16, margin = margin(0, 15, 0, 0)),
    axis.title.x = element_blank(),
    panel.grid.major = element_line(colour = "#ebebeb", linewidth = 0.5),
    axis.line = element_blank()
  )
resprouting_phylo_plot_2
```


Plot with confidence intervals

```{r}
# Extract model coefficients
coefficients <- coef(pglm_resprouting)

# Create the design matrix for the new data
design_matrix <- model.matrix(~ log10(mean_fires_100) + I(log10(mean_fires_100)^2) + log10(mean_fires_100):woody_or_herb +
  I(log10(mean_fires_100)^2):woody_or_herb + woody_or_herb, data = sorted_data)

# Calculate predicted values on the logit scale
logit_values <- design_matrix %*% coefficients

# Convert logit values to probabilities
predicted_probabilities <- 1 / (1 + exp(-logit_values))

# Calculate standard errors on the logit scale
vcov_matrix <- vcov(pglm_resprouting)
standard_errors_logit <- sqrt(diag(design_matrix %*% vcov_matrix %*% t(design_matrix)))

# Calculate confidence intervals on the logit scale
confidence_level <- 0.95
z_value <- qnorm((1 + confidence_level) / 2)
CI_low_logit <- logit_values - z_value * standard_errors_logit
CI_high_logit <- logit_values + z_value * standard_errors_logit

# Convert confidence intervals to probabilities
CI_low <- 1 / (1 + exp(-CI_low_logit))
CI_high <- 1 / (1 + exp(-CI_high_logit))

# Add predicted probabilities and confidence intervals to the sorted_data dataframe
sorted_data$predicted_resprouting <- predicted_probabilities
sorted_data$CI_low <- CI_low
sorted_data$CI_high <- CI_high

# Plot the predicted probabilities with confidence intervals
(
  resprouting_phylo_plot <-
    ggplot(
      sorted_data,
      aes(x = mean_fires_100, y = predicted_resprouting, color = woody_or_herb)
    ) +
    geom_line(linewidth = 1.3) +
    geom_ribbon(aes(ymin = CI_low, ymax = CI_high, fill = woody_or_herb), alpha = 0.2, linetype = 0) +
    scale_x_log10(
      expand = c(0, 0),
      limits = c(0.01, 100),
      labels = c("0.01", "0.1", "1", "10", "100")
    ) +
    scale_y_continuous(
      limits = c(0, 1),
      expand = c(0, 0)
    ) +
    ylab("Resprouting probability") +
    xlab("Fire frequency (per century)") +
    scale_colour_manual(
      values = c("#5e3a25", "#8aa540"),
      name = "",
      breaks = c("woody", "herb"),
      labels = c("Woody", "Herb")
    ) +
    scale_fill_manual(
      values = c("#b19586", "#9fb846"),
      name = "",
      breaks = c("woody", "herb"),
      labels = c("Woody", "Herb")
    ) +
    ggtitle("a)") +
    theme_classic() +
    theme(
      panel.border = element_rect(colour = "black", fill = NA),
      legend.position = "none",
      axis.text = element_text(size = 10),
      plot.margin = unit(c(0.1, 1, 0.1, 0.1), "cm"),
      plot.title.position = "plot",
      plot.title = element_text(size = 22, margin = margin(0, 0, 20, 0)),
      axis.title.y = element_text(size = 16, margin = margin(0, 15, 0, 0)),
      axis.title.x = element_text(size = 16, margin = margin(15, 0, 0, 0)),
      panel.grid.major = element_line(colour = "#ebebeb", linewidth = 0.5),
      axis.line = element_blank()
    )
)
```

To calculate R2 Tjur

```{r}
pred <- fitted(pglm_resprouting)
y <- pglm_resprouting$y

categories <- unique(y)
mean1 <- mean(pred[which(y == categories[1])], na.rm = TRUE)
mean2 <- mean(pred[which(y == categories[2])], na.rm = TRUE)

abs(mean1 - mean2) # 0.0362
```


Fit logistic curve

```{r}
coef2 <- coef(pglm_resprouting)
plot.new()
plot(
  seq(-100, 100, by = 20),
  seq(0, 1, by = 0.1),
  xlab = "Mean fires",
  ylab = "Resprouting probability"
)

curve(plogis(coef2[1] + coef2[2] * x), col = "red", add = TRUE)

sorted_data |>
  dplyr::filter(!is.na(resprouting_binomial)) |>
  mutate(resprouting_binomial = if_else(resprouting_binomial == 1, 1, 0)) |>
  ggplot(aes(x = mean_fires_100, y = resprouting_binomial)) +
  geom_jitter(height = 0.01, alpha = 0.25) +
  stat_smooth(method = "glm", method.args = list(family = binomial)) +
  facet_wrap(~woody_or_herb) +
  scale_x_log10() +
  theme_classic()
```


#### Seeding binomial


Model without herb data

```{r}
df_woody_only <- df_woody_only[df_woody_only$taxon_name %in% tree$tip.label, ]
df_woody_only$seeding_binomial <- ifelse(df_woody_only$seeding_binomial == TRUE, 1, 0) # make numeric

pruned_tree <- drop.tip(tree, setdiff(tree$tip.label, df_woody_only$taxon_name))
pruned_tree <- as.phylo(pruned_tree)

sorted_data <- df_woody_only[order(match(df_woody_only$taxon_name, pruned_tree$tip.label)), ]

sorted_data <-
  sorted_data %>%
  select(taxon_name, mean_fires_100, seeding_binomial, woody_or_herb) %>%
  column_to_rownames("taxon_name") |>
  mutate(across(c("woody_or_herb", "seeding_binomial"), as.factor))

pglm_seeding <- phyloglm(
  seeding_binomial ~ log10(mean_fires_100) + I(log10(mean_fires_100)^2),
  method = "logistic_MPLE",
  data = sorted_data,
  phy = pruned_tree,
  btol = 100,
  boot = 0
)

summary(pglm_seeding)
sjPlot::tab_model(pglm_seeding)
```


To calculate R2 Tjur

```{r}
pred <- fitted(pglm_seeding)
y <- pglm_seeding$y

categories <- unique(y)
mean1 <- mean(pred[which(y == categories[1])], na.rm = TRUE)
mean2 <- mean(pred[which(y == categories[2])], na.rm = TRUE)

abs(mean1 - mean2) # 0.039
```


Plot

```{r}
# coef2 <- coef(pglm_seeding)
# x1 <- log10(sorted_data$mean_fires_100)
# x2 <- sorted_data$woody_or_herb
# x2 <- ifelse(x2 == "woody", 1, 0)
# pred_y <- coef2[1] + coef2[2] * x1 + coef2[3] * x1^2
# pred_p <- exp(pred_y) / (1 + exp(pred_y))
# df1 <- data.frame(x1, x2, sorted_data$woody_or_herb, pred_y, pred_p)
# df1 <- df1 %>% rename(woody_or_herb = sorted_data.woody_or_herb)

# seeding_phylo_plot <-
#   ggplot(data = df1, aes(x = 10^x1, y = pred_p, col = woody_or_herb)) +
#   geom_line(linewidth = 1.3) +
#   scale_x_log10(
#     expand = c(0, 0),
#     limits = c(0.01, 100),
#     labels = c("0.01", "0.1", "1", "10", "100")
#   ) +
#   scale_y_continuous(
#     limits = c(0, 1),
#     expand = c(0, 0)
#   ) +
#   scale_colour_manual(
#     values = c("#5e3a25"),
#     name = "",
#     breaks = c("woody"),
#     labels = c("Woody")
#   ) +
#   ggtitle("b)") +
#   ylab("Seeding probability") +
#   xlab("Fire frequency (per century)") +
#   theme_classic() +
#   theme(
#     panel.border = element_rect(colour = "black", fill = NA),
#     legend.position = "none",
#     axis.text = element_text(size = 10),
#     plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"),
#     plot.title.position = "plot",
#     plot.title = element_text(size = 22, margin = margin(0, 0, 20, 0)),
#     axis.title.y = element_text(size = 16, margin = margin(0, 15, 0, 0)),
#     axis.title.x = element_text(size = 16, margin = margin(15, 0, 0, 0)),
#     panel.grid.major = element_line(colour = "#ebebeb", linewidth = 0.5),
#     axis.line = element_blank()
#   )

# seeding_phylo_plot


seeding_phylo_plot_2 <-
  ggplot(data = df1, aes(x = 10^x1, y = pred_p, col = woody_or_herb)) +
  geom_line(linewidth = 1.3) +
  scale_x_log10(
    expand = c(0, 0),
    limits = c(0.01, 100),
    labels = c("0.01", "0.1", "1", "10", "100")
  ) +
  scale_y_continuous(
    limits = c(0, 1),
    expand = c(0, 0)
  ) +
  scale_colour_manual(
    values = c("#5e3a25"),
    name = "",
    breaks = c("woody"),
    labels = c("Woody")
  ) +
  ggtitle("b)") +
  ylab("Seeding probability") +
  xlab("Fire frequency (per century)") +
  theme_classic() +
  theme(
    panel.border = element_rect(colour = "black", fill = NA),
    legend.position = "none",
    axis.text = element_text(size = 10),
    plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"),
    plot.title.position = "plot",
    plot.title = element_text(size = 22, margin = margin(0, 0, 20, 0)),
    axis.title.y = element_text(size = 16, margin = margin(0, 15, 0, 0)),
    axis.title.x = element_blank(),
    panel.grid.major = element_line(colour = "#ebebeb", linewidth = 0.5),
    axis.line = element_blank()
  )

seeding_phylo_plot_2
```


Plot with confidence intervals

```{r}
# Extract model coefficients
coefficients <- coef(pglm_seeding)

# Create the design matrix for the new data
design_matrix <- model.matrix(~ log10(mean_fires_100) + I(log10(mean_fires_100)^2), data = sorted_data)

# Calculate predicted values on the logit scale
logit_values <- design_matrix %*% coefficients

# Convert logit values to probabilities
predicted_probabilities <- 1 / (1 + exp(-logit_values))

# Calculate standard errors on the logit scale
vcov_matrix <- vcov(pglm_seeding)
standard_errors_logit <- sqrt(diag(design_matrix %*% vcov_matrix %*% t(design_matrix)))

# Calculate confidence intervals on the logit scale
confidence_level <- 0.95
z_value <- qnorm((1 + confidence_level) / 2)
CI_low_logit <- logit_values - z_value * standard_errors_logit
CI_high_logit <- logit_values + z_value * standard_errors_logit

# Convert confidence intervals to probabilities
CI_low <- 1 / (1 + exp(-CI_low_logit))
CI_high <- 1 / (1 + exp(-CI_high_logit))

# Add predicted probabilities and confidence intervals to the sorted_data dataframe
sorted_data$predicted_seeding <- predicted_probabilities
sorted_data$CI_low <- CI_low
sorted_data$CI_high <- CI_high

# Plot the predicted probabilities with confidence intervals
(
  seeding_phylo_plot <-
    ggplot(
      sorted_data,
      aes(x = mean_fires_100, y = predicted_seeding, color = woody_or_herb)
    ) +
    geom_line(linewidth = 1.3) +
    geom_ribbon(aes(ymin = CI_low, ymax = CI_high, fill = woody_or_herb), alpha = 0.2, linetype = 0) +
    scale_x_log10(
      expand = c(0, 0),
      limits = c(0.01, 100),
      labels = c("0.01", "0.1", "1", "10", "100")
    ) +
    scale_y_continuous(
      limits = c(0, 1),
      expand = c(0, 0)
    ) +
    ylab("Seeding probability") +
    xlab("Fire frequency (per century)") +
    scale_colour_manual(
      values = c("#5e3a25"),
      name = "",
      breaks = c("woody"),
      labels = c("Woody")
    ) +
    scale_fill_manual(
      values = c("#b19586"),
      name = "",
      breaks = c("woody"),
      labels = c("Woody")
    ) +
    ggtitle("a)") +
    theme_classic() +
    theme(
      panel.border = element_rect(colour = "black", fill = NA),
      legend.position = "none",
      axis.text = element_text(size = 10),
      plot.margin = unit(c(0.1, 1, 0.1, 0.1), "cm"),
      plot.title.position = "plot",
      plot.title = element_text(size = 22, margin = margin(0, 0, 20, 0)),
      axis.title.y = element_text(size = 16, margin = margin(0, 15, 0, 0)),
      axis.title.x = element_text(size = 16, margin = margin(15, 0, 0, 0)),
      panel.grid.major = element_line(colour = "#ebebeb", linewidth = 0.5),
      axis.line = element_blank()
    )
)
```

```{r}
resprouting_phylo_plot + seeding_phylo_plot

ggsave("figures/fire_response_probabilities_phyloglm.png", height = 4.5, width = 9)
```


Main figure but with phylo plots

```{r}
design <- "12\n34\n35"

resprouting_phylo_plot_2 + seeding_phylo_plot_2 + resprouting_density + seeding_density + density_legend +
  plot_layout(design = design, heights = c(8, 3.75, 2.5, 3.5))

ggsave("figures/fire_response_probabilities_density_functions_phylo.png", width = 10, height = 8)
```


Check if subsetting to the same taxa as the phyloglm and running a normal lm has the same result

```{r}
df_woody_herb_only <- read_csv("outputs/final_df_predicted_fires_and_fire_response.csv")
df_woody_only <- df_woody_herb_only %>% filter(woody_or_herb == "woody")

df_woody_herb_only <- df_woody_herb_only[df_woody_herb_only$taxon_name %in% tree$tip.label, ]
df_woody_only <- df_woody_only[df_woody_only$taxon_name %in% tree$tip.label, ]

glm_resprouting <- glm(
  resprouting_binomial ~
    log10(mean_fires_100) + I(log10(mean_fires_100)^2) + log10(mean_fires_100):woody_or_herb +
    I(log10(mean_fires_100)^2):woody_or_herb + woody_or_herb,
  data = df_woody_herb_only,
  family = binomial
)

summary(glm_resprouting)
sjPlot::tab_model(glm_resprouting)

ggeffects::ggpredict(glm_resprouting, c("mean_fires_100", "woody_or_herb")) %>%
  ggplot(aes(x, predicted, color = group)) +
  geom_line(linewidth = 1) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group), alpha = 0.2, linetype = 0) +
  scale_x_log10() +
  ylab("Resprouting probability") +
  xlab("Mean fires") +
  theme_classic()

df_woody_only <-
  df_woody_herb_only %>%
  filter(woody_or_herb == "woody") |>
  mutate(across(c("resprouting_binomial", "seeding_binomial", "woody_or_herb"), as.factor))


glm_seeding_woody <- glm(
  seeding_binomial ~ log10(mean_fires_100) + I(log10(mean_fires_100)^2),
  data = df_woody_only,
  family = binomial
)

summary(glm_seeding_woody)
sjPlot::tab_model(glm_seeding_woody)

ggeffects::ggpredict(glm_seeding_woody, c("mean_fires_100")) %>%
  ggplot(aes(x, predicted)) +
  geom_line(linewidth = 1) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2, linetype = 0) +
  scale_x_log10() +
  ylab("Seeding probability") +
  xlab("Mean fires") +
  theme_classic()
```

Increasing the sample size of the phyloglm (by updating the phylogeny tree with APCalign) lead to the same
pattern as the normal glm (but I'm not sure this is valid because that may screw up the relationships in the
phylogeny tree). Then I subsetted the dataset to the taxa used in the phyloglm and ran a normal glm
and it also came out with the same pattern. I think this means that phylogeny doesn't really have an effect
after all, and the specific subset of taxa that are matched to the phylogeny tree doesn't have an effect either.


Test what happens to the results if you only do the species level, not the infraspecific level

```{r}
df_woody_herb_only$taxon_rank |> table()

df_woody_herb_only_sp_only <- df_woody_herb_only |> filter(taxon_rank == "species")

glm_resprouting_sp_only <- glm(
  resprouting_binomial ~
    log10(mean_fires_100) + I(log10(mean_fires_100)^2) + log10(mean_fires_100):woody_or_herb +
    I(log10(mean_fires_100)^2):woody_or_herb + woody_or_herb,
  data = df_woody_herb_only_sp_only,
  family = binomial
)

df_woody_only_sp_only <- df_woody_only |> filter(taxon_rank == "species")

glm_seeding_woody_sp_only <- glm(
  seeding_binomial ~ log10(mean_fires_100) + I(log10(mean_fires_100)^2),
  data = df_woody_only_sp_only,
  family = binomial
)

summary(glm_resprouting)
summary(glm_resprouting_sp_only)

summary(glm_seeding_woody)
summary(glm_seeding_woody_sp_only)

# How many cases do the subspecies have different traits to the species?
small_df <- df_woody_herb_only |>
  select(all_of(c(
    "taxon_name", "mean_fires_100", "taxon_rank", "resprouting_binomial",
    "seeding_binomial", "num_pixels"
  )))

species_only <- small_df |>
  filter(taxon_rank == "species") |>
  rename(all_of(c(
    "mean_fires_100_sp" = "mean_fires_100",
    "resprouting_binomial_sp" = "resprouting_binomial",
    "seeding_binomial_sp" = "seeding_binomial",
    "num_pixels_sp" = "num_pixels",
    "species" = "taxon_name"
  ))) |>
  select(-"taxon_rank")

infra_only <- small_df |>
  filter(taxon_rank != "species") |>
  mutate(species = str_extract(taxon_name, "^\\S+\\s+\\S+")) |>
  left_join(species_only, by = "species") |>
  mutate(
    mean_fires_100_diff = mean_fires_100_sp - mean_fires_100,
    resprouting_diff = if_else(resprouting_binomial != resprouting_binomial_sp, TRUE, FALSE),
    seeding_diff = if_else(seeding_binomial != seeding_binomial_sp, TRUE, FALSE)
  )

infra_only |> nrow() # 1416

table(infra_only$resprouting_diff) # 971 FALSE, 143 TRUE (12.8% where data available for both species and infraspecies)
infra_only |>
  filter(!is.na(resprouting_binomial)) |>
  nrow() # 1389
infra_only |>
  filter(!is.na(resprouting_binomial), !is.na(resprouting_binomial_sp)) |>
  nrow() # 1114

table(infra_only$seeding_diff) # 192 FALSE, 43 TRUE (18.3% where data available for both species and infraspecies)
infra_only |>
  filter(!is.na(seeding_binomial)) |>
  nrow() # 394
infra_only |>
  filter(!is.na(seeding_binomial), !is.na(seeding_binomial_sp)) |>
  nrow() # 235

# How different are the fire frequencies experienced by the subspecies?
(
  infra_diffs_hist <- ggplot(infra_only) +
    geom_histogram(aes(x = mean_fires_100_diff), colour = "#739c3e", fill = "#739c3e") +
    scale_x_continuous(
      expand = c(0, 0),
      limits = c(-20, 30)
    ) +
    scale_y_continuous(
      expand = c(0, 0),
      limits = c(0, 800)
    ) +
    labs(
      x = "Difference in mean fire frequency (per century)",
      y = "Count"
    ) +
    theme_classic() +
    theme(
      plot.margin = unit(c(0.5, 1, 0.5, 0.5), "cm"),
      axis.text = element_text(size = 10, colour = "black"),
      axis.title.y = element_text(size = 16, margin = margin(0, 15, 0, 0)),
      axis.title.x = element_text(size = 16, margin = margin(15, 0, 0, 0))
    )
)

ggsave("figures/infra_diffs_hist.png", width = 7, height = 5.5)
sum(abs(na.omit(infra_only$mean_fires_100_diff)) > 3) # 119
```

In conclusion, running with species only does not change the relationships.

---

## Comparison of survival analysis and poisson regression methods


### Survival analysis method

Read in median fire return interval data from survival analysis

```{r}
median_fris <- read_csv("outputs/median_fris_with_survival_analysis.csv")
median_fris <- median_fris %>% left_join(resprouting_seeding, by = c("taxon_name"))

median_fris |>
  dplyr::filter(num_pixels >= 10) |>
  nrow()
median_fris |>
  dplyr::filter(!is.na(med_w_warnings), num_pixels >= 10) |>
  nrow()

median_fires_weibull <- median_fris %>%
  dplyr::filter(is.na(med_w_warnings), num_pixels >= 10) %>%
  mutate(across(
    c("med_w", "med_w_lwr", "med_w_upr", "med_wo", "med_wo_lwr", "med_wo_upr"),
    ~ .x / 365.2422 # Convert intervals to years not days
  )) |>
  mutate(across(
    c("med_w", "med_w_lwr", "med_w_upr", "med_wo", "med_wo_lwr", "med_wo_upr"),
    ~ 100 / .x # Convert to fire frequency per century
  )) |>
  mutate(confint_size = med_w_lwr - med_w_upr) |> # Lower minus upper because original FRI was upper minus lower
  select(all_of(c(
    "taxon_name", "med_w", "med_w_lwr", "med_w_upr", "confint_size", "num_unburnt_pixels", "num_pixels",
    "resprouting_binomial", "seeding_binomial", "data_on_both", "woody_or_herb", "taxon_rank"
  ))) |>
  dplyr::filter(confint_size < 30) # Filter out confidence intervals greater than 30 years
```

With the survival analysis method, 1099 models did not converge (incl. unburnt
pixels) out of 10406 species with GBIF occurrences >= 10. The warning was "ran out of iterations and did not converge". This occurred
because most or all pixels were unburnt.


Histogram of median fire frequencies and confidence intervals

```{r}
hist(median_fires_weibull$med_w, breaks = 30)
hist(median_fires_weibull$confint_size, breaks = 30)
```


### Poisson regression method

```{r}
mean_fires <- read_csv("outputs/mean_fires_df_with_poisson_glm.csv", guess_max = 10871)
mean_fires <- mean_fires %>% left_join(resprouting_seeding, by = c("taxon_name"))

mean_fires |>
  dplyr::filter(num_pixels >= 10) |>
  nrow()
mean_fires |>
  dplyr::filter(num_pixels >= 10) |>
  dplyr::filter(!is.na(warnings) | !is.na(messages) | !is.na(error)) |>
  nrow()

mean_fires_poisson <- mean_fires |>
  # Convert to mean number of fires in 100 years, instead of mean number of fires in the sampling period
  mutate(across(
    c("mean_fires", "upr_conf_int", "lwr_conf_int"),
    ~ .x / 22.17 * 100
  )) |>
  dplyr::filter(is.na(warnings), is.na(error), is.na(messages), num_pixels >= 10) |>
  mutate(confint_size = upr_conf_int - lwr_conf_int) |>
  select(-all_of(c(
    "mean_fri", "warnings", "messages", "error", "resprouting", "post_fire_seeding",
    "resprouter_or_seeder", "woodiness_detailed", "obligate_resprouting_or_seeding",
    "resprouting_seeding_binomial"
  )))

hist(mean_fires_poisson$mean_fires, breaks = 30)
hist(mean_fires_poisson$confint_size, breaks = 30)
```


In the Poisson regression method, out of 10406 species with GBIF occurrences >= 10, 166 did not fit and they were all because all pixels were unburnt.


Join data from the two methods

```{r}
method_comparison <- mean_fires_poisson |>
  select(-c("confint_size", "upr_conf_int", "lwr_conf_int", "dispersion_ratio", "dispersion_p_value")) |>
  mutate(method = "poisson") |>
  rename("avg_fires" = "mean_fires") |>
  full_join(
    median_fires_weibull |>
      select(-c("confint_size", "med_w_lwr", "med_w_upr")) |>
      mutate(method = "survival") |>
      rename("avg_fires" = "med_w")
  )
```


Plot methods against each other

```{r}
method_comparison |>
  ggplot() +
  geom_density(
    aes(avg_fires, fill = method),
    alpha = 0.6,
    size = 0.8,
    col = NA
  ) +
  scale_x_log10(
    expand = c(0, 0),
    limits = c(0.001, 100),
    breaks = c(0.001, 0.01, 0.1, 1, 10, 100),
    labels = c("0.001", "0.01", "0.1", "1", "10", "100")
  ) +
  scale_y_continuous(
    expand = c(0, 0),
    limits = c(0, 1)
  ) +
  scale_fill_manual(
    values = c("#d1b2eb", "#5e804a"),
    labels = c("Poisson regression", "Survival analysis")
  ) +
  labs(
    x = "Fire frequency (per century)",
    y = "Density",
    fill = "Method"
  ) +
  theme_classic() +
  theme(
    axis.line = element_blank(),
    strip.background = element_blank(),
    strip.text = element_blank(),
    plot.margin = unit(c(0.5, 1, 0.1, 0.1), "cm"),
    panel.border = element_rect(colour = "black", fill = NA),
    panel.spacing = unit(0.5, "lines"),
    axis.text = element_text(size = 10),
    axis.title.y = element_text(size = 16, margin = margin(0, 15, 0, 0)),
    axis.title.x = element_text(size = 16, margin = margin(15, 0, 0, 0)),
    legend.title = element_text(size = 15),
    legend.text = element_text(size = 13),
    legend.position = "bottom"
  )

ggsave("figures/method_comparison.png", width = 7, height = 4)
```


```{r}
method_comparison |>
  filter(woody_or_herb %in% c("woody", "herb")) |>
  mutate(woody_or_herb = if_else(woody_or_herb == "woody", "Woody", "Herbs")) |>
  ggplot() +
  geom_density(
    aes(avg_fires, fill = method),
    alpha = 0.6,
    size = 0.8,
    col = NA
  ) +
  facet_wrap(
    ~ factor(woody_or_herb, ordered = TRUE, levels = c("Woody", "Herbs")),
    ncol = 1
  ) +
  scale_x_log10(
    expand = c(0, 0),
    limits = c(0.001, 100),
    breaks = c(0.001, 0.01, 0.1, 1, 10, 100),
    labels = c("0.001", "0.01", "0.1", "1", "10", "100")
  ) +
  scale_y_continuous(
    expand = c(0, 0),
    limits = c(0, 1)
  ) +
  scale_fill_manual(
    values = c("#d1b2eb", "#5e804a"),
    labels = c("Poisson regression", "Survival analysis")
  ) +
  labs(
    x = "Fire frequency (per century)",
    y = "Density",
    fill = "Method"
  ) +
  theme_classic() +
  theme(
    strip.background = element_blank(),
    strip.text = element_text(size = 17, hjust = 0, margin = margin(10, 0, 10, 5), face = "bold"),
    axis.line = element_blank(),
    plot.margin = unit(c(0.5, 1, 0.1, 0.1), "cm"),
    panel.border = element_rect(colour = "black", fill = NA),
    panel.spacing = unit(0.5, "lines"),
    axis.text = element_text(size = 12),
    axis.title.y = element_text(size = 18, margin = margin(0, 15, 0, 0)),
    axis.title.x = element_text(size = 18, margin = margin(15, 0, 0, 0)),
    legend.title = element_text(size = 17),
    legend.text = element_text(size = 15),
    legend.position = "bottom"
  )

ggsave("figures/method_comparison_woody_herb.png", width = 7, height = 7)
```


## Summary statistics of fire frequency for woody and herbaceous species


Woody resprouters and non-resprouters

```{r}
df_woody_herb_only <-
  mean_fires_clean %>%
  dplyr::filter(woody_or_herb %in% c("woody", "herb")) |>
  mutate(across(c("resprouting_binomial", "seeding_binomial", "woody_or_herb"), as.factor))

df_woody_herb_only |>
  filter(woody_or_herb == "woody", resprouting_binomial == TRUE) |>
  summarise(median_fires = median(mean_fires_100), sd = sd(mean_fires_100))

df_woody_herb_only |>
  filter(woody_or_herb == "woody", resprouting_binomial == FALSE) |>
  summarise(median_fires = median(mean_fires_100), sd = sd(mean_fires_100))

wilcox.test(mean_fires_100 ~ resprouting_binomial, data = df_woody_herb_only, subset = woody_or_herb == "woody")
```

Herbaceous resprouters and non-resprouters

```{r}
df_woody_herb_only |>
  filter(woody_or_herb == "herb", resprouting_binomial == TRUE) |>
  summarise(median_fires = median(mean_fires_100), sd = sd(mean_fires_100))

df_woody_herb_only |>
  filter(woody_or_herb == "herb", resprouting_binomial == FALSE) |>
  summarise(median_fires = median(mean_fires_100), sd = sd(mean_fires_100))

wilcox.test(mean_fires_100 ~ resprouting_binomial, data = df_woody_herb_only, subset = woody_or_herb == "herb")
```

Woody seeders and non-seeders

```{r}
df_woody_herb_only |>
  filter(woody_or_herb == "woody", seeding_binomial == TRUE) |>
  summarise(median_fires = median(mean_fires_100), sd = sd(mean_fires_100))

df_woody_herb_only |>
  filter(woody_or_herb == "woody", seeding_binomial == FALSE) |>
  summarise(median_fires = median(mean_fires_100), sd = sd(mean_fires_100))
```


Median fire frequency experienced by woody and herbaceous plants

```{r}
df_woody_herb_only |>
  filter(woody_or_herb == "woody") |>
  summarise(
    max_fires = max(mean_fires_100),
    min_fires = min(mean_fires_100),
    median_fires = median(mean_fires_100),
    sd = sd(mean_fires_100)
  )

df_woody_herb_only |>
  filter(woody_or_herb == "herb") |>
  summarise(
    max_fires = max(mean_fires_100),
    min_fires = min(mean_fires_100),
    median_fires = median(mean_fires_100),
    sd = sd(mean_fires_100)
  )

# Highest fire frequency species
df_woody_herb_only |>
  arrange(-mean_fires_100) |>
  head(10) |>
  View()
```


Fire frequencies of facultative herbs

```{r}
df_woody_herb_only |>
  filter(woody_or_herb == "herb", resprouting_binomial == TRUE, seeding_binomial == TRUE) |>
  View()

# Looks like most of these are rarely burnt species
# So maybe supports the hypothesis that in moderately or frequently burnt systems, herbs are
# usually only a resprouter or a seeder, not both?
```


Investigate facultative species and post-fire seeding in frequently burnt species

```{r}
df_woody_herb_only |>
  filter(mean_fires_100 > 25, woody_or_herb == "woody", !is.na(seeding_binomial)) |>
  pull(seeding_binomial) |>
  length() # 68
df_woody_herb_only |>
  filter(mean_fires_100 > 25, woody_or_herb == "woody", !is.na(seeding_binomial)) |>
  pull(seeding_binomial) |>
  table() # 41 false, 27 true
df_woody_herb_only |>
  filter(mean_fires_100 > 25, woody_or_herb == "woody", !is.na(seeding_binomial), seeding_binomial == TRUE, resprouting_binomial == TRUE) |>
  nrow() # 25
df_woody_herb_only |>
  filter(mean_fires_100 > 25, woody_or_herb == "woody", !is.na(seeding_binomial), seeding_binomial == TRUE, resprouting_binomial == FALSE) |>
  nrow() # 2
```

When I filter to plants that have mean fire frequency per century of > 25, are woody and have seeding data available,
40% are seeders and of which 93% are also resprouters (facultative species).


Rerun analyses without annuals

```{r}
df_woody_herb_only_no_annuals <- df_woody_herb_only_life_history |>
  filter(annual_perennial != "annual")

glm_resprouting <- glm(
  resprouting_binomial ~
    log10(mean_fires_100) + I(log10(mean_fires_100)^2) + log10(mean_fires_100):woody_or_herb +
    I(log10(mean_fires_100)^2):woody_or_herb + woody_or_herb,
  data = df_woody_herb_only_no_annuals,
  family = binomial
)

summary(glm_resprouting)

ggeffects::ggpredict(glm_resprouting, c("mean_fires_100", "woody_or_herb")) %>%
  ggplot(aes(x, predicted, color = group)) +
  geom_line(linewidth = 1) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group), alpha = 0.2, linetype = 0) +
  scale_x_log10() +
  ylab("Resprouting probability") +
  xlab("Mean fires") +
  theme_classic()

glm_resprouting_herb <- glm(
  resprouting_binomial ~
    log10(mean_fires_100) + I(log10(mean_fires_100)^2) + log10(mean_fires_100):woody_or_herb +
    I(log10(mean_fires_100)^2):woody_or_herb + woody_or_herb,
  data = df_woody_herb_only_no_annuals |> filter(woody_or_herb == "herb"),
  family = binomial
)

summary(glm_resprouting_herb)
ggeffects::ggpredict(glm_resprouting_herb, c("mean_fires_100", "woody_or_herb")) %>%
  ggplot(aes(x, predicted, color = group)) +
  geom_line(linewidth = 1) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group), alpha = 0.2, linetype = 0) +
  scale_x_log10() +
  ylab("Resprouting probability") +
  xlab("Mean fires") +
  theme_classic()
```
