---
title: "Phylogeny Figure"
output: html_notebook
editor_options:
  chunk_output_type: console
---



Load libraries

```{r}
library(tidyverse)
library(patchwork)
library(ggtreeExtra)
library(ggtree)
```


### Resprouting data

Join to lookup table

```{r}
resprouting_seeding <- readRDS("data/resprouting_seeding.rds")

austraits <- read_rds("data/austraits-5.0.0.rds")

resprouting_seeding_new <- resprouting_seeding |>
  left_join(
    austraits$taxa |> select(all_of(c("taxon_name", "genus", "family"))),
    by = "taxon_name"
  )

resprouting_seeding_new$family %>% unique() %>% length() # 223 unique families
x <- resprouting_seeding_new %>% group_by(family) %>% summarise(count = n())
hist(x$count, breaks = 100)
```


Summarise resprouting by family

```{r}
resprouting_summary <-
  resprouting_seeding_new %>%
  drop_na(resprouting_binomial) %>%
  group_by(family) %>%
  summarise(
    prop_resprouting = sum(resprouting_binomial) / n(),
    data_count = n()
  ) %>%
  ungroup() |>
  arrange(-data_count) |>
  slice_head(n = 20)
```


Seeding data

```{r}
seeding_summary <-
  resprouting_seeding_new %>%
  drop_na(seeding_binomial, family) %>%
  group_by(family) %>%
  summarise(prop_seeding = sum(seeding_binomial) / n()) %>%
  ungroup()
```


Mean fire frequency data

```{r}
mean_fires <- read_csv("outputs/final_df_predicted_fires_and_fire_response.csv")
```


```{r}
mean_fires_new <-
  mean_fires %>%
  left_join(
    austraits$taxa |> select(all_of(c("taxon_name", "genus", "family"))),
    by = "taxon_name"
  )

median_fires_summary <-
  mean_fires_new %>%
  group_by(family) %>%
  summarise(median_fires_100 = median(mean_fires_100)) %>%
  ungroup()
```



TODO Add frequency distribution of fire frequencies to figure

```{r}
(
  fires_histogram <-
    mean_fires_new |>
    ggplot(aes(x = mean_fires_100)) +
    geom_histogram(colour = "#517226", fill = "#517226") +
    scale_x_log10(
      expand = c(0, 0),
      limits = c(0.01, 100),
      labels = c("0.01", "0.1", "1", "10", "100")
    ) +
    scale_y_continuous(
      expand = c(0, 0),
      limits = c(0, 1500)
    ) +
    labs(
      title = "a)",
      x = "Mean fire frequency (per century)",
      y = "Count"
    ) +
    theme_classic() +
    theme(
      plot.title.position = "plot",
      plot.title = element_text(size = 22, margin = margin(0, 0, 20, 0)),
      plot.margin = unit(c(0.5, 1, 0.5, 0.5), "cm"),
      axis.text = element_text(size = 10, colour = "black"),
      axis.title.y = element_text(size = 16, margin = margin(0, 15, 0, 0)),
      axis.title.x = element_text(size = 16, margin = margin(15, 0, 0, 0))
    )
)

ggsave("figures/mean_fire_frequency_histogram.png", width = 10, height = 4)
```



**Summary statistics for manuscript**

Range and mean fire frequencies in the dataset

TODO Look at more summary statistics

```{r}
mean_fires$mean_fires_100 |> min()
mean_fires$mean_fires_100 |> max()
mean_fires$mean_fires_100 |> median()
median_fires_summary$median_fires_100 |> min()
median_fires_summary$median_fires_100 |> max()
median_fires_summary |>
  filter(family %in% c("Fabaceae", "Myrtaceae", "Orchidaceae", "Proteaceae", "Asteraceae", "Poaceae")) |>
  arrange(-median_fires_100)

resprouting_summary |> filter(family == "Asteraceae")
```


LMA data

```{r}
aus_native_lookup <- read_csv("data/aus_native_lookup.csv")

lma <-
  austraits$traits %>%
  filter(trait_name == "leaf_mass_per_area") %>%
  pivot_wider(names_from = "trait_name", values_from = "value") |>
  mutate(leaf_mass_per_area = as.numeric(leaf_mass_per_area))

# Filter out non-native taxa
lma <-
  lma %>%
  left_join(
    aus_native_lookup,
    by = c("taxon_name" = "canonical_name")
  ) %>%
  filter(aus_native == TRUE) %>%
  select(-aus_native)

mean_lma <-
  lma %>%
  group_by(taxon_name) %>%
  summarise(mean_lma = mean(leaf_mass_per_area)) %>%
  ungroup()

mean_lma_new <-
  mean_lma %>%
  left_join(
    austraits$taxa |> select(all_of(c("taxon_name", "genus", "family"))),
    by = "taxon_name"
  )

mean_lma_new$family %>% unique() %>% length() # 186
x <- mean_lma_new %>% group_by(family) %>% summarise(count = n())
hist(x$count, breaks = 100)

median_lma_summary <-
  mean_lma_new %>%
  drop_na(mean_lma, family) %>%
  group_by(family) %>%
  summarise(median_lma = median(mean_lma)) %>%
  ungroup()
```


Leaf N per dry mass data

```{r}
leaf_n <-
  austraits$traits %>%
  filter(trait_name == "leaf_N_per_dry_mass") %>%
  pivot_wider(names_from = "trait_name", values_from = "value") |>
  mutate(leaf_N_per_dry_mass = as.numeric(leaf_N_per_dry_mass))


# Filter out non-native taxa
leaf_n <-
  leaf_n %>%
  left_join(
    aus_native_lookup,
    by = c("taxon_name" = "canonical_name")
  ) %>%
  filter(aus_native == TRUE) %>%
  select(-aus_native)

mean_leaf_n <-
  leaf_n %>%
  group_by(taxon_name) %>%
  summarise(mean_leaf_n = mean(leaf_N_per_dry_mass)) %>%
  ungroup()

mean_leaf_n_new <-
  mean_leaf_n %>%
  left_join(
    austraits$taxa |> select(all_of(c("taxon_name", "genus", "family"))),
    by = "taxon_name"
  )

mean_leaf_n_new$family %>% unique() %>% length() # 145
x <- mean_leaf_n_new %>% group_by(family) %>% summarise(count = n())
hist(x$count, breaks = 100)

median_leaf_n_summary <-
  mean_leaf_n_new %>%
  drop_na(mean_leaf_n, family) %>%
  group_by(family) %>%
  summarise(median_leaf_n = median(mean_leaf_n)) %>%
  ungroup()
```


Number of woody species

```{r}
woodiness <-
  austraits$traits %>%
  filter(dataset_id == "Wenk_2022", trait_name == "woodiness_detailed") %>%
  pivot_wider(values_from = value, names_from = trait_name)

woody <- c("woody", "semi_woody woody", "woody woody_base")
herbaceous <- c(
  "herbaceous", "herbaceous woody", "herbaceous woody_base", "herbaceous woody_root",
  "herbaceous semi_woody", "herbaceous semi_woody woody"
)
semi_woody <- c("semi_woody", "semi_woody woody_base")
unclear <- c("woody_like_stem", "woody_base")


# Filter out non-native taxa
woodiness <-
  woodiness %>%
  left_join(
    aus_native_lookup,
    by = c("taxon_name" = "canonical_name")
  ) %>%
  filter(aus_native == TRUE) %>%
  select(-aus_native)

woodiness_new <-
  woodiness %>%
  left_join(
    austraits$taxa |> select(all_of(c("taxon_name", "genus", "family"))),
    by = "taxon_name"
  )

woodiness_new <-
  woodiness_new %>%
  mutate(
    woody_or_herb = case_when(
      woodiness_detailed %in% woody ~ "woody",
      woodiness_detailed %in% herbaceous ~ "herb",
      woodiness_detailed %in% semi_woody ~ "semi-woody",
      woodiness_detailed %in% unclear ~ "unclear"
    )
  )

woodiness_summary <-
  woodiness_new %>%
  drop_na(woody_or_herb) %>%
  group_by(family) %>%
  summarise(
    prop_woody = sum(woody_or_herb == "woody") / n()
  )
```


---

## Pruned tree with only families with resprouting and/or seeding data

```{r}
tt <- ape::read.tree("data/v0.1-big-seed-plant-trees/ALLMB.tre")

a <- taxonlookup::lookup_table(unique(tt$tip.label), by_species = TRUE, missing_action = "NA") |>
  select(-family, -order, -group) # Get genus
taxonlookup_table <- taxonlookup::lookup_table(unique(tt$tip.label), by_species = TRUE, missing_action = "NA") |>
  select(genus, family) |>
  distinct() |>
  rename(all_of(c("taxonlookup_genus" = "genus", "taxonlookup_family" = "family")))
a <- rownames_to_column(a, "species")
a <- a %>%
  left_join(
    austraits$taxa |>
      filter(taxonomic_status == "accepted") |>
      select(all_of(c("genus", "family"))) |> # Get family from AusTraits taxa table
      distinct(),
    by = "genus"
  ) |>
  left_join(taxonlookup_table, by = c("genus" = "taxonlookup_genus")) |>
  mutate(
    family = if_else(is.na(family), taxonlookup_family, family)
  ) |>
  select(-taxonlookup_family)
take_tenth <- function(x) x[10] # Not [1] because the first species of some families is taxonomically ambiguous
one_tip <- a |> group_by(family) %>% summarise(species = take_tenth(species))
out <- ape::drop.tip(tt, tt$tip.label[!tt$tip.label %in% one_tip$species])
out$tip.label <- one_tip$family[match(out$tip.label, one_tip$species)]

out_pruned <- ape::drop.tip(out, out$tip.label[!out$tip.label %in% resprouting_summary$family])

ggtree(out_pruned, ladderize = FALSE) +
  geom_tiplab() +
  xlim(0, 300)
```

Number of native Australian taxa

```{r}
taxonomic_resources <- APCalign::load_taxonomic_resources()

apc <- taxonomic_resources$APC |>
  filter(
    taxonomic_status == "accepted",
    taxon_rank %in% c("species", "form", "variety", "subspecies")
  )

native_lookup <- APCalign::native_anywhere_in_australia(apc$canonical_name, taxonomic_resources)

apc_native <-
  apc %>%
  left_join(native_lookup, by = c("canonical_name" = "species"))

species_diversity <-
  apc_native %>%
  filter(
    taxon_rank == "species", # Filter for species only
    native_anywhere_in_aus == "native" # Filter for native species only
  ) %>%
  group_by(family) %>%
  summarise(n_species = n_distinct(canonical_name)) %>%
  ungroup() %>%
  arrange(-n_species)

species_diversity
```



New data frame to label and colour key families

```{r}
extra_data <- tibble(label = out_pruned$tip.label)

extra_data <-
  extra_data %>%
  left_join(species_diversity, by = c("label" = "family"))

new_tree <- full_join(out_pruned, extra_data, by = "label")
```



```{r}
new_tree@phylo |>
  as_tibble() |>
  filter(label %in% c("Fabaceae", "Myrtaceae", "Orchidaceae", "Proteaceae", "Asteraceae", "Poaceae"))
```

```{r}
ggtree(new_tree, layout = "rectangular", size = 0.7) +
  geom_tiplab(
    aes(subset = (!node %in% c(3, 7, 10, 15, 16, 20))),
    offset = 3,
    size = 6,
    show.legend = FALSE
  ) +
  geom_tiplab(
    aes(subset = (node %in% c(3, 7, 10, 15, 16, 20)), fontface = "bold"),
    offset = 3,
    size = 6,
    show.legend = FALSE
  ) +
  geom_text(
    aes(label = n_species),
    hjust = 1,
    nudge_x = 0,
    nudge_y = -0.25,
    size = 4.5
  ) +
  labs(title = "b)") +
  theme(
    plot.margin = unit(c(5, 5, 5, 5), "mm"),
    plot.title.position = "plot",
    plot.title = element_text(size = 22, margin = margin(0, 0, 20, 0)),
    legend.position = "none",
    legend.margin = margin(-0, 0, 0, 0, unit = "cm"),
    legend.text = element_text(size = 8.5),
    legend.justification = c("right"),
    legend.box.margin = margin(0, 0, 0, -20)
  ) +
  geom_fruit(
    data = median_fires_summary,
    geom = geom_tile,
    mapping = aes(x = median_fires_100, y = family, fill = median_fires_100),
    offset = 0.7,
    pwidth = 0,
    width = 14,
    height = 1
  ) +
  scale_fill_gradient(
    low = "white",
    high = "#7d7d7d"
  ) +
  geom_fruit(
    data = median_fires_summary,
    geom = geom_text,
    mapping = aes(x = median_fires_100, y = family, label = round(median_fires_100, digits = 1)),
    colour = "black",
    offset = 0,
    pwidth = 0
  ) +
  ggnewscale::new_scale_fill() +
  geom_fruit(
    data = resprouting_summary,
    geom = geom_tile,
    mapping = aes(x = prop_resprouting, y = family, fill = prop_resprouting),
    offset = 0.1,
    pwidth = 0,
    width = 14,
    height = 1
  ) +
  scale_fill_gradient(
    low = "white",
    high = "#7d7d7d"
  ) +
  geom_fruit(
    data = resprouting_summary,
    geom = geom_text,
    mapping = aes(x = prop_resprouting, y = family, label = round(prop_resprouting, digits = 1)),
    colour = "black",
    offset = 0,
    pwidth = 0
  ) +
  ggnewscale::new_scale_fill() +
  geom_fruit(
    data = seeding_summary,
    geom = geom_tile,
    mapping = aes(x = prop_seeding, y = family, fill = prop_seeding),
    offset = 0.1,
    pwidth = 0,
    width = 14,
    height = 1
  ) +
  scale_fill_gradient(
    low = "white",
    high = "#7d7d7d"
  ) +
  geom_fruit(
    data = seeding_summary,
    geom = geom_text,
    mapping = aes(x = prop_seeding, y = family, label = round(prop_seeding, digits = 1)),
    colour = "black",
    offset = 0,
    pwidth = 0
  ) +
  ggnewscale::new_scale_fill() +
  geom_fruit(
    data = median_lma_summary,
    geom = geom_tile,
    mapping = aes(x = median_lma, y = family, fill = median_lma),
    offset = 0.1,
    pwidth = 0,
    width = 14,
    height = 1
  ) +
  scale_fill_gradient(
    low = "white",
    high = "#7d7d7d"
  ) +
  geom_fruit(
    data = median_lma_summary,
    geom = geom_text,
    mapping = aes(x = median_lma, y = family, label = round(median_lma, digits = 0)),
    colour = "black",
    offset = 0,
    pwidth = 0
  ) +
  ggnewscale::new_scale_fill() +
  geom_fruit(
    data = median_leaf_n_summary,
    geom = geom_tile,
    mapping = aes(x = median_leaf_n, y = family, fill = median_leaf_n),
    offset = 0.1,
    pwidth = 0,
    width = 14,
    height = 1
  ) +
  scale_fill_gradient(
    low = "white",
    high = "#7d7d7d"
  ) +
  geom_fruit(
    data = median_leaf_n_summary,
    geom = geom_text,
    mapping = aes(x = median_leaf_n, y = family, label = round(median_leaf_n, digits = 1)),
    colour = "black",
    offset = 0,
    pwidth = 0
  ) +
  ggnewscale::new_scale_fill() +
  geom_fruit(
    data = woodiness_summary,
    geom = geom_tile,
    mapping = aes(x = prop_woody, y = family, fill = prop_woody),
    offset = 0.1,
    pwidth = 0,
    width = 14,
    height = 1
  ) +
  scale_fill_gradient(
    low = "white",
    high = "#7d7d7d"
  ) +
  geom_fruit(
    data = woodiness_summary,
    geom = geom_text,
    mapping = aes(x = prop_woody, y = family, label = round(prop_woody, digits = 1)),
    colour = "black",
    offset = 0,
    pwidth = 0
  )

ggsave("figures/phylogeny_figure_greyscale.png", width = 10, height = 9)
```


With colour

```{r}
(
  phylogeny_tree <- ggtree(new_tree, layout = "rectangular", size = 0.7) +
    geom_tiplab(
      aes(subset = (!node %in% c(3, 7, 10, 15, 16, 20))),
      offset = 3,
      size = 6,
      show.legend = FALSE
    ) +
    geom_tiplab(
      aes(subset = (node %in% c(3, 7, 10, 15, 16, 20)), fontface = "bold"),
      offset = 3,
      size = 6,
      show.legend = FALSE
    ) +
    geom_text(
      aes(label = n_species),
      hjust = 1,
      nudge_x = 0,
      nudge_y = -0.25,
      size = 4.5
    ) +
    labs(title = "b)") +
    theme(
      plot.margin = unit(c(5, 5, 5, 5), "mm"),
      plot.title.position = "plot",
      plot.title = element_text(size = 22, margin = margin(0, 0, 20, 0)),
      legend.position = "none",
      legend.margin = margin(-0, 0, 0, 0, unit = "cm"),
      legend.text = element_text(size = 8.5),
      legend.justification = c("right"),
      legend.box.margin = margin(0, 0, 0, -20)
    ) +
    geom_fruit(
      data = median_fires_summary,
      geom = geom_tile,
      mapping = aes(x = median_fires_100, y = family, fill = median_fires_100),
      offset = 0.7,
      pwidth = 0,
      width = 14,
      height = 1
    ) +
    scale_fill_gradient(
      low = "white",
      high = "#e04d3d"
    ) +
    geom_fruit(
      data = median_fires_summary,
      geom = geom_text,
      mapping = aes(x = median_fires_100, y = family, label = round(median_fires_100, digits = 1)),
      colour = "black",
      offset = 0,
      pwidth = 0
    ) +
    ggnewscale::new_scale_fill() +
    geom_fruit(
      data = resprouting_summary,
      geom = geom_tile,
      mapping = aes(x = prop_resprouting, y = family, fill = prop_resprouting),
      offset = 0.1,
      pwidth = 0,
      width = 14,
      height = 1
    ) +
    scale_fill_gradient(
      low = "white",
      high = "#e384b8"
    ) +
    geom_fruit(
      data = resprouting_summary,
      geom = geom_text,
      mapping = aes(x = prop_resprouting, y = family, label = round(prop_resprouting, digits = 1)),
      colour = "black",
      offset = 0,
      pwidth = 0
    ) +
    ggnewscale::new_scale_fill() +
    geom_fruit(
      data = seeding_summary,
      geom = geom_tile,
      mapping = aes(x = prop_seeding, y = family, fill = prop_seeding),
      offset = 0.1,
      pwidth = 0,
      width = 14,
      height = 1
    ) +
    scale_fill_gradient(
      low = "white",
      high = "#e3b13d"
    ) +
    geom_fruit(
      data = seeding_summary,
      geom = geom_text,
      mapping = aes(x = prop_seeding, y = family, label = round(prop_seeding, digits = 1)),
      colour = "black",
      offset = 0,
      pwidth = 0
    ) +
    ggnewscale::new_scale_fill() +
    geom_fruit(
      data = median_lma_summary,
      geom = geom_tile,
      mapping = aes(x = median_lma, y = family, fill = median_lma),
      offset = 0.1,
      pwidth = 0,
      width = 14,
      height = 1
    ) +
    scale_fill_gradient(
      low = "white",
      high = "#66914d"
    ) +
    geom_fruit(
      data = median_lma_summary,
      geom = geom_text,
      mapping = aes(x = median_lma, y = family, label = round(median_lma, digits = 0)),
      colour = "black",
      offset = 0,
      pwidth = 0
    ) +
    ggnewscale::new_scale_fill() +
    geom_fruit(
      data = median_leaf_n_summary,
      geom = geom_tile,
      mapping = aes(x = median_leaf_n, y = family, fill = median_leaf_n),
      offset = 0.1,
      pwidth = 0,
      width = 14,
      height = 1
    ) +
    scale_fill_gradient(
      low = "white",
      high = "#5894cc"
    ) +
    geom_fruit(
      data = median_leaf_n_summary,
      geom = geom_text,
      mapping = aes(x = median_leaf_n, y = family, label = round(median_leaf_n, digits = 1)),
      colour = "black",
      offset = 0,
      pwidth = 0
    ) +
    ggnewscale::new_scale_fill() +
    geom_fruit(
      data = woodiness_summary,
      geom = geom_tile,
      mapping = aes(x = prop_woody, y = family, fill = prop_woody),
      offset = 0.1,
      pwidth = 0,
      width = 14,
      height = 1
    ) +
    scale_fill_gradient(
      low = "white",
      high = "#8f6856"
    ) +
    geom_fruit(
      data = woodiness_summary,
      geom = geom_text,
      mapping = aes(x = prop_woody, y = family, label = round(prop_woody, digits = 1)),
      colour = "black",
      offset = 0,
      pwidth = 0
    )
)
ggsave("figures/phylogeny_figure_colour.png", width = 10, height = 12)
# Modify in PowerPoint
```
