---
title: "fire_response_strategy_and_predictors"
author: "Sophie"
date: "22/11/2021"
output: html_document
editor_options:
  chunk_output_type: console
---


# Fire response strategy and fire frequency

Load libraries

```{r}
library(tidyverse)
library(patchwork)
```

Read in fire response strategy data

```{r}
resprouting_seeding <- readRDS("data/resprouting_seeding.rds")
```



## Mean fires from Poisson regression

Read in mean fire return interval data from GLMs

```{r}
mean_fires <- read_csv("outputs/mean_fires_df_with_poisson_glm.csv", guess_max = 10000)
mean_fires <- mean_fires %>% left_join(resprouting_seeding, by = c("taxon_name"))
```

Filter out taxa with warning messages

```{r}
mean_fires_no_errors <-
  mean_fires %>%
  filter(is.na(warnings), is.na(error), is.na(messages)) %>%
  select(-warnings, -messages, -error)
```

Clean data

```{r}
# Eliminate species with less than 10 pixels
mean_fires_clean <- mean_fires_no_errors |>
  dplyr::filter(num_pixels >= 10) |> # Removed 284
  # Convert to mean number of fires in 100 years, instead of mean number of fires in the sampling period
  mutate(mean_fires_100 = mean_fires / 22.17 * 100)

df_woody_herb_only <-
  mean_fires_clean %>%
  dplyr::filter(woody_or_herb %in% c("woody", "herb")) |>
  mutate(across(c("resprouting_binomial", "seeding_binomial", "woody_or_herb"), as.factor))

#write_csv(df_woody_herb_only, "outputs/final_df_predicted_fires_and_fire_response.csv")
#df_woody_herb_only <- read_csv("outputs/final_df_predicted_fires_and_fire_response.csv")
```

Histogram of mean fires

```{r}
hist(mean_fires_clean$mean_fires_100, breaks = 40)
```



### Binomial logistic regression


#### Resprouting binomial

```{r}
glm_resprouting <- glm(
  resprouting_binomial ~
    log10(mean_fires_100) + I(log10(mean_fires_100)^2) + log10(mean_fires_100):woody_or_herb +
    I(log10(mean_fires_100)^2):woody_or_herb + woody_or_herb,
  data = df_woody_herb_only,
  family = binomial
)

summary(glm_resprouting)
sjPlot::tab_model(glm_resprouting)

ggeffects::ggpredict(glm_resprouting, c("mean_fires_100", "woody_or_herb")) %>%
  ggplot(aes(x, predicted, color = group)) +
  geom_line(linewidth = 1) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group), alpha = 0.2, linetype = 0) +
  scale_x_log10() +
  ylab("Resprouting probability") +
  xlab("Mean fires") +
  theme_classic()

modelbased::estimate_expectation(glm_resprouting) |>
  ggplot(aes(x = mean_fires_100, y = Predicted, color = woody_or_herb)) +
  geom_line(linewidth = 1) +
  geom_ribbon(aes(ymin = CI_low, ymax = CI_high, fill = woody_or_herb), alpha = 0.2, linetype = 0) +
  scale_x_log10() +
  ylab("Resprouting probability") +
  xlab("Mean fires") +
  theme_classic()
```


Check accuracy of model

```{r}
index <-
  sample(
    seq_along(df_woody_herb_only$resprouting_binomial),
    0.7 * length(df_woody_herb_only$resprouting_binomial)
  )
training_x <- df_woody_herb_only[index, c("mean_fires_100", "woody_or_herb")]
training_y <- df_woody_herb_only[index, "resprouting_binomial"]
test_x <- df_woody_herb_only[-index, c("mean_fires_100", "woody_or_herb")]
test_y <- df_woody_herb_only[-index, "resprouting_binomial"]
```


```{r}
set.seed(1)

glm_resprouting_preds <- glm(
  resprouting_binomial ~
    log10(mean_fires_100) + I(log10(mean_fires_100)^2) + log10(mean_fires_100):woody_or_herb +
    I(log10(mean_fires_100)^2):woody_or_herb + woody_or_herb,
  data = data.frame(training_x, training_y),
  family = binomial
)
summary(glm_resprouting_preds)

predictions <- predict(
  glm_resprouting_preds,
  newdata = data.frame(test_x, test_y),
  newx = as.matrix(test_x), type = "link"
) # Make predictions

predictions_vs_real <- data.frame(test_x, test_y, predictions)

predictions_vs_real <-
  predictions_vs_real %>%
  mutate(
    preds_transformed = 1 / (1 + exp(-predictions)),
    preds_binomial = case_when(
      preds_transformed >= 0.5 ~ TRUE,
      preds_transformed < 0.5 ~ FALSE
    ),
    preds_binomial = as.factor(preds_binomial)
  )

predictions_vs_real %>%
  drop_na(resprouting_binomial) %>%
  ggplot(aes(resprouting_binomial, preds_transformed)) +
  geom_boxplot() +
  facet_wrap(~ woody_or_herb) +
  theme_classic()

caret::confusionMatrix(predictions_vs_real$preds_binomial, predictions_vs_real$resprouting_binomial)
```


```{r}
coef2 <- coef(glm_resprouting)
x1 <- log10(df_woody_herb_only$mean_fires_100)
x2 <- df_woody_herb_only$woody_or_herb
x2 <- ifelse(x2 == "woody", 1, 0)
pred_y <- coef2[1] + coef2[2] * x1 + coef2[3] * x1^2 + coef2[4] * x2 + coef2[5] * x1 * x2 + coef2[6] * x1^2 * x2
pred_p <- exp(pred_y) / (1 + exp(pred_y))
df1 <- data.frame(x1, x2, df_woody_herb_only$woody_or_herb, pred_y, pred_p)
df1 <- df1 %>% rename(woody_or_herb = df_woody_herb_only.woody_or_herb)

resprouting_plot <-
  ggplot(data = df1, aes(x = 10^x1, y = pred_p, col = woody_or_herb)) +
  geom_line(linewidth = 1.3) +
  scale_x_log10(
    expand = c(0, 0),
    limits = c(0.01, 100),
    labels = c("0.01", "0.1", "1", "10", "100")
  ) +
  scale_y_continuous(
    limits = c(0, 0.8),
    expand = c(0, 0)) +
  ylab("Resprouting probability") +
  xlab("Fire frequency (per century)") +
  scale_colour_manual(
    values = c("#7e5137", "#b9ce6f"),
    name = "",
    breaks = c("woody", "herb"),
    labels = c("Woody", "Herb")) +
  ggtitle("a)") +
  theme_classic() +
  theme(
    panel.border = element_rect(colour = "black", fill = NA),
    legend.position = "none",
    axis.text = element_text(size = 10),
    plot.margin = unit(c(0.1, 1, 0.1, 0.1), "cm"),
    plot.title.position = "plot",
    plot.title = element_text(size = 22, margin = margin(0, 0, 20, 0)),
    axis.title.y = element_text(size = 16, margin = margin(0, 15, 0, 0)),
    axis.title.x = element_blank(),
    panel.grid.major = element_line(colour = "#ebebeb", linewidth = 0.5),
    axis.line = element_blank())

resprouting_plot
```


Density distributions

```{r}
medians <- df_woody_herb_only %>%
  drop_na(resprouting_binomial) %>%
  group_by(woody_or_herb, resprouting_binomial) %>%
  summarise(median = median(mean_fires_100)) # Medians are robust to log scale and
  # also more suited to non-normal distributions

data_labels <- tibble(
  label = c("Woody", "Herb"),
  woody_or_herb = c("woody", "herb"),
  x =  c(0.015, 0.015),
  y = c(1, 1)
)

resprouting_density <-
  df_woody_herb_only %>%
  drop_na(resprouting_binomial) %>%
  ggplot() +
  geom_density(
    aes(mean_fires_100, fill = resprouting_binomial),
    alpha = 0.6,
    size = 0.8,
    col = NA
  ) +
  #geom_vline(
  #  data = medians,
  #  aes(xintercept = median, col = resprouting_binomial),
  #  alpha = 1,
  #  size = 1,
  #  linetype = "longdash"
  #) +
  facet_wrap(~ factor(woody_or_herb, ordered = TRUE, levels = c("woody", "herb")), ncol = 1, scales = "free") +
  geom_text(
    data = data_labels,
    aes(x = x, y = y, label = label, colour = label),
    hjust = 0,
    size = 7,
    fontface = "bold"
  ) +
  scale_x_log10(
    expand = c(0, 0),
    limits = c(0.01, 100),
    labels = c("0.01", "0.1", "1", "10", "100")
  ) +
  scale_y_continuous(
    expand = c(0, 0),
    limits = c(0, 1.25),
    breaks = seq(0, 1.25, by = 0.625),
    labels = c("0", "0.625", "1.25")
  ) +
  scale_fill_manual(
    values = c("#eb8775", "#172e6d"),
    labels = c("No", "Yes")
  ) +
  scale_colour_manual(
    values = c("#b9ce6f", "#7e5137")
  ) +
  labs(
    x = "Fire frequency (per century)",
    y = "Density",
    fill = "Resprouting/seeding"
  ) +
  theme_classic() +
  theme(
    axis.line = element_blank(),
    strip.background = element_blank(),
    strip.text = element_blank(),
    plot.margin = unit(c(0.5, 1, 0.1, 0.1), "cm"),
    panel.border = element_rect(colour = "black", fill = NA),
    panel.spacing = unit(0.5, "lines"),
    axis.text = element_text(size = 10),
    axis.title.y = element_text(size = 16, margin = margin(0, 15, 0, 0)),
    axis.title.x = element_text(size = 16, margin = margin(15, 0, 0, 0)),
    legend.position = "none")
resprouting_density
```





#### Seeding binomial


```{r}
glm_seeding <- glm(
  seeding_binomial ~
    log10(mean_fires_100) + I(log10(mean_fires_100)^2) + log10(mean_fires_100):woody_or_herb +
    I(log10(mean_fires_100)^2):woody_or_herb + woody_or_herb,
  data = df_woody_herb_only,
  family = binomial
)

summary(glm_seeding)
sjPlot::tab_model(glm_seeding)

ggeffects::ggpredict(glm_seeding, c("mean_fires_100", "woody_or_herb")) %>%
  ggplot(aes(x, predicted, color = group)) +
  geom_line(linewidth = 1) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group), alpha = 0.2, linetype = 0) +
  scale_x_log10() +
  ylab("Seeding probability") +
  xlab("Mean fires") +
  theme_classic()

modelbased::estimate_expectation(glm_seeding) |>
  ggplot(aes(x = mean_fires_100, y = Predicted, color = woody_or_herb)) +
  geom_line(linewidth = 1) +
  geom_ribbon(aes(ymin = CI_low, ymax = CI_high, fill = woody_or_herb), alpha = 0.2, linetype = 0) +
  scale_x_log10() +
  ylab("Seeding probability") +
  xlab("Mean fires") +
  theme_classic()

# Herb data does not look statistically significant
```


Model without herb data

```{r}
df_woody_only <-
  df_woody_herb_only %>%
  filter(woody_or_herb == "woody")

glm_seeding_woody <- glm(
  seeding_binomial ~ log10(mean_fires_100) + I(log10(mean_fires_100)^2),
  data = df_woody_only,
  family = binomial
)

summary(glm_seeding_woody)
sjPlot::tab_model(glm_seeding_woody)

ggeffects::ggpredict(glm_seeding_woody, c("mean_fires_100")) %>%
  ggplot(aes(x, predicted)) +
  geom_line(linewidth = 1) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2, linetype = 0) +
  scale_x_log10() +
  ylab("Seeding probability") +
  xlab("Mean fires") +
  theme_classic()
```


Check accuracy of model

```{r}
index <-
  sample(seq_along(df_woody_herb_only$seeding_binomial),
       0.7 * length(df_woody_herb_only$seeding_binomial))
training_x <- df_woody_herb_only[index, c("mean_fires_100", "woody_or_herb")]
training_y <- df_woody_herb_only[index, "seeding_binomial"]
test_x <- df_woody_herb_only[-index, c("mean_fires_100", "woody_or_herb")]
test_y <- df_woody_herb_only[-index, "seeding_binomial"]
```


```{r}
set.seed(1)

glm_seeding_preds <- glm(
  seeding_binomial ~ log10(mean_fires_100) + I(log10(mean_fires_100)^2),
  data = data.frame(training_x, training_y),
  family = binomial
)
summary(glm_seeding_preds)

predictions <- predict(
  glm_seeding_preds,
  newdata = data.frame(test_x, test_y),
  newx = as.matrix(test_x), type = "link"
) # Make predictions

predictions_vs_real <- data.frame(test_x, test_y, predictions)

predictions_vs_real <-
  predictions_vs_real %>%
  mutate(
    preds_transformed = 1 / (1 + exp(-predictions)),
    preds_binomial = case_when(
      preds_transformed >= 0.5 ~ TRUE,
      preds_transformed < 0.5 ~ FALSE
    ),
    preds_binomial = as.factor(preds_binomial)
  )

predictions_vs_real %>%
  drop_na(seeding_binomial) %>%
  ggplot(aes(seeding_binomial, preds_transformed)) +
  geom_boxplot() +
  facet_wrap(~ woody_or_herb) +
  theme_classic()

caret::confusionMatrix(predictions_vs_real$preds_binomial, predictions_vs_real$seeding_binomial)
```


Plot fitted model

```{r}
coef2 <- coef(glm_seeding_woody)
x1 <- log10(df_woody_only$mean_fires_100)
x2 <- df_woody_only$woody_or_herb
x2 <- ifelse(x2 == "woody", 1, 0)
pred_y <- coef2[1] + coef2[2] * x1 + coef2[3] * x1^2
pred_p <- exp(pred_y) / (1 + exp(pred_y))
df1 <- data.frame(x1, x2, df_woody_only$woody_or_herb, pred_y, pred_p)
df1 <- df1 %>% rename(woody_or_herb = df_woody_only.woody_or_herb)

seeding_plot <-
  ggplot(data = df1, aes(x = 10^x1, y = pred_p, col = woody_or_herb)) +
  geom_line(linewidth = 1.3) +
  scale_x_log10(
    expand = c(0, 0),
    limits = c(0.01, 100),
    labels = c("0.01", "0.1", "1", "10", "100")
  ) +
  scale_y_continuous(
    limits = c(0, 1.0),
    expand = c(0, 0)) +
  ylab("Seeding probability") +
  xlab("Fire frequency (per century)") +
  scale_colour_manual(
    values = c("#7e5137"),
    name = "",
    breaks = c("woody"),
    labels = c("Woody")) +
  ggtitle("b)") +
  theme_classic() +
  theme(
    panel.border = element_rect(colour = "black", fill = NA),
    legend.position = "none",
    axis.text = element_text(size = 10),
    plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"),
    plot.title.position = "plot",
    plot.title = element_text(size = 22, margin = margin(0, 0, 20, 0)),
    axis.title.y = element_text(size = 16, margin = margin(0, 15, 0, 0)),
    axis.title.x = element_blank(),
    panel.grid.major = element_line(colour = "#ebebeb", linewidth = 0.5),
    axis.line = element_blank()
  )

seeding_plot
```


Density distributions

```{r}
medians <- df_woody_only %>%
  drop_na(seeding_binomial) %>%
  group_by(seeding_binomial) %>%
  summarise(median = median(mean_fires_100))

data_labels <- tibble(
  label = c("Woody"),
  woody_or_herb = c("woody"),
  x =  c(0.015),
  y = c(1)
)

seeding_density <-
  df_woody_only %>%
  drop_na(seeding_binomial) %>%
  ggplot(aes(mean_fires_100, fill = seeding_binomial)) +
  geom_density(alpha = 0.6, size = 0.8, col = NA) +
  #geom_vline(
  #  data = medians,
  #  aes(xintercept = median, col = seeding_binomial),
  #  alpha = 1,
  #  size = 1,
  #  linetype = "longdash") +
  scale_x_log10(
    expand = c(0, 0),
    limits = c(0.01, 100),
    labels = c("0.01", "0.1", "1", "10", "100")
  ) +
  scale_y_continuous(
    expand = c(0, 0),
    limits = c(0, 1.25),
    breaks = seq(0, 1.25, by = 0.625)
  ) +
  scale_fill_manual(
    values = c("#eb8775", "#172e6d"),
    labels = c("No", "Yes")) +
  scale_colour_manual(values = c("#eb8775", "#172e6d")) +
  guides(fill = guide_legend(ncol = 2)) +
  labs(
    x = "Fire frequency (per century)",
    y = "Density",
    fill = "Resprouting/seeding"
  ) +
  theme_classic() +
  theme(
    axis.line = element_blank(),
    plot.margin = unit(c(0.5, 0.1, 0.1, 0.1), "cm"),
    panel.border = element_rect(colour = "black", fill = NA),
    panel.spacing = unit(0.5, "lines"),
    axis.text = element_text(size = 10),
    axis.title.y = element_text(size = 16, margin = margin(0, 15, 0, 0)),
    axis.title.x = element_text(size = 16, margin = margin(15, 0, 0, 0)),
    legend.position = "bottom",
    legend.title = element_text(size = 16),
    legend.text = element_text(size = 14)
  )

# Get legend
density_legend <- cowplot::get_legend(seeding_density)

seeding_density <-
  df_woody_only %>%
  drop_na(seeding_binomial) %>%
  ggplot() +
  geom_density(
    aes(mean_fires_100, fill = seeding_binomial),
    alpha = 0.6, size = 0.8, col = NA
  ) +
  geom_text(
    data = data_labels,
    aes(x = x, y = y, label = label, colour = label),
    hjust = 0,
    size = 7,
    fontface = "bold"
  ) +
  #geom_vline(
  #  data = medians,
  #  aes(xintercept = median, col = seeding_binomial),
  #  alpha = 1,
  #  size = 1,
  #  linetype = "longdash") +
  scale_x_log10(
    expand = c(0, 0),
    limits = c(0.01, 100),
    labels = c("0.01", "0.1", "1", "10", "100")
  ) +
  scale_y_continuous(
    expand = c(0, 0),
    limits = c(0, 1.25),
    breaks = seq(0, 1.25, by = 0.625),
    labels = c("0", "0.625", "1.25")
  ) +
  scale_fill_manual(
    values = c("#eb8775", "#172e6d"),
    labels = c("No", "Yes")) +
  scale_colour_manual(
    values = c("#7e5137")
  ) +
  guides(fill = guide_legend(ncol = 2)) +
  labs(
    x = "Fire frequency (per century)",
    y = "Density",
    fill = "Resprouting/seeding"
  ) +
  theme_classic() +
  theme(
    axis.line = element_blank(),
    plot.margin = unit(c(0.5, 0.1, 0.1, 0.1), "cm"),
    panel.border = element_rect(colour = "black", fill = NA),
    panel.spacing = unit(0.5, "lines"),
    axis.text = element_text(size = 10),
    axis.title.y = element_text(size = 16, margin = margin(0, 15, 0, 0)),
    axis.title.x = element_text(size = 16, margin = margin(15, 0, 0, 0)),
    legend.position = "none",
    legend.title = element_text(size = 16),
    legend.text = element_text(size = 14)
  )

seeding_density
```


Plot resprouting and seeding probability graphs together

```{r}
design <- "12\n34\n35"

resprouting_plot + seeding_plot + resprouting_density + seeding_density + density_legend +
  plot_layout(design = design, heights = c(8, 3.75, 2.5, 3.5))

ggsave("figures/fire_response_probabilities_density_functions.png", width = 10, height = 8)
```




### Phylogenetic logistic regression

#### Resprouting binomial


Phylogenetic regression

```{r}
library(phylolm)
library(future)
library(ape)
plan(multisession)

tree <- read.tree("data/v0.1-big-seed-plant-trees/ALLMB.tre")
df_woody_herb_only$taxon_name <- gsub(" ", "_", df_woody_herb_only$taxon_name)
df_woody_herb_only <- df_woody_herb_only[df_woody_herb_only$taxon_name %in% tree$tip.label, ]
df_woody_herb_only$resprouting_binomial <- ifelse(df_woody_herb_only$resprouting_binomial == TRUE, 1, 0) # Make numeric

pruned_tree <- drop.tip(tree, setdiff(tree$tip.label, df_woody_herb_only$taxon_name))
pruned_tree <- as.phylo(pruned_tree)
```


With sorted data

```{r}
sorted_data <- df_woody_herb_only[order(match(df_woody_herb_only$taxon_name, pruned_tree$tip.label)), ]

sorted_data <-
  sorted_data %>%
  select(taxon_name, mean_fires_100, resprouting_binomial, woody_or_herb) %>%
  column_to_rownames("taxon_name") |>
  mutate(across(c("woody_or_herb", "resprouting_binomial"), as.factor))

pglm_resprouting <- phyloglm(
  resprouting_binomial ~
    log10(mean_fires_100) + I(log10(mean_fires_100)^2) + log10(mean_fires_100):woody_or_herb +
    I(log10(mean_fires_100)^2):woody_or_herb + woody_or_herb,
  method = "logistic_MPLE",
  data = sorted_data,
  phy = pruned_tree,
  btol = 100,
  boot = 0
)

summary(pglm_resprouting)
sjPlot::tab_model(pglm_resprouting)

modelbased::estimate_prediction(pglm_resprouting) |>
  ggplot(aes(x = mean_fires_100, y = Predicted, color = woody_or_herb)) +
  geom_line(linewidth = 1) +
  scale_x_log10() +
  ylab("Resprouting probability") +
  xlab("Mean fires") +
  theme_classic()

coef2 <- coef(pglm_resprouting)
x1 <- log10(sorted_data$mean_fires_100)
x2 <- sorted_data$woody_or_herb
x2 <- ifelse(x2 == "woody", 1, 0)
pred_y <- coef2[1] + coef2[2] * x1 + coef2[3] * x1^2 + coef2[4] * x2 + coef2[5] * x1 * x2 + coef2[6] * x1^2 * x2
pred_p <- exp(pred_y) / (1 + exp(pred_y))
df1 <- data.frame(x1, x2, sorted_data$woody_or_herb, pred_y, pred_p)
df1 <- df1 %>% rename(woody_or_herb = sorted_data.woody_or_herb)

resprouting_phylo_plot <-
  ggplot(data = df1, aes(x = 10^x1, y = pred_p, col = woody_or_herb)) +
  geom_line(linewidth = 1.3) +
  scale_x_log10(
    expand = c(0, 0),
    limits = c(0.01, 100),
    labels = c("0.01", "0.1", "1", "10", "100")
  ) +
  scale_y_continuous(
    limits = c(0, 1),
    expand = c(0, 0)) +
  ylab("Resprouting probability") +
  xlab("Fire frequency (per century)") +
  scale_colour_manual(
    values = c("#7e5137", "#b9ce6f"),
    name = "",
    breaks = c("woody", "herb"),
    labels = c("Woody", "Herb")) +
  ggtitle("a)") +
  theme_classic() +
  theme(
    panel.border = element_rect(colour = "black", fill = NA),
    legend.position = "none",
    axis.text = element_text(size = 10),
    plot.margin = unit(c(0.1, 1, 0.1, 0.1), "cm"),
    plot.title.position = "plot",
    plot.title = element_text(size = 22, margin = margin(0, 0, 20, 0)),
    axis.title.y = element_text(size = 16, margin = margin(0, 15, 0, 0)),
    axis.title.x = element_text(size = 16, margin = margin(15, 0, 0, 0)),
    panel.grid.major = element_line(colour = "#ebebeb", linewidth = 0.5),
    axis.line = element_blank())
resprouting_phylo_plot

resprouting_phylo_plot_2 <-
  ggplot(data = df1, aes(x = 10^x1, y = pred_p, col = woody_or_herb)) +
  geom_line(linewidth = 1.3) +
  scale_x_log10(
    expand = c(0, 0),
    limits = c(0.01, 100),
    labels = c("0.01", "0.1", "1", "10", "100")
  ) +
  scale_y_continuous(
    limits = c(0, 1),
    expand = c(0, 0)) +
  ylab("Resprouting probability") +
  xlab("Fire frequency (per century)") +
  scale_colour_manual(
    values = c("#7e5137", "#b9ce6f"),
    name = "",
    breaks = c("woody", "herb"),
    labels = c("Woody", "Herb")) +
  ggtitle("a)") +
  theme_classic() +
  theme(
    panel.border = element_rect(colour = "black", fill = NA),
    legend.position = "none",
    axis.text = element_text(size = 10),
    plot.margin = unit(c(0.1, 1, 0.1, 0.1), "cm"),
    plot.title.position = "plot",
    plot.title = element_text(size = 22, margin = margin(0, 0, 20, 0)),
    axis.title.y = element_text(size = 16, margin = margin(0, 15, 0, 0)),
    axis.title.x = element_blank(),
    panel.grid.major = element_line(colour = "#ebebeb", linewidth = 0.5),
    axis.line = element_blank())
resprouting_phylo_plot_2
```


To calculate R2 Tjur

```{r}
pred <- fitted(pglm_resprouting)
y <- pglm_resprouting$y

categories <- unique(y)
mean1 <- mean(pred[which(y == categories[1])], na.rm = TRUE)
mean2 <- mean(pred[which(y == categories[2])], na.rm = TRUE)

abs(mean1 - mean2) # 0.0312
```


Fit logistic curve

```{r}
coef2 <- coef(pglm_resprouting)
plot.new()
plot(
  seq(-100, 100, by = 20),
  seq(0, 1, by = 0.1),
  xlab = "Mean fires",
  ylab = "Resprouting probability"
)
curve(plogis(coef2[1] + coef2[2] * x), col = "red", add = TRUE)

sorted_data |>
  dplyr::filter(!is.na(resprouting_binomial)) |>
  mutate(resprouting_binomial = if_else(resprouting_binomial == 1, 1, 0)) |>
  ggplot(aes(x = mean_fires_100, y = resprouting_binomial)) +
  geom_jitter(height = 0.01, alpha = 0.25) +
  stat_smooth(method = "glm", method.args = list(family = binomial)) +
  facet_wrap(~ woody_or_herb) +
  scale_x_log10() +
  theme_classic()
```


#### Seeding binomial


Model without herb data

```{r}
df_woody_only$taxon_name <- gsub(" ", "_", df_woody_only$taxon_name)
df_woody_only <- df_woody_only[df_woody_only$taxon_name %in% tree$tip.label, ]
df_woody_only$seeding_binomial <- ifelse(df_woody_only$seeding_binomial == TRUE, 1, 0) # make numeric

pruned_tree <- drop.tip(tree, setdiff(tree$tip.label, df_woody_only$taxon_name))
pruned_tree <- as.phylo(pruned_tree)

sorted_data <- df_woody_only[order(match(df_woody_only$taxon_name, pruned_tree$tip.label)), ]

sorted_data <-
  sorted_data %>%
  select(taxon_name, mean_fires_100, seeding_binomial, woody_or_herb) %>%
  column_to_rownames("taxon_name") |>
  mutate(across(c("woody_or_herb", "seeding_binomial"), as.factor))

pglm_seeding <- phyloglm(
  seeding_binomial ~ log10(mean_fires_100) + I(log10(mean_fires_100)^2),
  method = "logistic_MPLE",
  data = sorted_data,
  phy = pruned_tree,
  btol = 100,
  boot = 0)

summary(pglm_seeding)
sjPlot::tab_model(pglm_seeding)
```


To calculate R2 Tjur

```{r}
pred <- fitted(pglm_seeding)
y <- pglm_seeding$y

categories <- unique(y)
mean1 <- mean(pred[which(y == categories[1])], na.rm = TRUE)
mean2 <- mean(pred[which(y == categories[2])], na.rm = TRUE)

abs(mean1 - mean2) # 0.039
```


Plot

```{r}
coef2 <- coef(pglm_seeding)
x1 <- log10(sorted_data$mean_fires_100)
x2 <- sorted_data$woody_or_herb
x2 <- ifelse(x2 == "woody", 1, 0)
pred_y <- coef2[1] + coef2[2] * x1 + coef2[3] * x1^2
pred_p <- exp(pred_y) / (1 + exp(pred_y))
df1 <- data.frame(x1, x2, sorted_data$woody_or_herb, pred_y, pred_p)
df1 <- df1 %>% rename(woody_or_herb = sorted_data.woody_or_herb)

seeding_phylo_plot <-
  ggplot(data = df1, aes(x = 10^x1, y = pred_p, col = woody_or_herb)) +
  geom_line(linewidth = 1.3) +
  scale_x_log10(
    expand = c(0, 0),
    limits = c(0.01, 100),
    labels = c("0.01", "0.1", "1", "10", "100")
  ) +
  scale_y_continuous(
    limits = c(0, 1),
    expand = c(0, 0)) +
  scale_colour_manual(
    values = c("#7e5137"),
    name = "",
    breaks = c("woody"),
    labels = c("Woody")) +
  ggtitle("b)") +
  ylab("Seeding probability") +
  xlab("Fire frequency (per century)") +
  theme_classic() +
  theme(
    panel.border = element_rect(colour = "black", fill = NA),
    legend.position = "none",
    axis.text = element_text(size = 10),
    plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"),
    plot.title.position = "plot",
    plot.title = element_text(size = 22, margin = margin(0, 0, 20, 0)),
    axis.title.y = element_text(size = 16, margin = margin(0, 15, 0, 0)),
    axis.title.x = element_text(size = 16, margin = margin(15, 0, 0, 0)),
    panel.grid.major = element_line(colour = "#ebebeb", linewidth = 0.5),
    axis.line = element_blank()
  )

seeding_phylo_plot


seeding_phylo_plot_2 <-
  ggplot(data = df1, aes(x = 10^x1, y = pred_p, col = woody_or_herb)) +
  geom_line(linewidth = 1.3) +
  scale_x_log10(
    expand = c(0, 0),
    limits = c(0.01, 100),
    labels = c("0.01", "0.1", "1", "10", "100")
  ) +
  scale_y_continuous(
    limits = c(0, 1),
    expand = c(0, 0)) +
  scale_colour_manual(
    values = c("#7e5137"),
    name = "",
    breaks = c("woody"),
    labels = c("Woody")) +
  ggtitle("b)") +
  ylab("Seeding probability") +
  xlab("Fire frequency (per century)") +
  theme_classic() +
  theme(
    panel.border = element_rect(colour = "black", fill = NA),
    legend.position = "none",
    axis.text = element_text(size = 10),
    plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"),
    plot.title.position = "plot",
    plot.title = element_text(size = 22, margin = margin(0, 0, 20, 0)),
    axis.title.y = element_text(size = 16, margin = margin(0, 15, 0, 0)),
    axis.title.x = element_blank(),
    panel.grid.major = element_line(colour = "#ebebeb", linewidth = 0.5),
    axis.line = element_blank()
  )

seeding_phylo_plot_2
```


```{r}
resprouting_phylo_plot + seeding_phylo_plot

ggsave("figures/fire_response_probabilities_phyloglm.png", height = 4.5, width = 9)
```


Main figure but with phylo plots

```{r}
design <- "12\n34\n35"

resprouting_phylo_plot_2 + seeding_phylo_plot_2 + resprouting_density + seeding_density + density_legend +
  plot_layout(design = design, heights = c(8, 3.75, 2.5, 3.5))

ggsave("figures/fire_response_probabilities_density_functions_phylo.png", width = 10, height = 8)
```


---

## Comparison of survival analysis and poisson regression methods


### Survival analysis method

Read in median fire return interval data from survival analysis

```{r}
median_fris <- read_csv("outputs/median_fris_with_survival_analysis.csv")
median_fris <- median_fris %>% left_join(resprouting_seeding, by = c("taxon_name"))

median_fris |> dplyr::filter(num_pixels >= 10) |> nrow()
median_fris |> dplyr::filter(!is.na(med_w_warnings), num_pixels >= 10) |> nrow()

median_fires_weibull <- median_fris %>%
  dplyr::filter(is.na(med_w_warnings), num_pixels >= 10) %>%
  mutate(across(
    c("med_w", "med_w_lwr", "med_w_upr", "med_wo", "med_wo_lwr", "med_wo_upr"),
    ~ .x / 365.2422 # Convert intervals to years not days
  )) |>
  mutate(across(
    c("med_w", "med_w_lwr", "med_w_upr", "med_wo", "med_wo_lwr", "med_wo_upr"),
    ~ 100 / .x # Convert to fire frequency per century
  )) |>
  mutate(confint_size =  med_w_lwr - med_w_upr) |> # Lower minus upper because original FRI was upper minus lower
  select(all_of(c(
    "taxon_name", "med_w", "med_w_lwr", "med_w_upr", "confint_size", "num_unburnt_pixels", "num_pixels",
    "resprouting_binomial", "seeding_binomial", "data_on_both", "woody_or_herb", "taxon_rank"
  ))) |>
  dplyr::filter(confint_size < 30) # Filter out confidence intervals greater than 30 years
```

With the survival analysis method, 1099 models did not converge (incl. unburnt
pixels) out of 10406 species with GBIF occurrences >= 10. The warning was "ran out of iterations and did not converge". This occurred
because most or all pixels were unburnt.


Histogram of median fire frequencies and confidence intervals

```{r}
hist(median_fires_weibull$med_w, breaks = 30)
hist(median_fires_weibull$confint_size, breaks = 30)
```


### Poisson regression method

```{r}
mean_fires <- read_csv("outputs/mean_fires_df_with_poisson_glm.csv", guess_max = 10871)
mean_fires <- mean_fires %>% left_join(resprouting_seeding, by = c("taxon_name"))

mean_fires |> dplyr::filter(num_pixels >= 10) |> nrow()
mean_fires |>
  dplyr::filter(num_pixels >= 10) |>
  dplyr::filter(!is.na(warnings) | !is.na(messages) | !is.na(error)) |>
  nrow()

mean_fires_poisson <- mean_fires |>
  # Convert to mean number of fires in 100 years, instead of mean number of fires in the sampling period
  mutate(across(
    c("mean_fires", "upr_conf_int", "lwr_conf_int"),
    ~ .x / 22.17 * 100
  )) |>
  dplyr::filter(is.na(warnings), is.na(error), is.na(messages), num_pixels >= 10) |>
  mutate(confint_size = upr_conf_int - lwr_conf_int) |>
  select(-all_of(c(
    "mean_fri", "warnings", "messages", "error", "resprouting", "post_fire_seeding",
    "resprouter_or_seeder", "woodiness_detailed", "obligate_resprouting_or_seeding",
    "resprouting_seeding_binomial"
  )))

hist(mean_fires_poisson$mean_fires, breaks = 30)
hist(mean_fires_poisson$confint_size, breaks = 30)
```


In the Poisson regression method, out of 10406 species with GBIF occurrences >= 10, 166 did not fit and they were all because all pixels were unburnt.


Join data from the two methods

```{r}
method_comparison <- mean_fires_poisson |>
  select(-c("confint_size", "upr_conf_int", "lwr_conf_int", "dispersion_ratio", "dispersion_p_value")) |>
  mutate(method = "poisson") |>
  rename("avg_fires" = "mean_fires") |>
  full_join(
    median_fires_weibull |>
      select(-c("confint_size", "med_w_lwr", "med_w_upr")) |>
      mutate(method = "survival") |>
      rename("avg_fires" = "med_w")
  )
```


Plot methods against each other

```{r}
method_comparison |>
  ggplot() +
  geom_density(
    aes(avg_fires, fill = method),
    alpha = 0.6,
    size = 0.8,
    col = NA
  ) +
  scale_x_log10(
    expand = c(0, 0),
    limits = c(0.001, 100),
    breaks = c(0.001, 0.01, 0.1, 1, 10, 100),
    labels = c("0.001", "0.01", "0.1", "1", "10", "100")
  ) +
  scale_y_continuous(
    expand = c(0, 0),
    limits = c(0, 1)
  ) +
  scale_fill_manual(
    values = c("#d1b2eb", "#5e804a"),
    labels = c("Poisson regression", "Survival analysis")
  ) +
  labs(
    x = "Fire frequency (per century)",
    y = "Density",
    fill = "Method"
  ) +
  theme_classic() +
  theme(
    axis.line = element_blank(),
    strip.background = element_blank(),
    strip.text = element_blank(),
    plot.margin = unit(c(0.5, 1, 0.1, 0.1), "cm"),
    panel.border = element_rect(colour = "black", fill = NA),
    panel.spacing = unit(0.5, "lines"),
    axis.text = element_text(size = 10),
    axis.title.y = element_text(size = 16, margin = margin(0, 15, 0, 0)),
    axis.title.x = element_text(size = 16, margin = margin(15, 0, 0, 0)),
    legend.title = element_text(size = 15),
    legend.text = element_text(size = 13),
    legend.position = "bottom"
  )

ggsave("figures/method_comparison.png", width = 7, height = 4)
```


```{r}
method_comparison |>
  filter(woody_or_herb %in% c("woody", "herb")) |>
  mutate(woody_or_herb = if_else(woody_or_herb == "woody", "Woody", "Herbs")) |>
  ggplot() +
  geom_density(
    aes(avg_fires, fill = method),
    alpha = 0.6,
    size = 0.8,
    col = NA
  ) +
  facet_wrap(
    ~ factor(woody_or_herb, ordered = TRUE, levels = c("Woody", "Herbs")),
    ncol = 1
  ) +
  scale_x_log10(
    expand = c(0, 0),
    limits = c(0.001, 100),
    breaks = c(0.001, 0.01, 0.1, 1, 10, 100),
    labels = c("0.001", "0.01", "0.1", "1", "10", "100")
  ) +
  scale_y_continuous(
    expand = c(0, 0),
    limits = c(0, 1)
  ) +
  scale_fill_manual(
    values = c("#d1b2eb", "#5e804a"),
    labels = c("Poisson regression", "Survival analysis")
  ) +
  labs(
    x = "Fire frequency (per century)",
    y = "Density",
    fill = "Method"
  ) +
  theme_classic() +
  theme(
    strip.background = element_blank(),
    strip.text = element_text(size = 17, hjust = 0, margin = margin(10, 0, 10, 5), face = "bold"),
    axis.line = element_blank(),
    plot.margin = unit(c(0.5, 1, 0.1, 0.1), "cm"),
    panel.border = element_rect(colour = "black", fill = NA),
    panel.spacing = unit(0.5, "lines"),
    axis.text = element_text(size = 12),
    axis.title.y = element_text(size = 18, margin = margin(0, 15, 0, 0)),
    axis.title.x = element_text(size = 18, margin = margin(15, 0, 0, 0)),
    legend.title = element_text(size = 17),
    legend.text = element_text(size = 15),
    legend.position = "bottom"
  )

ggsave("figures/method_comparison_woody_herb.png", width = 7, height = 7)
```

