---
title: "Fire response data"
output: html_notebook
editor_options:
  chunk_output_type: console
---


```{r}
library(tidyverse)

austraits <- read_rds("data/austraits-5.0.0.rds")
```


# Extract `resprouting_capacity` trait records

```{r}
resprouting_capacity <-
  austraits$traits %>%
  filter(trait_name == "resprouting_capacity") %>%
  pivot_wider(values_from = value, names_from = trait_name)
```


Filter out non-native taxa

```{r}
aus_native_lookup <- read_csv("data/aus_native_lookup.csv")
# Only includes accepted species and species or below taxon ranks

resprouting_capacity <-
  resprouting_capacity %>%
  left_join(aus_native_lookup,
            by = c("taxon_name" = "canonical_name")) %>%
  filter(aus_native == TRUE) %>%
  select(-aus_native) # Lost 1757 taxa
```


Join taxonomic information

```{r}
resprouting_capacity <-
  resprouting_capacity %>%
  left_join(
    austraits$taxa %>%
      select(taxon_name, genus, family, taxon_rank, taxonomic_status, taxon_distribution),
    by = "taxon_name")
```


Create dataframe with tallies of the number of observations for each value

```{r}
resprouting_capacity_summary <-
  resprouting_capacity %>%
  group_by(taxon_name, resprouting_capacity, taxon_rank) %>%
  summarise(count = n()) %>%
  pivot_wider(names_from = resprouting_capacity,
              values_from = count) %>%
  ungroup() %>%
  mutate(across(where(is.integer), ~ replace_na(., 0)))

names(resprouting_capacity_summary) <-
  names(resprouting_capacity_summary) %>%
  str_replace_all(c(" " = "_"))
```


Calculate a binary column for whether there is variation in observations

```{r}
col <- colnames(resprouting_capacity_summary)
col <- col[!col %in% c("taxon_name", "taxon_rank")]

resprouting_capacity_summary <-
  resprouting_capacity_summary %>%
  mutate(variation_in_strat = (rowSums(resprouting_capacity_summary[, col] > 0) > 1))
```


## For taxa with no variation in observation of resprouting capacity

```{r}
resprouting_capacity_summary <-
  resprouting_capacity_summary %>%
  mutate(
    resprouting = if_else(
      variation_in_strat == FALSE,
      case_when(
        fire_killed >= 1 ~ "does not resprout",
        resprouts >= 1 ~ "mostly resprouts",
        fire_killed_partial_resprouting >= 1 ~ "sometimes resprouts",
        fire_killed_resprouts >= 1 ~ "sometimes resprouts",
        partial_resprouting >= 1 ~ "sometimes resprouts",
        fire_killed_partial_resprouting_resprouts >= 1 ~ "sometimes resprouts",
        partial_resprouting_resprouts >= 1 ~ "mostly resprouts"),
      NA_character_
      )
    )
```


## For taxa with variation

Make columns that tally the number of observations for fire-killed, resprouts, mixed
fire responses, fire resistors and no fire exposure

```{r}
resprouting_capacity_summary <-
  resprouting_capacity_summary %>%
  mutate(
    resprout_tally =
      resprouts + fire_killed_partial_resprouting + fire_killed_resprouts + partial_resprouting +
      fire_killed_partial_resprouting_resprouts + partial_resprouting_resprouts,
    does_not_resprout_tally =
      fire_killed + fire_killed_partial_resprouting + fire_killed_resprouts +
      fire_killed_partial_resprouting_resprouts
  )
```


Plot histogram of percentage of resprout observations

```{r}
resprouting_capacity_summary <-
  resprouting_capacity_summary %>%
  mutate(resprout_prop = resprout_tally / (resprout_tally + does_not_resprout_tally))

resprouting_capacity_summary %>%
  filter(variation_in_strat == TRUE) %>%
  ggplot(aes(x = resprout_prop)) +
  geom_histogram() +
  scale_x_continuous(breaks = seq(0.0, 1.0, by = 0.1)) +
  theme_classic()
```


If over 70% of observations are resprouts, then can designate as mostly resprouts.
If less than 30% are resprouts, designate as does not resprout.
If less than 70% but more than 30% observations are resprouts, then designate as sometimes resprouts.

```{r}
resprouting_capacity_summary <-
  resprouting_capacity_summary %>%
  mutate(
    resprouting = if_else(
      variation_in_strat == TRUE,
      case_when(
        resprout_prop < 0.3 ~ "does not resprout",
        resprout_prop < 0.7 ~ "sometimes resprouts",
        resprout_prop >= 0.7 ~ "mostly resprouts"),
      resprouting
      )
    )
```


Summary of final resprouting values and counts

```{r}
resprouting_capacity_summary %>%
  group_by(resprouting) %>%
  summarise(count = n()) %>%
  ungroup()

# Final tibble for resprouting variable
resprouting <-
  resprouting_capacity_summary %>%
  select(taxon_name, taxon_rank, resprouting)
```



---

# Extract post_fire_recruitment trait records


Subset traits to post_fire_recruitment records only

```{r}
post_fire_recruitment <-
  austraits$traits %>%
  filter(trait_name == "post_fire_recruitment") %>%
  pivot_wider(values_from = value, names_from = trait_name)
```


Filter out non-native species

```{r}
post_fire_recruitment <-
  post_fire_recruitment %>%
  left_join(aus_native_lookup,
            by = c("taxon_name" = "canonical_name")) %>%
  filter(aus_native == TRUE) %>%
  select(-aus_native) # Lost 336
```


Join taxonomic information

```{r}
post_fire_recruitment <-
  post_fire_recruitment %>%
  left_join(
    austraits$taxa %>%
      select(taxon_name, genus, family, taxon_rank, taxonomic_status, taxon_distribution),
    by = "taxon_name")
```


Create dataframe with tallies of the number of observations for each value

```{r}
post_fire_recruitment_summary <-
  post_fire_recruitment %>%
  group_by(taxon_name, post_fire_recruitment, taxon_rank) %>%
  summarise(count = n()) %>%
  pivot_wider(names_from = post_fire_recruitment, values_from = count) %>%
  ungroup() %>%
  mutate(across(where(is.integer), ~ replace_na(., 0)))

names(post_fire_recruitment_summary) <-
  names(post_fire_recruitment_summary) %>%
  str_replace_all(c(" " = "_"))
```


Calculate a binary column for whether there is variation in observations

```{r}
col <- colnames(post_fire_recruitment_summary)
col <- col[!col %in% c("taxon_name", "taxon_rank")]

post_fire_recruitment_summary <-
  post_fire_recruitment_summary %>%
  mutate(variation_in_seeding = (rowSums(post_fire_recruitment_summary[, col] > 0) > 1))
```


## For taxa with no variation in observation of post-fire recruitment

```{r}
post_fire_recruitment_summary <-
  post_fire_recruitment_summary %>%
  mutate(
    post_fire_seeding = if_else(
      variation_in_seeding == FALSE,
      case_when(
        post_fire_recruitment >= 1 ~ "robust post-fire seeding",
        post_fire_recruitment_absent >= 1 ~ "no post-fire seeding",
        post_fire_recruitment_post_fire_recruitment_absent >= 1 ~ "some post-fire seeding"),
      NA_character_
      )
    )
```


## For taxa with no variation

Make columns that tally the number of observations for post-fire recruitment and no
post-fire recruitment

```{r}
post_fire_recruitment_summary <-
  post_fire_recruitment_summary %>%
  mutate(post_fire_recruitment_tally =
           post_fire_recruitment +
           post_fire_recruitment_post_fire_recruitment_absent,
         no_post_fire_recruitment_tally =
           post_fire_recruitment_absent +
           post_fire_recruitment_post_fire_recruitment_absent)
```


Plot histogram of percentage of seeding observations

```{r}
post_fire_recruitment_summary <-
  post_fire_recruitment_summary %>%
  mutate(seeding_prop = post_fire_recruitment_tally / (post_fire_recruitment_tally + no_post_fire_recruitment_tally))

post_fire_recruitment_summary %>%
  filter(variation_in_seeding == TRUE) %>%
  ggplot(aes(x = seeding_prop)) +
  geom_histogram() +
  scale_x_continuous(breaks = seq(0.0, 1.0, by = 0.1)) +
  theme_bw()
```


If over 70% of observations are post-fire recruitment, then can designate as robust
post-fire seeding. If less than 30% of observations are no post-fire recruitment, designate as no post-fire seeding. If less than 70% but more than 30% observations are post-fire recruitment, then designate as some post-fire seeding. (Decided 70% based on previous histogram.)

```{r}
post_fire_recruitment_summary <-
  post_fire_recruitment_summary %>%
  mutate(
    post_fire_seeding = if_else(
      variation_in_seeding == TRUE,
      case_when(
        seeding_prop < 0.3 ~ "no post-fire seeding",
        seeding_prop < 0.7 ~ "some post-fire seeding",
        seeding_prop >= 0.7 ~ "robust post-fire seeding"),
      post_fire_seeding)
    )
```


Summary of final seeding values and counts

```{r}
post_fire_recruitment_summary %>%
  group_by(post_fire_seeding) %>%
  summarise(count = n()) %>%
  ungroup()

# Final tibble for seeding variable
post_fire_seeding <- post_fire_recruitment_summary %>% select(taxon_name, taxon_rank, post_fire_seeding)
```



---

# Join resprouting_capacity and post_fire_recruitment

```{r}
resprouting_taxa <- resprouting$taxon_name %>% unique() # 11060 taxa
seeding_taxa <- post_fire_seeding$taxon_name %>% unique() # 3615 taxa

intersecting_taxa <- intersect(resprouting_taxa, seeding_taxa)
# 3474 taxa intersect between resprouting_capacity and post_fire_recruitment
```


```{r}
resprouting_seeding <-
  resprouting %>%
  full_join(post_fire_seeding, by = c("taxon_name", "taxon_rank"))

resprouting_seeding <-
  resprouting_seeding %>%
  mutate(data_on_both = !is.na(resprouting) & !is.na(post_fire_seeding))
```


Make new resprouter or seeder variable

```{r}
resprouting_seeding <-
  resprouting_seeding %>%
  mutate(
    resprouter_or_seeder = case_when(
      (resprouting == "mostly resprouts" | resprouting == "sometimes resprouts") & post_fire_seeding == "no post-fire seeding" ~ "resprouter",
      (post_fire_seeding == "some post-fire seeding" | post_fire_seeding == "robust post-fire seeding") & resprouting == "does not resprout" ~ "seeder",
      (resprouting == "mostly resprouts" | resprouting == "sometimes resprouts") & (post_fire_seeding == "some post-fire seeding" | post_fire_seeding == "robust post-fire seeding") ~ "facultative",
      (resprouting == "does not resprout" | is.na(resprouting)) & post_fire_seeding == "no post-fire seeding" ~ "neither")
  )
```


Clean data

```{r}
# Remove hybrids
resprouting_seeding <-
  resprouting_seeding %>%
  filter(!str_detect(taxon_name, pattern = " x "))

# Make variables ordered factors
resprouting_seeding$resprouting <-
  factor(
    resprouting_seeding$resprouting,
    ordered = TRUE,
    levels = c("does not resprout", "sometimes resprouts", "mostly resprouts")
  )

resprouting_seeding$post_fire_seeding <-
  factor(
    resprouting_seeding$post_fire_seeding,
    ordered = TRUE,
    levels = c("no post-fire seeding", "some post-fire seeding", "robust post-fire seeding")
  )

resprouting_seeding$resprouter_or_seeder <-
  factor(
    resprouting_seeding$resprouter_or_seeder,
    ordered = TRUE,
    levels = c("resprouter", "seeder", "facultative", "neither")
  )
```


Summarise values

```{r}
# Filter for taxa with data on both variables
resprouting_seeding %>%
  filter(data_on_both == TRUE) %>%
  group_by(resprouter_or_seeder) %>%
  summarise(count = n()) %>%
  ungroup()
```


# Extract woodiness data

**Use Lizzy's curated dataset instead (Wenk_2022)**

```{r}
woodiness <-
  austraits$traits %>%
  filter(dataset_id == "Wenk_2022", trait_name == "woodiness_detailed") %>%
  pivot_wider(values_from = value, names_from = trait_name) %>%
  select(taxon_name, woodiness_detailed)
```


```{r}
resprouting_seeding_woodiness <-
  resprouting_seeding %>%
  left_join(woodiness, by = "taxon_name")

woody <- c("woody", "semi_woody woody", "woody woody_base")
herbaceous <- c("herbaceous", "herbaceous woody", "herbaceous woody_base", "herbaceous woody_root",
                "herbaceous semi_woody", "herbaceous semi_woody woody")
semi_woody <- c("semi_woody", "semi_woody woody_base")
unclear <- c("woody_like_stem", "woody_base")

resprouting_seeding_woodiness <-
  resprouting_seeding_woodiness %>%
  mutate(
    woody_or_herb = case_when(
      woodiness_detailed %in% woody ~ "woody",
      woodiness_detailed %in% herbaceous ~ "herb",
      woodiness_detailed %in% semi_woody ~ "semi-woody",
      woodiness_detailed %in% unclear ~ "unclear")
    )

resprouting_seeding_woodiness %>%
  group_by(woody_or_herb) %>%
  summarise(count = n()) %>%
  ungroup()

# 4641 herbs, 6390 woody, 67 semi-woody, 47 unclear, 16 NA
```



Make binomial fire response variables

```{r}
resprouting_seeding <-
  resprouting_seeding_woodiness %>%
  mutate(
    resprouting_binomial = case_when(
      resprouting %in% c("mostly resprouts", "sometimes resprouts") ~ TRUE,
      resprouting %in% c("does not resprout") ~ FALSE),
    seeding_binomial = case_when(
      post_fire_seeding %in% c("robust post-fire seeding", "some post-fire seeding") ~ TRUE,
      post_fire_seeding %in% c("no post-fire seeding") ~ FALSE),
    obligate_resprouting_or_seeding = if_else(
      data_on_both == TRUE, case_when(
        resprouter_or_seeder == "seeder" ~ "obligate seeder",
        resprouter_or_seeder == "resprouter" ~ "obligate resprouter"
        ), NA_character_),
    resprouting_seeding_binomial = if_else(data_on_both == TRUE, case_when(
      resprouting == "mostly resprouts" & post_fire_seeding == "some post-fire seeding" ~ "resprouter",
      resprouting == "sometimes resprouts" & post_fire_seeding == "robust post-fire seeding" ~ "seeder",
      resprouter_or_seeder == "seeder" ~ "seeder",
      resprouter_or_seeder == "resprouter" ~ "resprouter"
      ), NA_character_)
  ) |>
  distinct()
```


```{r}
saveRDS(resprouting_seeding, "data/resprouting_seeding.rds")
resprouting_seeding <- read_rds("data/resprouting_seeding.rds")

fire_response_woody_herb_only <-
  resprouting_seeding %>%
  filter(woody_or_herb %in% c("woody", "herb"))
saveRDS(
  fire_response_woody_herb_only, "data/fire_response_woody_herb_only.rds")
fire_response_woody_herb_only <-
  read_rds("data/fire_response_woody_herb_only.rds")
```


---

# Read in gbif data

The GBIF download and cleaning process is in R/gbif_process.R

```{r}
gbif <- read.csv("data/filtered_aus_obs.csv", skipNul = TRUE)
```

## Try use ausflora to align GBIF names with APC names

```{r}
#remotes::install_github("traitecoevo/APCalign")
library(APCalign)
```


```{r}
gbif_taxa <- gbif$verbatimscientificname %>% unique()
gbif_taxa <- gbif_taxa[gbif_taxa != "Microcarpaea  R.Br."]

resources <- load_taxonomic_resources()

lookup_table <- create_taxonomic_update_lookup(
  taxa = gbif_taxa,
  resources = resources
)

lookup_table %>% saveRDS("data/lookup_table.rds")
lookup_table <- readRDS("data/lookup_table.rds")
```


```{r}
gbif <-
  gbif %>%
  left_join(lookup_table, by = c("verbatimscientificname" = "original_name"))
```


## Join gbif species occurrences to austraits data

```{r}
# Does not include austraits taxa that cannot be matched to gbif records
joined_fire_data <-
  resprouting_seeding %>%
  inner_join(gbif, by = c("taxon_name" = "accepted_name"))

joined_fire_data %>% saveRDS("data/joined_fire_data.rds")
joined_fire_data <- readRDS("data/joined_fire_data.rds")

resprouting_seeding_taxa <- resprouting_seeding$taxon_name %>% unique()
joined_fire_data_taxa <- joined_fire_data$taxon_name %>% unique()
# Lost 280 taxa
```



# Maps of resprouting and seeding

Try new method like with Australian annuals project

```{r}
joined_fire_data_vect <- joined_fire_data %>% terra::vect(geom = c("decimalLongitude", "decimalLatitude"))
base_raster <- joined_fire_data_vect %>% terra::ext() %>% terra::rast()
terra::res(base_raster) <- 2

joined_fire_data$cell <-
  terra::extract(
    base_raster,
    cbind(joined_fire_data$decimalLongitude, joined_fire_data$decimalLatitude),
    cell = TRUE)$cell # Tracking cell numbers

# Take single record per species per cell, to avoid oversampling
data_grid <-
  joined_fire_data %>%
  group_by(cell, taxon_name) %>%
  mutate(n_obs = n()) %>% # Not summarise, keep all columns
  slice(1) %>%
  ungroup()

# Replace "x", which is the location of the obs, with grid-based x which is better just for plotting
xy_from_grid <- as_tibble(terra::xyFromCell(base_raster, data_grid$cell))
data_grid$x <- xy_from_grid$x
data_grid$y <- xy_from_grid$y

# Save for later use in analysis
data_grid %>% saveRDS("outputs/data_grid_2_res.rds")
```

```{r}
data_grid <-
  readRDS("outputs/data_grid_2_res.rds")

output_grid <-
  data_grid %>%
  filter(
    woody_or_herb %in% c("herb", "woody"),
    x < 155, y > -45
  ) %>%
  # Summaries for each cell
  group_by(cell, woody_or_herb) %>%
  summarise(
    x = x[1],
    y = y[1],
    n_obs = sum(n_obs),
    n_species = n_distinct(taxon_name),
    n_resprouters = sum(resprouting_binomial, na.rm = TRUE),
    n_seeders = sum(seeding_binomial, na.rm = TRUE),
    frac_resprouters = n_resprouters / n_species,
    frac_seeders = n_seeders / n_species
  ) %>%
  ungroup()
```

```{r}
library(patchwork)
data_ozmaps <- ozmaps::ozmap_country
data_sea <- rnaturalearth::ne_download(scale = 10, type = "ocean", category = "physical", returnclass = "sf")
data_woody <- output_grid %>% filter(woody_or_herb == "woody")
data_herb <- output_grid %>% filter(woody_or_herb == "herb")

n_woody_resprouters <-
  ggplot(data_ozmaps) +
  geom_tile(data = data_woody, aes(x, y, fill = n_resprouters)) +
  geom_sf(data = data_sea, fill = "white", colour = NA) +
  geom_sf(fill = NA, colour = "#1a1a1a", linewidth = 0.4) +
  coord_sf(ylim = c(-45, -10), xlim = c(110, 155), expand = FALSE) +
  scale_fill_continuous(type = "viridis", na.value = NA) +
  labs(fill = "Number of\nresprouting species", x = "Longitude", y = "Latitude", tag = "a)") +
  theme_bw() +
  theme(plot.tag = element_text()) +
  ggtitle("Woody")

n_herb_resprouters <-
  ggplot(data_ozmaps) +
  geom_tile(data = data_herb, aes(x, y, fill = n_resprouters)) +
  geom_sf(data = data_sea, fill = "white", colour = NA) +
  geom_sf(fill = NA, colour = "#1a1a1a", linewidth = 0.4) +
  coord_sf(ylim = c(-45, -10), xlim = c(110, 155), expand = FALSE) +
  scale_fill_continuous(type = "viridis", na.value = NA) +
  labs(fill = "Number of\nresprouting species", x = "Longitude", y = "Latitude", tag = "a)") +
  theme_bw() +
  theme(plot.tag = element_text()) +
  ggtitle("Herbaceous")

n_woody_resprouters + n_herb_resprouters
ggsave("figures/sr_resprouters.png", width = 8.5, height = 3)
```

```{r}
frac_woody_resprouters <-
  ggplot(data_ozmaps) +
  geom_tile(
    data = data_woody |> filter(n_obs >= 10, frac_resprouters >= 0.4, y < -10), # 0.3 value is outside of border
    aes(x, y, fill = frac_resprouters), width = 2.1, height = 2.1) +
  geom_sf(data = data_sea, fill = "white", colour = NA) +
  geom_sf(fill = NA, colour = "#1a1a1a", linewidth = 0.2) +
  coord_sf(ylim = c(-45, -8), xlim = c(110, 157), expand = FALSE) +
  scale_x_continuous(breaks = seq(110, 150, by = 20)) +
  scale_y_continuous(breaks = seq(-40, -10, by = 10)) +
  scale_fill_continuous(type = "viridis", na.value = NA, limits = c(0.3, 0.95)) +
  theme_bw() +
  labs(subtitle = "Woody") +
  ggtitle("a) Proportion of resprouting taxa") +
  theme(
    plot.title = element_text(size = 13, margin = margin(0, 0, 10, 0)),
    plot.title.position = "plot",
    plot.subtitle = element_text(hjust = 0.5, size = 11),
    axis.title = element_blank(),
    axis.text = element_text(size = 7, colour = "#1a1a1a"),
    legend.position = "none",
    panel.grid = element_blank(),
    plot.background = element_rect(fill = NA, colour = NA)
  )

frac_herbaceous_resprouters <-
  ggplot(data_ozmaps) +
  geom_tile(
    data = data_herb |> filter(n_obs >= 10, frac_resprouters >= 0.3, y < -10),
    aes(x, y, fill = frac_resprouters), width = 2.1, height = 2.1) +
  geom_sf(data = data_sea, fill = "white", colour = NA) +
  geom_sf(fill = NA, colour = "#1a1a1a", linewidth = 0.2) +
  coord_sf(ylim = c(-45, -8), xlim = c(110, 157), expand = FALSE) +
  scale_x_continuous(breaks = seq(110, 150, by = 20)) +
  scale_y_continuous(breaks = seq(-40, -10, by = 10)) +
  scale_fill_continuous(
    type = "viridis",
    na.value = NA,
    limits = c(0.3, 0.95),
    breaks = seq(0.3, 0.9, by = 0.2)
  ) +
  theme_bw() +
  labs(subtitle = "Herbaceous") +
  guides(
    fill = guide_colourbar(
      barwidth = 0.7,
      barheight = 8,
      frame.linewidth = 0.3,
      frame.colour = "#1a1a1a",
      ticks.linewidth = 0.3,
      ticks.colour = "#1a1a1a")
  ) +
  theme(
    plot.subtitle = element_text(hjust = 0.5, size = 11),
    axis.title = element_blank(),
    axis.text = element_text(size = 7, colour = "#1a1a1a"),
    legend.title = element_blank(),
    panel.grid = element_blank(),
    plot.background = element_rect(fill = NA, colour = NA)
  )

frac_woody_resprouters + frac_herbaceous_resprouters
ggsave("figures/frac_resprouters.png", width = 8.5, height = 3)
```


## Calculate number of fires per pixel

Get list of rasters from MODIS

Win20 = Australia
A list of geoTIFFs for each year and month
See pg 15 of the documentation/user guide for MCD64:
https://lpdaac.usgs.gov/documents/875/MCD64_User_Guide_V6.pdf

Instructions for obtaining the data begins on pg 22.

```{r}
list_of_files <-
  list.files("data/MODIS_burn_data_australia/Win20", recursive = TRUE, full.names = TRUE) %>%
  str_subset("burndate") # Filter for the burndate layers not the QA layers
```


```{r}
# Function to extract burn dates
extract_burn_dates <- function(file_path) {

  # `file_path` argument must be a string
  raster <- terra::rast(file_path)
  # Extract year from file_path
  year <- str_match(file_path, "(?<=.A)[0-9]{4}")
  # Convert into a data.frame
  df <-
    raster %>%
    as.data.frame(xy = TRUE) %>%
    rename(burn_date = contains("burndate")) %>%
    # Filter out -2 and 0 values
    filter(burn_date > 0) %>%
    # Convert Julian date to calendar date
    mutate(burn_date = as.Date(burn_date - 1, origin = str_glue("{year}-01-01")))
  # Remove raster and run garbage collection
  remove(raster)
  gc()
  # Return the data.frame
  return(df)

}
```


Apply function to extract burn dates

```{r}
burn_dates_df <- map_dfr(list_of_files, extract_burn_dates)
burn_dates_df <- burn_dates_df %>% arrange(x, y, burn_date)
```


Use group_by and summarise to count the number of burns per pixel

```{r}
num_of_fires_df <-
  burn_dates_df %>%
  group_by(x, y) %>%
  summarise(num_of_fires = sum(burn_date != 0)) %>%
  ungroup()
```


```{r}
init_raster <- terra::rast("data/MODIS_burn_data_australia/Win20/2000/MCD64monthly.A2000306.Win20.006.burndate.tif")
# Resolution = 0.004
# Extent: 112, 155.0005, 0.004394531, -10 (xmin, xmax, ymin, ymax)

init_df <- init_raster %>% as.data.frame(xy = TRUE)
init_df <-
  init_df %>%
  rename(burn_date = contains("burndate")) %>%
  filter(burn_date >= 0) %>%
  arrange(x, y) %>%
  select(x, y)
full_fires_df <- full_join(init_df, num_of_fires_df, by = c("x", "y"))

nrow(full_fires_df)
nrow(init_df)
nrow(num_of_fires_df)
# 18.6% of pixels burned (but this includes water pixels)

full_fires_df <- full_fires_df %>% replace_na(list(num_of_fires = 0))

# Filter out values outside of Australia
full_fires_no_water <-
  full_fires_df %>%
  filter(x > 143 | y > -39) %>%
  filter(x > 120 | y > -36) %>%
  filter(x > 126 | y < -12) %>%
  filter(x < 145 | y < -14)

num_fires_raster <- terra::rast(full_fires_no_water, type = "xyz")

terra::writeRaster(num_fires_raster, "data/num_fires_map.tif", overwrite = TRUE)
```


```{r}
num_fires <- terra::rast("data/num_fires_map.tif")
num_fires <- num_fires %>% as.data.frame(xy = TRUE) |> filter(!is.na(num_of_fires))

num_fires_plot <-
  ggplot(num_fires) +
  geom_raster(aes(x = x, y = y, fill = num_of_fires), interpolate = FALSE) +
  geom_sf(data = data_sea, fill = "white", colour = NA) +
  geom_sf(data = data_ozmaps, fill = NA, colour = "#1a1a1a", linewidth = 0.2) +
  coord_sf(ylim = c(-45, -8), xlim = c(110, 157), expand = FALSE) +
  scale_x_continuous(breaks = seq(110, 150, by = 20)) +
  scale_y_continuous(breaks = seq(-40, -10, by = 10)) +
  scale_fill_distiller(
    palette = "YlOrRd",
    direction = 1,
    trans = "log",
    na.value = "white",
    breaks = c(1, 3, 8, 21)
  ) +
  guides(
    fill = guide_colourbar(
      barwidth = 0.7,
      barheight = 8,
      frame.linewidth = 0.3,
      frame.colour = "#1a1a1a",
      ticks.linewidth = 0.3,
      ticks.colour = "#1a1a1a")) +
  ggtitle("b) Number of fires") +
  theme_bw() +
  theme(
    plot.subtitle = element_text(hjust = 0.5, size = 11),
    axis.title = element_blank(),
    axis.text = element_text(size = 7, colour = "#1a1a1a"),
    legend.title = element_blank(),
    panel.grid = element_blank(),
    plot.background = element_rect(fill = NA, colour = NA)
  )
num_fires_plot

frac_woody_resprouters + frac_herbaceous_resprouters + num_fires_plot

ggsave("figures/frac_resprouters_and_num_fires.png", width = 13, height = 3)
```

```{r}
library(grid)
grid.newpage()
grid.draw(cbind(ggplotGrob(p1), ggplotGrob(p2)))
grid.draw(cbind(ggplotGrob(p1), ggplotGrob(p2), ggplotGrob(num_fires_plot)))
ggarrange(cbind(ggplotGrob(p1), ggplotGrob(p2)), num_fires_plot, ncol = 2, nrow = 1)

plot <- cbind(ggplotGrob(p1), ggplotGrob(p2), ggplotGrob(num_fires_plot))
ggsave("maps.png", plot = plot, width = 8, height = 2.4)

p3 <- cbind(ggplotGrob(p1), ggplotGrob(p2))
ggarrange(p3, num_fires_plot, ncol = 2, nrow = 1, heights = c(2, 1)) # Fiddle with heights argument or align argument
# common.legend = TRUE
# legend = "bottom"
# labels = c("a)", "b)")
# It's because you are plotting p3 as one plot and the other as another so they try make them the same size
plot <- ggarrange(p1, p2, num_fires_plot, ncol = 3, nrow = 1)
ggsave("maps.png", plot = plot, width = 10, height = 3)
ggarrange(
  p1, p2, num_fires_plot,
  ncol = 3, nrow = 1,
  labels = c("a) Proportion of Resprouting Taxa", "", "b) Number of Fires")
)
ggarrange(
  ggarrange(p1, p2, ncol = 2, nrow = 1, common.legend = TRUE, legend = "right"), num_fires_plot,
  ncol = 2, nrow = 1, labels = c("a) Proportion of Resprouting Taxa", "b) Number of Fires")
)
```

