---
title: "fire_response_strategy_and_predictors"
author: "Sophie"
date: "22/11/2021"
output: html_document
editor_options:
  chunk_output_type: console
---


# Fire response strategy and fire frequency

Load libraries

```{r}
library(tidyverse)
```


Read in fire response strategy data

```{r}
resprouting_seeding <- readRDS("data/resprouting_seeding.rds")
```


Number of resprouters and seeders in woody and herb taxa

```{r}
resprouting_seeding_woody_herb_only <- readRDS("data/fire_response_woody_herb_only.rds")

resprouting_seeding_woody_herb_only %>%
  drop_na(resprouting_binomial) %>%
  group_by(resprouting_binomial, woody_or_herb) %>%
  summarise(count = n()) %>%
  ggplot(aes(x = resprouting_binomial, y = count)) +
  geom_col() +
  facet_wrap(~woody_or_herb) +
  theme_classic()
# 4022 (64.1%) resprouting and 2254 non-resprouting in woody taxa, 3006 resprouting (65.3%) and 1599
# non-resprouting in herby taxa

resprouting_seeding_woody_herb_only %>%
  drop_na(seeding_binomial) %>%
  group_by(seeding_binomial, woody_or_herb) %>%
  summarise(count = n()) %>%
  ggplot(aes(x = seeding_binomial, y = count)) +
  geom_col() +
  facet_wrap(~woody_or_herb) +
  theme_classic()
# 2092 seeding and 569 non-seeding (78.6%) in woody taxa, 712 seeding (78.3%) and 197
# non-seeding in herby taxa
```



---

# Survival Analysis Method


Read in median fire return interval data from survival analysis

```{r}
median_fris <- read_csv("outputs/median_fris_with_survival_analysis.csv")

median_fris <- median_fris %>% left_join(resprouting_seeding, by = c("taxon_name"))


# Convert intervals to years not days
median_fris <- median_fris %>%
  mutate(across(
  c(med_w, med_w_lwr, med_w_upr, med_wo, med_wo_lwr, med_wo_upr),
  ~ .x / 365.25
  ))

median_fires_per_22_years <- median_fris %>%
  mutate(across(
    c(med_w, med_w_lwr, med_w_upr, med_wo, med_wo_lwr, med_wo_upr),
    ~ 22.17 / .x
  ))
```



# Median FRIs calculated with unburnt pixels

Filter out taxa with warning messages

```{r}
median_fris_w <-
  median_fris %>%
  filter(is.na(med_w_warnings)) %>%
  select(
    taxon_name, med_w, med_w_lwr, med_w_upr, num_unburnt_pixels, num_pixels,
    resprouting_binomial, seeding_binomial, data_on_both, woody_or_herb, taxon_rank
  )

# Maybe should do this more elegantly by filtering out values that have very large confidence intervals
```


Histogram of median FRIs

```{r}
hist(median_fris_w$med_w)
```


Plot

```{r}
median_fris_w %>%
  filter(!is.na(resprouting_binomial)) %>%
  group_by(resprouting_binomial) %>%
  mutate(n = n()) %>%
  mutate(label = paste0(resprouting_binomial, "\nN = ", n)) %>%
  ggplot(aes(med_w, factor(label))) +
  geom_point(alpha = 0.025) +
  scale_x_log10() +
  theme_classic()
```


---

# GLM Method


Read in mean fire return interval data from GLMs

```{r}
mean_fires <- read_csv("outputs/mean_fires_df_with_poisson_glm.csv", guess_max = 10000)

mean_fires <- mean_fires %>% left_join(resprouting_seeding, by = c("taxon_name"))
```


Filter out taxa with warning messages

```{r}
mean_fires_no_errors <-
  mean_fires %>%
  filter(is.na(warnings), is.na(error), is.na(messages)) %>%
  select(-warnings, -messages, -error)
```


```{r}
write_csv(mean_fires_no_errors, "outputs/final_df_predicted_fris_and_fire_response.csv")
mean_fires_no_errors <- read_csv("outputs/final_df_predicted_fris_and_fire_response.csv")
```


Histogram of mean FRIs

```{r}
hist(mean_fires_no_errors$mean_fri)
```


## New Stacked Plots


Binary resprouting variable

```{r}
mean_fires_no_errors %>%
  drop_na(resprouting_binomial) %>%
  filter(woody_or_herb %in% c("woody", "herb")) %>%
  ggplot(aes(mean_fri, group = resprouting_binomial, fill = resprouting_binomial)) +
  geom_density(alpha = 0.6) +
  facet_wrap(~woody_or_herb) +
  scale_x_log10() +
  scale_fill_brewer(palette = "Set1") +
  theme_classic()

mean_fires_no_errors %>%
  drop_na(resprouting_binomial) %>%
  ggplot(aes(mean_fri, group = resprouting_binomial, fill = resprouting_binomial)) +
  geom_density(adjust = 1.5, position = "fill", alpha = 0.6) +
  scale_x_log10() +
  scale_fill_brewer(palette = "Set1") +
  theme_classic()

mean_fires_no_errors %>%
  drop_na(resprouting_binomial) %>%
  ggplot(aes(mean_fri, fill = resprouting_binomial)) +
  geom_histogram(alpha = 0.6, position = "identity", binwidth = 0.1) +
  scale_x_log10() +
  scale_fill_brewer(palette = "Set1") +
  theme_classic()

mean_fires_no_errors %>%
  drop_na(resprouting_binomial) %>%
  filter(woody_or_herb %in% c("woody", "herb")) %>%
  ggplot(aes(mean_fri, fill = resprouting_binomial)) +
  geom_histogram(alpha = 0.6, position = "identity", binwidth = 0.1) +
  facet_wrap(~woody_or_herb) +
  scale_x_log10() +
  scale_fill_brewer(palette = "Set1") +
  theme_classic()

mean_fires_no_errors %>%
  drop_na(resprouting_binomial) %>%
  filter(mean_fri < 6000) %>%
  ggplot(aes(mean_fri, fill = resprouting_binomial)) +
  geom_histogram(alpha = 0.6, position = "fill", binwidth = 0.1) +
  scale_x_log10() +
  scale_fill_brewer(palette = "Set1") +
  theme_classic()

mean_fires_no_errors %>%
  drop_na(resprouting_binomial) %>%
  filter(mean_fri < 6000) %>%
  filter(woody_or_herb %in% c("woody", "herb")) %>%
  ggplot(aes(mean_fri, fill = resprouting_binomial)) +
  geom_histogram(alpha = 0.6, position = "fill", binwidth = 0.2) +
  facet_wrap(~woody_or_herb) +
  scale_x_log10() +
  scale_fill_brewer(palette = "Set1") +
  theme_classic()

mean_fires_no_errors %>%
  drop_na(resprouting_binomial) %>%
  filter(woody_or_herb %in% c("woody", "herb")) %>%
  ggplot(aes(mean_fri, ..count.., fill = resprouting_binomial)) +
  geom_density(adjust = 1.5, position = "fill", alpha = 0.6) +
  facet_wrap(~woody_or_herb) +
  scale_x_log10() +
  scale_fill_brewer(palette = "Set1") +
  theme_classic()

mean_fires_no_errors %>%
  drop_na(resprouting_binomial) %>%
  ggplot(aes(mean_fri, ..count.., fill = resprouting_binomial)) +
  geom_density(adjust = 1.5, position = "fill", alpha = 0.6) +
  scale_x_log10() +
  scale_fill_brewer(palette = "Set1") +
  theme_classic()
```


Binary seeding variable

```{r}
mean_fires_no_errors %>%
  drop_na(seeding_binomial) %>%
  filter(woody_or_herb %in% c("woody", "herb")) %>%
  ggplot(aes(mean_fri, group = seeding_binomial, fill = seeding_binomial)) +
  geom_density(alpha = 0.6) +
  facet_wrap(~woody_or_herb) +
  scale_x_log10() +
  scale_fill_brewer(palette = "Set1") +
  theme_classic()

mean_fires_no_errors %>%
  drop_na(seeding_binomial) %>%
  ggplot(aes(mean_fri, group = seeding_binomial, fill = seeding_binomial)) +
  geom_density(adjust = 1.5, position = "fill", alpha = 0.6) +
  scale_x_log10() +
  scale_fill_brewer(palette = "Set1") +
  theme_classic()

mean_fires_no_errors %>%
  drop_na(seeding_binomial) %>%
  ggplot(aes(mean_fri, fill = seeding_binomial)) +
  geom_histogram(alpha = 0.6, position = "identity", binwidth = 0.1) +
  scale_x_log10() +
  scale_fill_brewer(palette = "Set1") +
  theme_classic()

mean_fires_no_errors %>%
  drop_na(seeding_binomial) %>%
  filter(woody_or_herb %in% c("woody", "herb")) %>%
  ggplot(aes(mean_fri, fill = seeding_binomial)) +
  geom_histogram(alpha = 0.6, position = "identity", binwidth = 0.1) +
  facet_wrap(~woody_or_herb) +
  scale_x_log10() +
  scale_fill_brewer(palette = "Set1") +
  theme_classic()

mean_fires_no_errors %>%
  drop_na(seeding_binomial) %>%
#  filter(mean_fri < 1100) %>% # Filter out high mean FRIs
  ggplot(aes(mean_fri, fill = seeding_binomial)) +
  geom_histogram(alpha = 0.6, position = "fill", binwidth = 0.1) +
  scale_x_log10() +
  scale_fill_brewer(palette = "Set1") +
  theme_classic()

mean_fires_no_errors %>%
  drop_na(seeding_binomial) %>%
  filter(woody_or_herb %in% c("woody", "herb")) %>%
  ggplot(aes(mean_fri, fill = seeding_binomial)) +
  geom_histogram(alpha = 0.6, position = "fill", binwidth = 0.2) +
  facet_wrap(~woody_or_herb) +
  scale_x_log10() +
  scale_fill_brewer(palette = "Set1") +
  theme_classic()

mean_fires_no_errors %>%
  drop_na(seeding_binomial) %>%
  filter(woody_or_herb %in% c("woody", "herb")) %>%
  ggplot(aes(mean_fri, ..count.., fill = seeding_binomial)) +
  geom_density(adjust = 1.5, position = "fill", alpha = 0.6) +
  facet_wrap(~woody_or_herb) +
  scale_x_log10() +
  scale_fill_brewer(palette = "Set1") +
  theme_classic()

mean_fires_no_errors %>%
  drop_na(seeding_binomial) %>%
  ggplot(aes(mean_fri, ..count.., fill = seeding_binomial)) +
  geom_density(adjust = 1.5, position = "fill", alpha = 0.6) +
  scale_x_log10() +
  scale_fill_brewer(palette = "Set1") +
  theme_classic()
```


---

# Binomial logistic regression


## Resprouting binomial

Clean mean FRI data

```{r}
# Eliminate species with less than 10 pixels
mean_fires_clean <- mean_fires_no_errors |>
  dplyr::filter(num_pixels >= 10) |> # Removed 284
  # Convert to mean number of fires in 100 years, instead of mean number of fires in the sampling period
  mutate(mean_fires_100 = mean_fires / 22.17 * 100)

df_woody_herb_only <-
  mean_fires_clean %>%
  filter(woody_or_herb %in% c("woody", "herb"))

df_woody_herb_only$resprouting_binomial <- df_woody_herb_only$resprouting_binomial %>% as.factor()
df_woody_herb_only$seeding_binomial <- df_woody_herb_only$seeding_binomial %>% as.factor()
```


```{r}
library(ggeffects)
library(RColorBrewer)
library(sjPlot)

m1 <- glm(
  resprouting_binomial ~
    log10(mean_fires_100) + I(log10(mean_fires_100)^2) + log10(mean_fires_100):woody_or_herb +
    I(log10(mean_fires_100)^2):woody_or_herb + woody_or_herb,
  data = df_woody_herb_only,
  family = binomial
)

summary(m1)

ggpredict(m1, c("mean_fires_100", "woody_or_herb")) %>%
  ggplot(aes(x, predicted, color = group)) +
  geom_line(size = 1) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group), alpha = 0.2, linetype = 0) +
  scale_x_log10() +
  ylab("Resprouting Probability") +
  xlab("Mean fires") +
  theme_classic()

tab_model(m1)
```


```{r}
coef2 <- coef(m1)

x1 <- log10(df_woody_herb_only$mean_fires_100)
x2 <- df_woody_herb_only$woody_or_herb
x2 <- ifelse(x2 == "woody", 1, 0)
pred_y <- coef2[1] + coef2[2] * x1 + coef2[3] * x1^2 + coef2[4] * x2 + coef2[5] * x1 * x2 + coef2[6] * x1^2 * x2
pred_p <- exp(pred_y) / (1 + exp(pred_y))
df1 <- data.frame(x1, x2, df_woody_herb_only$woody_or_herb, pred_y, pred_p)
df1 <- df1 %>% rename(woody_or_herb = df_woody_herb_only.woody_or_herb)

p1 <-
  ggplot(data = df1, aes(x = 10^x1, y = pred_p, col = woody_or_herb)) +
  geom_line(size = 1.3) +
  scale_x_log10(
    expand = c(0, 0),
    limits = c(0.008, 100)
  ) +
  scale_y_continuous(
    limits = c(0, 0.8),
    expand = c(0, 0)) +
  ylab("Resprouting Probability") +
  xlab("Fire Frequency (Per Century)") +
  scale_colour_manual(values = c("#915330", "#aedb65"),
            name = "",
            breaks = c("woody", "herb"),
            labels = c("Woody", "Herb")) +
  ggtitle("a)") +
  theme_classic() +
  theme(panel.border = element_rect(colour = "black", fill = NA),
      legend.position = "none",
      axis.text = element_text(colour = "black", size = 8),
      plot.margin = unit(c(0.1, 0.5, 0.1, 0.1), "cm"),
      plot.title.position = "plot",
      plot.title = element_text(size = 13, margin = margin(0, 0, 10, 0)),
      axis.title.y = element_text(size = 10, margin = margin(0, 8, 0, 0)),
      axis.title.x = element_text(size = 10, margin = margin(8, 0, 0, 0)),
      panel.grid.major = element_line(colour = "#ebebeb", size = 0.5),
      axis.line = element_blank())

p1
```


Density distributions

```{r}
df_woody_herb_only$woody_or_herb <- factor(df_woody_herb_only$woody_or_herb,
                       levels = c("woody", "herb"))

means <- df_woody_herb_only %>%
  drop_na(resprouting_binomial) %>%
  group_by(woody_or_herb, resprouting_binomial) %>%
  summarise(mean = mean(mean_fri))

medians <- df_woody_herb_only %>%
  drop_na(resprouting_binomial) %>%
  group_by(woody_or_herb, resprouting_binomial) %>%
  summarise(median = median(mean_fri))


resprouting_density <-
  df_woody_herb_only %>%
  drop_na(resprouting_binomial) %>%
  ggplot(aes(mean_fri, fill = resprouting_binomial)) +
  geom_density(alpha = 0.6, size = 0.8, col = NA) +
  geom_vline(data = means, aes(xintercept = mean, col = resprouting_binomial),
         alpha = 1, size = 1, linetype = "longdash") +
  facet_wrap(~woody_or_herb, ncol = 1, scales = "free") +
  scale_x_log10(expand = c(0, 0),
          limits = c(1.05, 5000),
          breaks = c(10, 100, 1000),
          labels = c(10, 100, 1000)) +
  scale_y_continuous(expand = c(0, 0),
             limits = c(0, 1),
             breaks = c(0, 0.5, 1.0)) +
  scale_fill_manual(values = c("#EA294C", "#2166AC")) +
  scale_colour_manual(values = c("#EA294C", "#2166AC")) +
  xlab("Fire Return Interval (Yrs)") +
  ylab("Density") +
  theme_classic() +
  theme(axis.line = element_blank(),
      strip.background = element_blank(),
      strip.text = element_blank(),
      plot.margin = unit(c(0.5, 0.5, 0.1, 0.1), "cm"),
      panel.border = element_rect(colour = "black", fill = NA),
      panel.spacing = unit(0.5, "lines"),
      axis.text = element_text(colour = "black", size = 8),
      axis.title.y = element_text(size = 10, margin = margin(0, 8, 0, 0)),
      axis.title.x = element_text(size = 10, margin = margin(8, 0, 0, 0)),
      legend.position = "none")
resprouting_density
```




Check accuracy of model

```{r}
index <-
  sample(1:length(df_woody_herb_only$resprouting_binomial),
       0.7*length(df_woody_herb_only$resprouting_binomial))
training_x <- df_woody_herb_only[index, c("mean_fri", "woody_or_herb")]
training_y <- df_woody_herb_only[index, "resprouting_binomial"]
test_x <- df_woody_herb_only[-index, c("mean_fri", "woody_or_herb")]
test_y <- df_woody_herb_only[-index, "resprouting_binomial"]
```


```{r}
set.seed(1)

m1 <- glm(resprouting_binomial ~ poly(log10(mean_fri),2) * woody_or_herb,
      data = data.frame(training_x, training_y),
      family = binomial)
summary(m1)
predictions <- predict(m1, newdata = data.frame(test_x, test_y),
             newx = as.matrix(test_x), type = "link") # Make predictions

predictions_vs_real <- data.frame(test_x, test_y, predictions)
predictions_vs_real <-
  predictions_vs_real %>%
  mutate(preds_transformed = 1/(1 + exp(-predictions)),
       preds_binomial = case_when(preds_transformed >= 0.5 ~ TRUE,
                    preds_transformed < 0.5 ~ FALSE))
predictions_vs_real$preds_binomial <- predictions_vs_real$preds_binomial %>% as.factor()

predictions_vs_real %>%
  drop_na(resprouting_binomial) %>%
  ggplot(aes(resprouting_binomial, preds_transformed)) +
  geom_boxplot() +
  facet_wrap(~woody_or_herb) +
  theme_classic()

library(caret)
confusionMatrix(predictions_vs_real$preds_binomial, predictions_vs_real$resprouting_binomial)
```



Phylogenetic regression

```{r}
library(phylolm)
library(future)
library(ape)
plan(multisession)

tree <- read.tree("data/v0.1-big-seed-plant-trees/ALLMB.tre")
df_woody_herb_only$taxon_name <- gsub(" ", "_", df_woody_herb_only$taxon_name)
df_woody_herb_only <- df_woody_herb_only[df_woody_herb_only$taxon_name %in% tree$tip.label, ]
df_woody_herb_only$resprouting_binomial <-
  ifelse(df_woody_herb_only$resprouting_binomial == TRUE, 1, 0) # Make numeric

pruned_tree <- drop.tip(tree, setdiff(tree$tip.label, df_woody_herb_only$taxon_name))
pruned_tree <- as.phylo(pruned_tree)
```


With sorted data (?)

```{r}
sorted_data <- df_woody_herb_only[order(match(df_woody_herb_only$taxon_name, pruned_tree$tip.label)), ]

sorted_data <-
  sorted_data %>%
  select(taxon_name, mean_fri, resprouting_binomial, woody_or_herb) %>%
  column_to_rownames("taxon_name")


pglm2 <- phyloglm(resprouting_binomial ~ log10(mean_fri) + I(log10(mean_fri)^2) + log10(mean_fri):woody_or_herb + I(log10(mean_fri)^2):woody_or_herb + woody_or_herb,
         method = "logistic_MPLE",
         data = sorted_data,
         phy = pruned_tree,
         btol = 100,
         boot = 0)

summary(pglm2)
coef2 <- coef(pglm2)

x1 <- log10(sorted_data$mean_fri)
x2 <- sorted_data$woody_or_herb
x2 <- ifelse(x2 == "woody", 1, 0)
pred_y <- coef2[1] + coef2[2]*x1 + coef2[3]*x1^2 + coef2[4]*x2 + coef2[5]*x1*x2 + coef2[6]*x1^2*x2
pred_p <- exp(pred_y) / (1 + exp(pred_y))
df1 <- data.frame(x1, x2, sorted_data$woody_or_herb, pred_y, pred_p)
df1 <- df1 %>% rename(woody_or_herb = sorted_data.woody_or_herb)

p3 <-
  ggplot(data = df1, aes(x = 10^x1, y = pred_p, col = woody_or_herb)) +
  geom_line(size = 1.3) +
  scale_x_log10(breaks = c(10, 100, 1000),
          labels = c(10, 100, 1000)) +
  scale_y_continuous(limits = c(0,1.0),
             expand = c(0,0)) +
  ylab("Resprouting Probability") +
  xlab("Fire Return Interval (Years)") +
  scale_colour_manual(values = c("#915330", "#aedb65"),
            name = "",
            breaks = c("woody", "herb"),
            labels = c("Woody", "Herb")) +
  ggtitle("a)") +
  theme_classic() +
  theme(panel.border = element_rect(colour = "black", fill = NA),
      legend.position = "none",
      axis.text = element_text(colour = "black", size = 8),
      plot.margin = unit(c(0.1,0.5,0.1,0.1), "cm"),
      plot.title.position = "plot",
      plot.title = element_text(size = 13, margin = margin(0,0,10,0)),
      axis.title.y = element_text(size = 10, margin = margin(0,8,0,0)),
      axis.title.x = element_text(size = 10, margin = margin(8,0,0,0)),
      panel.grid.major = element_line(colour = "#ebebeb", size = 0.5),
      axis.line = element_blank())
p3
```


To calculate R2 Tjur

```{r}
pred <- fitted(pglm2)
y <- pglm2$y

categories <- unique(y)
mean1 <- mean(pred[which(y == categories[1])], na.rm = TRUE)
mean2 <- mean(pred[which(y == categories[2])], na.rm = TRUE)

abs(mean1 - mean2) # 0.028
```



```{r}
coef2 <- coef(pglm2)
plot.new()
plot(seq(-100, 100, by = 20), seq(0,1, by = 0.1), xlab="Mean FRI",ylab="Resprouting Probability")
curve(plogis(coef2[1] + coef2[2]*x), col = "red", add = TRUE)

sorted_data %>%
  ggplot(aes(x = mean_fri, y = resprouting_binomial)) +
  geom_jitter(height = 0.01, alpha = 0.25) +
  stat_smooth(method = "glm", method.args = list(family = binomial)) +
  #facet_wrap(~woody_or_herb) +
  scale_x_log10() +
  theme_classic()
```




## Seeding binomial


```{r}
df_woody_herb_only$seeding_binomial <-
  df_woody_herb_only$seeding_binomial %>% as.factor()

m2 <- glm(seeding_binomial ~ log10(mean_fri) + I(log10(mean_fri)^2) + log10(mean_fri):woody_or_herb + I(log10(mean_fri)^2):woody_or_herb + woody_or_herb,
      data = df_woody_herb_only,
      family = binomial)
summary(m2)
ggpredict(m2, c("mean_fri", "woody_or_herb")) %>% plot()

ggpredict(m2, c("mean_fri", "woody_or_herb")) %>%
  ggplot(aes(x, predicted, color = group)) +
  geom_line(size = 1) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group), alpha = 0.2, linetype = 0) +
  scale_x_log10() +
  ylab("Seeding Probability") +
  xlab("Mean FRI") +
  theme_classic()

#m2 <- glm(seeding_binomial ~ poly(log10(mean_fri),2),
#      data = df_woody_herb_only,
#      family = binomial)
#summary(m2)
ggpredict(m2, c("mean_fri")) %>%
  ggplot(aes(x, predicted)) +
  geom_line(size = 1) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2, linetype = 0) +
  scale_x_log10() +
  ylab("Seeding Probability") +
  xlab("Mean FRI") +
  theme_classic()
```


Model without herb data

```{r}
df_woody_only <-
  df_woody_herb_only %>%
  filter(woody_or_herb == "woody")

m2 <- glm(seeding_binomial ~ log10(mean_fri) + I(log10(mean_fri)^2),
      data = df_woody_only,
      family = binomial)
summary(m2)

ggpredict(m2, c("mean_fri")) %>%
  ggplot(aes(x, predicted, color = group)) +
  geom_line(size = 1) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group), alpha = 0.2, linetype = 0) +
  scale_x_log10() +
  ylab("Seeding Probability") +
  xlab("Mean FRI") +
  theme_classic()

coef2 <- coef(m2)

x1 <- log10(df_woody_only$mean_fri)
x2 <- df_woody_only$woody_or_herb
x2 <- ifelse(x2 == "woody", 1, 0)
pred_y <- coef2[1] + coef2[2]*x1 + coef2[3]*x1^2
pred_p <- exp(pred_y) / (1 + exp(pred_y))
df1 <- data.frame(x1, x2, df_woody_only$woody_or_herb, pred_y, pred_p)
df1 <- df1 %>% rename(woody_or_herb = df_woody_only.woody_or_herb)

p3 <-
  ggplot(data = df1, aes(x = 10^x1, y = pred_p, col = woody_or_herb)) +
  geom_line(size = 1.3) +
  scale_x_log10(expand = c(0,0),
          limits = c(1.05,5000),
          breaks = c(10,100,1000),
          labels = c(10,100,1000)) +
  scale_y_continuous(limits = c(0,1.0),
             expand = c(0,0)) +
  ylab("Seeding Probability") +
  xlab("Fire Return Interval (Years)") +
  scale_colour_manual(values = c("#915330"),
            name = "",
            breaks = c("woody"),
            labels = c("Woody")) +
  ggtitle("b)") +
  theme_classic() +
  theme(panel.border = element_rect(colour = "black", fill = NA),
      legend.position = "none",
      axis.text = element_text(colour = "black", size = 8),
      plot.margin = unit(c(0.1,0.5,0.1,0.1), "cm"),
      plot.title.position = "plot",
      plot.title = element_text(size = 13, margin = margin(0,0,10,0)),
      axis.title.y = element_text(size = 10, margin = margin(0,8,0,0)),
      axis.title.x = element_text(size = 10, margin = margin(8,0,0,0)),
      panel.grid.major = element_line(colour = "#ebebeb", size = 0.5),
      axis.line = element_blank())

p3
```


Density distributions

```{r}
means <- df_woody_herb_only %>%
  drop_na(seeding_binomial) %>%
  group_by(woody_or_herb, seeding_binomial) %>%
  summarise(mean = mean(mean_fri))

medians <- df_woody_herb_only %>%
  drop_na(seeding_binomial) %>%
  group_by(woody_or_herb, seeding_binomial) %>%
  summarise(median = median(mean_fri))

seeding_density <-
  df_woody_herb_only %>%
  drop_na(seeding_binomial) %>%
  ggplot(aes(mean_fri, fill = seeding_binomial)) +
  geom_density(alpha = 0.6, size = 0.8, col = NA) +
  geom_vline(data = means, aes(xintercept = mean, col = seeding_binomial),
         alpha = 1, size = 1, linetype = "longdash") +
  facet_wrap(~woody_or_herb, ncol = 1, scales = "free") +
  scale_x_log10(expand = c(0,0),
          limits = c(1.05,5000),
          breaks = c(10,100,1000),
          labels = c(10,100,1000)) +
  scale_y_continuous(expand = c(0,0),
             limits = c(0,1),
             breaks = c(0,0.5,1.0)) +
  scale_fill_manual(name = "Resprouting/Seeding",
            values = c("#EA294C", "#2166AC"),
            labels = c("No", "Yes")) +
  scale_colour_manual(values = c("#EA294C", "#2166AC")) +
  guides(fill = guide_legend(ncol = 2)) +
  xlab("Fire Return Interval (Yrs)") +
  ylab("Density") +
  theme_classic() +
  theme(axis.line = element_blank(),
      strip.background = element_blank(),
      strip.text = element_blank(),
      plot.margin = unit(c(0.5,0.5,0.1,0.1), "cm"),
      panel.border = element_rect(colour = "black", fill = NA),
      panel.spacing = unit(0.5, "lines"),
      axis.text = element_text(colour = "black", size = 8),
      axis.title.y = element_text(size = 10, margin = margin(0,8,0,0)),
      axis.title.x = element_text(size = 10, margin = margin(8,0,0,0)),
      legend.position = "none")

seeding_density
```



```{r}
coef2 <- coef(m2)

x1 <- log10(df_woody_herb_only$mean_fri)
x2 <- df_woody_herb_only$woody_or_herb
x2 <- ifelse(x2 == "woody", 1, 0)
pred_y <- coef2[1] + coef2[2]*x1 + coef2[3]*x1^2 + coef2[4]*x2 + coef2[5]*x1*x2 + coef2[6]*x1^2*x2
pred_p <- exp(pred_y) / (1 + exp(pred_y))
df1 <- data.frame(x1, x2, df_woody_herb_only$woody_or_herb, pred_y, pred_p)
df1 <- df1 %>% rename(woody_or_herb = df_woody_herb_only.woody_or_herb)


p2 <-
  ggplot(data = df1, aes(x = 10^x1, y = pred_p, col = woody_or_herb)) +
  geom_line(size = 1.3) +
  scale_x_log10(breaks = c(10,100,1000),
          labels = c(10,100,1000)) +
  scale_y_continuous(limits = c(0,1.0),
             expand = c(0,0)) +
  ylab("Seeding Probability") +
  xlab("Mean FRI") +
  scale_colour_manual(values = c("#B15928", "#B2DF8A"),
            name = "",
            breaks = c("woody", "herb"),
            labels = c("Woody", "Herb")) +
  theme_classic() +
  theme(panel.border = element_rect(colour = "black", fill = NA),
      legend.position = "none",
      axis.text = element_text(colour = "black", size = 9),
      plot.margin = unit(c(0.1,0.5,0.1,0.1), "cm"),
      axis.title.y = element_text(size = 11, margin = margin(0,8,0,0)),
      axis.title.x = element_text(size = 11, margin = margin(8,0,0,0)))
p2
```


Plot resprouting and seeding probability graphs together

```{r}
library(grid)
grid.newpage()
grid.draw(cbind(ggplotGrob(p1), ggplotGrob(p3)))
final_plot <- cbind(ggplotGrob(p1), ggplotGrob(p3))
ggsave("resprouting_seeding_probabilites.png", plot = final_plot, width = 6, height = 2.8)

cbind(ggplotGrob(resprouting_density), rbind(ggplotGrob(seeding_density1), ggplotGrob(seeding_density2)))

ggsave("resprouting_densities.png", plot = resprouting_density, width = 3, height = 2.3)
ggsave("seeding_densities.png", plot = seeding_density, width = 3, height = 2.3)

```



```{r}
index <-
  sample(1:length(df_woody_herb_only$seeding_binomial),
       0.7*length(df_woody_herb_only$seeding_binomial))
training_x <- df_woody_herb_only[index, c("mean_fri", "woody_or_herb")]
training_y <- df_woody_herb_only[index, "seeding_binomial"]
test_x <- df_woody_herb_only[-index, c("mean_fri", "woody_or_herb")]
test_y <- df_woody_herb_only[-index, "seeding_binomial"]
```


```{r}
set.seed(1)

m1 <- glm(seeding_binomial ~ poly(log10(mean_fri),2) * woody_or_herb,
      data = data.frame(training_x, training_y),
      family = binomial)
summary(m1)
predictions <- predict(m1, newdata = data.frame(test_x, test_y),
             newx = as.matrix(test_x), type = "link") # Make predictions

predictions_vs_real <- data.frame(test_x, test_y, predictions)
predictions_vs_real <-
  predictions_vs_real %>%
  mutate(preds_transformed = 1/(1 + exp(-predictions)),
       preds_binomial = case_when(preds_transformed >= 0.5 ~ TRUE,
                    preds_transformed < 0.5 ~ FALSE))
predictions_vs_real$preds_binomial <- predictions_vs_real$preds_binomial %>% as.factor()

predictions_vs_real %>%
  drop_na(seeding_binomial) %>%
  ggplot(aes(seeding_binomial, preds_transformed)) +
  geom_boxplot() +
  facet_wrap(~woody_or_herb) +
  theme_classic()

library(caret)
confusionMatrix(predictions_vs_real$preds_binomial, predictions_vs_real$resprouting_binomial)
```


Phylogenetic logistic regression

```{r}
tree <- read.tree("raw_data/v0.1-big-seed-plant-trees/ALLMB.tre")
df_woody_herb_only$taxon_name <- gsub(" ", "_", df_woody_herb_only$taxon_name)
df_woody_herb_only <- df_woody_herb_only[df_woody_herb_only$taxon_name %in% tree$tip.label,]
df_woody_herb_only$seeding_binomial <-
  ifelse(df_woody_herb_only$seeding_binomial == TRUE, 1, 0) # make numeric

pruned_tree <- drop.tip(tree, setdiff(tree$tip.label, df_woody_herb_only$taxon_name))
pruned_tree <- as.phylo(pruned_tree)

sorted_data <- df_woody_herb_only[order(match(df_woody_herb_only$taxon_name, pruned_tree$tip.label)),]

sorted_data <-
  sorted_data %>%
  select(taxon_name, mean_fri, seeding_binomial, woody_or_herb) %>%
  column_to_rownames("taxon_name")


pglm1 <- phyloglm(seeding_binomial ~ log10(mean_fri) + I(log10(mean_fri)^2) + log10(mean_fri):woody_or_herb + I(log10(mean_fri)^2):woody_or_herb + woody_or_herb,
         method = "logistic_MPLE",
         data = sorted_data,
         phy = pruned_tree,
         btol = 100,
         boot = 0)
summary(pglm1)

coef2 <- coef(pglm1)

x1 <- log10(sorted_data$mean_fri)
x2 <- sorted_data$woody_or_herb
x2 <- ifelse(x2 == "woody", 1, 0)
pred_y <- coef2[1] + coef2[2]*x1 + coef2[3]*x1^2 + coef2[4]*x2 + coef2[5]*x1*x2 + coef2[6]*x1^2*x2
pred_p <- exp(pred_y) / (1 + exp(pred_y))
df1 <- data.frame(x1, x2, sorted_data$woody_or_herb, pred_y, pred_p)
df1 <- df1 %>% rename(woody_or_herb = sorted_data.woody_or_herb)


p6 <-
  ggplot(data = df1, aes(x = 10^x1, y = pred_p, col = woody_or_herb)) +
  geom_line(size = 1.3) +
  scale_x_log10(breaks = c(10,100,1000),
          labels = c(10,100,1000)) +
  scale_y_continuous(limits = c(0,1.0),
             expand = c(0,0)) +
  scale_colour_manual(values = c("#B15928", "#B2DF8A"),
            name = "",
            breaks = c("woody", "herb"),
            labels = c("Woody", "Herb")) +
  ylab("Seeding Probability") +
  xlab("Mean FRI") +
  theme_classic() +
  theme(panel.border = element_rect(colour = "black", fill = NA),
      legend.position = "none",
      axis.text = element_text(colour = "black", size = 9),
      plot.margin = unit(c(0.1,0.5,0.1,0.1), "cm"),
      axis.title.y = element_text(size = 11, margin = margin(0,8,0,0)),
      axis.title.x = element_text(size = 11, margin = margin(8,0,0,0)))
```


Model without herb data

```{r}
df_woody_only <-
  df_woody_herb_only %>%
  filter(woody_or_herb == "woody")

tree <- read.tree("raw_data/v0.1-big-seed-plant-trees/ALLMB.tre")
df_woody_only$taxon_name <- gsub(" ", "_", df_woody_only$taxon_name)
df_woody_only <- df_woody_only[df_woody_only$taxon_name %in% tree$tip.label,]
df_woody_only$seeding_binomial <-
  ifelse(df_woody_only$seeding_binomial == TRUE, 1, 0) # make numeric

pruned_tree <- drop.tip(tree, setdiff(tree$tip.label, df_woody_only$taxon_name))
pruned_tree <- as.phylo(pruned_tree)

sorted_data <- df_woody_only[order(match(df_woody_only$taxon_name, pruned_tree$tip.label)),]

sorted_data <-
  sorted_data %>%
  select(taxon_name, mean_fri, seeding_binomial, woody_or_herb) %>%
  column_to_rownames("taxon_name")


pglm3 <- phyloglm(seeding_binomial ~ log10(mean_fri) + I(log10(mean_fri)^2),
         method = "logistic_MPLE",
         data = sorted_data,
         phy = pruned_tree,
         btol = 100,
         boot = 0)
summary(pglm3)

coef2 <- coef(pglm3)

x1 <- log10(sorted_data$mean_fri)
x2 <- sorted_data$woody_or_herb
x2 <- ifelse(x2 == "woody", 1, 0)
pred_y <- coef2[1] + coef2[2]*x1 + coef2[3]*x1^2
pred_p <- exp(pred_y) / (1 + exp(pred_y))
df1 <- data.frame(x1, x2, sorted_data$woody_or_herb, pred_y, pred_p)
df1 <- df1 %>% rename(woody_or_herb = sorted_data.woody_or_herb)


p5 <-
  ggplot(data = df1, aes(x = 10^x1, y = pred_p, col = woody_or_herb)) +
  geom_line(size = 1.3) +
  scale_x_log10(breaks = c(10,100,1000),
          labels = c(10,100,1000)) +
  scale_y_continuous(limits = c(0,1.0),
             expand = c(0,0)) +
  scale_colour_manual(values = c("#915330"),
            name = "",
            breaks = c("woody"),
            labels = c("Woody")) +
  ggtitle("b)") +
  ylab("Seeding Probability") +
  xlab("Fire Return Interval (Years)") +
  theme_classic() +
  theme(panel.border = element_rect(colour = "black", fill = NA),
      legend.position = "none",
      axis.text = element_text(colour = "black", size = 8),
      plot.margin = unit(c(0.1,0.5,0.1,0.1), "cm"),
      plot.title.position = "plot",
      plot.title = element_text(size = 13, margin = margin(0,0,10,0)),
      axis.title.y = element_text(size = 10, margin = margin(0,8,0,0)),
      axis.title.x = element_text(size = 10, margin = margin(8,0,0,0)),
      panel.grid.major = element_line(colour = "#ebebeb", size = 0.5),
      axis.line = element_blank())
p5

# To calculate odds ratio, exp(estimate)


grid.newpage()
grid.draw(cbind(ggplotGrob(p3), ggplotGrob(p5)))

final_plot <- cbind(ggplotGrob(p3), ggplotGrob(p5))
ggsave("resprouting_seeding_probabilites_phylo.png", plot = final_plot, width = 6, height = 2.8)
```


To calculate R2 Tjur

```{r}
pred <- fitted(pglm3)
y <- pglm3$y

categories <- unique(y)
mean1 <- mean(pred[which(y == categories[1])], na.rm = TRUE)
mean2 <- mean(pred[which(y == categories[2])], na.rm = TRUE)

abs(mean1 - mean2) # 0.062
```




---

# Comparison of survival analysis and poisson regression methods

Out of 10871 taxa, 359 did not fit and they were all because all pixels were unburnt.
Out of the 10512 with no errors, 419 had confidence intervals bounding zero, which
were discarded, leaving us with 7296 taxa. There were 131 taxa with mean FRI greater
than 1000 years. 9940 taxa had confidence intervals less than 100 years big. #TODO update

```{r}
mean_fires <- read_csv("outputs/mean_fires_df_with_glm.csv", guess_max = 10000)

mean_fires %>% filter(mean_fri > 1000) %>% nrow()
mean_fires_clean %>% filter(upr_conf_int - lwr_conf_int < 100) %>% nrow()

y <- mean_fires_clean %>% mutate(confint_size = upr_conf_int - lwr_conf_int)
hist(y$confint_size)
```



Comparatively, the survival analysis resulted in 1443 that did not fit (incl. unburnt
pixels). The warning was "ran out of iterations and did not converge". This occurred
because most or all pixels were unburnt. There were 7217 taxa with median FRI greater
than 5000 years. 221 taxa had confidence intervals less than 100 years big. #TODO update

```{r}
median_fris <- read_csv("outputs/median_fris_with_survival_analysis.csv", guess_max = 10871)
median_fris <-
  median_fris %>%
  mutate(across(
    c("med_w", "med_w_lwr", "med_w_upr", "med_wo", "med_wo_lwr", "med_wo_upr"),
    ~ .x / 365.2422
  ))

median_fris %>% filter(med_w > 1000) %>% nrow()
median_fris %>% filter(med_w_upr - med_w_lwr < 100) %>% nrow()
x <- median_fris %>% mutate(confint_size = med_w_upr - med_w_lwr)
hist(x$confint_size)

median_fris %>% filter(!is.na(med_wo_warnings)) %>% nrow()
median_fris %>% filter(med_wo > 5000) %>% nrow()
```

1282 that did not fit (when not including unburnt pixels). There were 6719 taxa with
median FRI greater than 5000 years. #TODO update

