---
title: "fire_response_strategy_and_leaf_traits"
author: "Sophie"
date: "11/01/2022"
output: html_document
editor_options:
  chunk_output_type: console
---

# Fire response strategies and leaf traits

```{r}
library(tidyverse)
library(ggtext)
library(patchwork)

austraits <- read_rds("data/austraits-5.0.0.rds") # This is an updated rds file
```


Read in LMA and leaf N data

```{r}
lma <-
  austraits$traits %>%
  filter(trait_name == "leaf_mass_per_area") %>%
  pivot_wider(names_from = "trait_name", values_from = "value") |>
  mutate(leaf_mass_per_area = as.numeric(leaf_mass_per_area))

leaf_n <-
  austraits$traits %>%
  filter(trait_name == "leaf_N_per_dry_mass") %>%
  pivot_wider(names_from = "trait_name", values_from = "value") |>
  mutate(leaf_N_per_dry_mass = as.numeric(leaf_N_per_dry_mass))
```


Calculate mean lma and leaf N for each taxa

```{r}
mean_lma <-
  lma %>%
  group_by(taxon_name) %>%
  summarise(mean_lma = mean(leaf_mass_per_area)) %>%
  ungroup()

mean_leaf_n <-
  leaf_n %>%
  group_by(taxon_name) %>%
  summarise(mean_leaf_n = mean(leaf_N_per_dry_mass)) %>%
  ungroup()
```


Join to fire response

```{r}
fire_response_and_fire_frequency <- read_csv("outputs/final_df_predicted_fris_and_fire_response.csv")

fire_response_and_leaf_traits <-
  fire_response_and_fire_frequency %>%
  left_join(mean_lma) %>%
  left_join(mean_leaf_n)
```


Run linear models

```{r}
lma_resp_model <- lm(
  log10(mean_lma) ~ resprouting_binomial * woody_or_herb,
  data = fire_response_and_leaf_traits)
summary(lma_resp_model)

lma_resp_emmeans <- emmeans::emmeans(
  lma_resp_model,
  specs = pairwise ~ resprouting_binomial:woody_or_herb)
lma_resp_emmeans
lma_resp_emmeans$contrasts %>% confint()
```


```{r}
leaf_n_resp_model <- lm(
  log10(mean_leaf_n) ~ resprouting_binomial * woody_or_herb,
  data = fire_response_and_leaf_traits)
summary(leaf_n_resp_model)

leaf_n_resp_emmeans <- emmeans::emmeans(
  leaf_n_resp_model,
  specs = pairwise ~ resprouting_binomial:woody_or_herb)
leaf_n_resp_emmeans
leaf_n_resp_emmeans$contrasts %>% confint()
```


```{r}
lma_seed_model <- lm(
  log10(mean_lma) ~ seeding_binomial * woody_or_herb,
  data = fire_response_and_leaf_traits)
summary(lma_seed_model)

lma_seed_emmeans <- emmeans::emmeans(
  lma_seed_model,
  specs = pairwise ~ seeding_binomial:woody_or_herb)
lma_seed_emmeans
lma_seed_emmeans$contrasts %>% confint()
```


```{r}
leaf_n_seed_model <- lm(
  log10(mean_leaf_n) ~ seeding_binomial * woody_or_herb,
  data = fire_response_and_leaf_traits)
summary(leaf_n_seed_model)

leaf_n_seed_emmeans <- emmeans::emmeans(
  leaf_n_seed_model,
  specs = pairwise ~ seeding_binomial:woody_or_herb)
leaf_n_seed_emmeans
leaf_n_seed_emmeans$contrasts %>% confint()
```


Plot

```{r}
lma_resprouting_p_values <- tibble(
  label = c("0.045", "<0.0001"),
  woody_or_herb = c("woody", "herb"),
  x = c(0.5, 0.5),
  y = c(3.5, 3.5)
)

lma_resprouting_plot <-
  fire_response_and_leaf_traits %>%
  drop_na(resprouting_binomial) %>%
  ggplot(aes(x = resprouting_binomial, y = mean_lma, fill = woody_or_herb)) +
  geom_violin(linewidth = 0.8) +
  geom_boxplot(width = 0.05, linewidth = 0.8) +
  geom_text(
    data = lma_resprouting_p_values,
    aes(x = x, y = y, label = label),
    hjust = 0,
    vjust = 0,
    size = 4,
    fontface = "bold"
  ) +
  facet_wrap(
    ~ factor(woody_or_herb, ordered = TRUE, levels = c("woody", "herb")),
    labeller = as_labeller(c(
    "woody" = "Woody",
    "herb" = "Herbaceous"
  ))) +
  scale_y_log10() +
  scale_fill_manual(
    values = c("#7e5137", "#b9ce6f"),
    breaks = c("woody", "herb")) +
  labs(
    x = "Resprouting",
    y = expression(LMA~(g~m^-2))
  ) +
  theme_classic() +
  theme(
    panel.background = element_rect(colour = "black", fill = NA, size = 0.5),
    plot.margin = unit(c(0.1, 1, 0.05, 0.1), "cm"),
    axis.text.y = element_text(colour = "black", size = 10),
    axis.title.y = element_text(size = 12, margin = margin(0, 10, 0, 0)),
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    legend.position = "none",
    strip.background = element_rect(colour = "black", size = 0.5, fill = "#f0f0f0"),
    strip.text = element_text(size = 11, face = "bold"),
    panel.spacing = unit(0.2, "cm"),
    axis.line = element_blank()
  )
lma_resprouting_plot
```



```{r}
leaf_n_resprouting_plot <-
  fire_response_and_leaf_traits %>%
  drop_na(resprouting_binomial) %>%
  ggplot(aes(x = resprouting_binomial, y = mean_leaf_n, fill = woody_or_herb)) +
  geom_violin(size = 0.8) +
  geom_boxplot(width = 0.05, size = 0.8) +
  facet_wrap(
    ~ factor(woody_or_herb, ordered = TRUE, levels = c("woody", "herb")),
    labeller = as_labeller(c(
    "woody" = "Woody",
    "herb" = "Herbaceous"
  ))) +
  scale_y_log10() +
  scale_x_discrete(labels = c("Non-\nresprouters", "Resprouters")) +
  scale_fill_manual(
    values = c("#7e5137", "#b9ce6f"),
    breaks = c("woody", "herb")) +
  labs(
    x = "Resprouting",
    y = expression(Leaf~N~(mg~g^-1))
  ) +
  theme_classic() +
  theme(
    panel.background = element_rect(colour = "black", fill = NA, size = 0.5),
    plot.margin = unit(c(0.1, 1, 0.1, 0.1), "cm"),
    axis.text = element_text(colour = "black", size = 10),
    axis.title.y = element_text(size = 12, margin = margin(0, 10, 0, 0)),
    axis.title.x = element_text(size = 12, margin = margin(10, 0, 0, 0), face = "bold"),
    legend.position = "none",
    strip.background = element_rect(colour = "black", size = 0.5, fill = "#f0f0f0"),
    strip.text = element_text(size = 11, face = "bold"),
    panel.spacing = unit(0.2, "cm"),
    axis.line = element_blank()
  )
leaf_n_resprouting_plot
```




```{r}
lma_seeding_p_values <- tibble(
  label = c("<0.0001", "<0.0001"),
  woody_or_herb = c("woody", "herb"),
  x = c(0.5, 0.5),
  y = c(3.5, 3.5)
)

lma_seeding_plot <-
  fire_response_and_leaf_traits %>%
  drop_na(seeding_binomial) %>%
  ggplot(aes(x = seeding_binomial, y = mean_lma, fill = woody_or_herb)) +
  geom_violin(size = 0.8) +
  geom_boxplot(width = 0.05, size = 0.8) +
  geom_text(
    data = lma_seeding_p_values,
    aes(x = x, y = y, label = label),
    hjust = 0,
    vjust = 0,
    size = 4,
    fontface = "bold"
  ) +
  facet_wrap(
    ~ factor(woody_or_herb, ordered = TRUE, levels = c("woody", "herb")),
    labeller = as_labeller(c(
    "woody" = "Woody",
    "herb" = "Herbaceous"
  ))) +
  scale_y_log10() +
  scale_fill_manual(
    values = c("#7e5137", "#b9ce6f"),
    breaks = c("woody", "herb")) +
  labs(
    x = "Post-fire seeding"
  ) +
  theme_classic() +
  theme(
    panel.background = element_rect(colour = "black", fill = NA, size = 0.5),
    plot.margin = unit(c(0.1, 0.1, 0.05, 0.1), "cm"),
    axis.ticks.y = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.text = element_text(colour = "black", size = 10),
    axis.title.y = element_blank(),
    axis.title.x = element_blank(),
    legend.position = "none",
    strip.background = element_rect(colour = "black", size = 0.5, fill = "#f0f0f0"),
    strip.text = element_text(size = 11, face = "bold"),
    panel.spacing = unit(0.2, "cm"),
    axis.line = element_blank())
lma_seeding_plot
```


```{r}
annotations_p4 <- data.frame(label = c("p = 0.050", ""), woody_or_herb = c("herb", "woody"))
annotations_p4$woody_or_herb <- annotations_p4$woody_or_herb %>% factor(ordered = TRUE, levels = c("woody", "herb"))

leaf_n_seeding_plot <-
  fire_response_and_leaf_traits %>%
  drop_na(seeding_binomial) %>%
  ggplot(aes(x = seeding_binomial, y = mean_leaf_n, fill = woody_or_herb)) +
  geom_violin(size = 0.8) +
  geom_boxplot(width = 0.05, size = 0.8) +
#  geom_text(aes(x = 2.1, y = 2.1, label = label), data = annotations_leaf_n_seeding_plot,
#        size = 4, fontface = "bold") +
  facet_wrap(
    ~ factor(woody_or_herb, ordered = TRUE, levels = c("woody", "herb")),
    labeller = as_labeller(c(
    "woody" = "Woody",
    "herb" = "Herbaceous"
  ))) +
  scale_x_discrete(labels = c("Non-seeders", "Seeders")) +
  scale_y_log10() +
  scale_fill_manual(
    values = c("#7e5137", "#b9ce6f"),
    breaks = c("woody", "herb")) +
  labs(
    x = "Post-fire seeding",
    y = "Leaf N (mg/g)") +
  theme_classic() +
  theme(
    panel.background = element_rect(colour = "black", fill = NA, size = 0.5),
    plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"),
    axis.ticks.y = element_blank(),
    axis.text.y = element_blank(),
    axis.text = element_text(colour = "black", size = 10),
    axis.title.y = element_blank(),
    axis.title.x = element_text(size = 12, margin = margin(10, 0, 0, 0), face = "bold"),
    legend.position = "none",
    strip.background = element_rect(colour = "black", size = 0.5, fill = "#f0f0f0"),
    strip.text = element_text(size = 11, face = "bold"),
    panel.spacing = unit(0.2, "cm"),
    axis.line = element_blank()
  )
leaf_n_seeding_plot
```


Plot together and save

```{r}
lma_resprouting_plot + lma_seeding_plot + leaf_n_resprouting_plot + leaf_n_seeding_plot
ggsave("figures/leaf_traits.png", width = 8.5, height = 4.8)
```



Rerun with phylogenetic linear models

```{r}
tree <- ape::read.tree("data/v0.1-big-seed-plant-trees/ALLMB.tre")
fire_response_and_leaf_traits$taxon_name <- gsub(" ", "_", fire_response_and_leaf_traits$taxon_name)
fire_response_and_leaf_traits <-
  fire_response_and_leaf_traits[fire_response_and_leaf_traits$taxon_name %in% tree$tip.label, ]

fire_response_and_leaf_traits$resprouting_binomial <-
  ifelse(fire_response_and_leaf_traits$resprouting_binomial == TRUE, 1, 0) # Make numeric

pruned_tree <- ape::drop.tip(tree, setdiff(tree$tip.label, fire_response_and_leaf_traits$taxon_name))
pruned_tree <- ape::as.phylo(pruned_tree)

sorted_data <-
  fire_response_and_leaf_traits[order(match(fire_response_and_leaf_traits$taxon_name, pruned_tree$tip.label)), ]

sorted_data <-
  sorted_data %>%
  select(taxon_name, mean_lma, mean_leaf_n, resprouting_binomial, seeding_binomial, woody_or_herb) %>%
  column_to_rownames("taxon_name")


# Return old variable to original state for running normal lm's
fire_response_and_leaf_traits <- fire_response_and_leaf_traits %>%
  filter(woody_or_herb %in% c("woody", "herb"))
fire_response_and_leaf_traits$woody_or_herb <-
  fire_response_and_leaf_traits$woody_or_herb %>%
  factor(levels = c("woody", "herb"))
```

```{r}
pm1 <-
  phylolm::phylolm(
    log10(mean_lma) ~ resprouting_binomial * woody_or_herb,
    data = sorted_data,
    phy = pruned_tree,
    model = "BM"
  )
summary(pm1)
```


```{r}
pm2 <-
  phylolm::phylolm(
    log10(mean_leaf_n) ~ resprouting_binomial * woody_or_herb,
    data = sorted_data,
    phy = pruned_tree,
    model = "BM"
  )
summary(pm2)
```


```{r}
pm3 <-
  phylolm::phylolm(
    log10(mean_lma) ~ seeding_binomial * woody_or_herb,
    data = sorted_data,
    phy = pruned_tree,
    model = "BM"
  )
summary(pm3)
```


```{r}
pm4 <-
  phylolm::phylolm(
    log10(mean_leaf_n) ~ seeding_binomial * woody_or_herb,
    data = sorted_data,
    phy = pruned_tree,
    model = "BM"
  )
summary(pm4)
```


---


# Check number of genera and families in our dataset

Total number of genera and families

```{r}
library(taxonlookup)

fire_response_and_leaf_traits <-
  fire_response_and_fire_frequency %>%
  left_join(mean_lma) %>%
  left_join(mean_leaf_n)

lookup_table <-
  fire_response_and_leaf_traits$taxon_name %>%
  lookup_table(by_species = TRUE, missing_action = "NA") %>%
  rownames_to_column("taxon_name")

fire_response_and_leaf_traits <-
  fire_response_and_leaf_traits %>%
  left_join(lookup_table, by = "taxon_name") %>%
  add_higher_order()

# Number of genera
fire_response_and_leaf_traits$genus %>% unique() %>% length() # 1423

# Number of families
fire_response_and_leaf_traits$family %>% unique() %>% length() # 213
```


Number of genera and families with resprouting and seeding data

```{r}
resprouting_data <- fire_response_and_leaf_traits %>% drop_na(resprouting_binomial)
resprouting_data$genus %>% unique() %>% length() # 1417
resprouting_data$family %>% unique() %>% length() # 213

seeding_data <- fire_response_and_leaf_traits %>% drop_na(seeding_binomial)
seeding_data$genus %>% unique() %>% length() # 732
seeding_data$family %>% unique() %>% length() # 149
```


Number of genera and families with lma and leaf N data

```{r}
lma_data <- fire_response_and_leaf_traits %>% drop_na(mean_lma)
lma_data$genus %>% unique() %>% length() # 720
lma_data$family %>% unique() %>% length() # 152

leaf_n_data <- fire_response_and_leaf_traits %>% drop_na(mean_leaf_n)
leaf_n_data$genus %>% unique() %>% length() # 465
leaf_n_data$family %>% unique() %>% length() # 114
```


Number of genera and families with data in all columns

```{r}
all_data <- fire_response_and_leaf_traits %>% drop_na(mean_lma, mean_leaf_n, resprouting_binomial, seeding_binomial)
all_data$genus %>% unique() %>% length() # 269
all_data$family %>% unique() %>% length() # 85
```
