---
title: "Phylogeny Figure"
output: html_notebook
editor_options:
  chunk_output_type: console
---



Load libraries

```{r}
library(tidyverse)
library(tidytree)
library(ggtreeExtra)
library(ggtree)
library(ggnewscale)
library(ape)
library(RColorBrewer)
library(phytools)
```



## Full tree (not pruned)

```{r}
make_phy_f <- function() {
  tt <- ape::read.tree("data/v0.1-big-seed-plant-trees/ALLMB.tre")
  a <- taxonlookup::lookup_table(tt$tip.label, by_species = TRUE)
  a <- rownames_to_column(a, "species")
  take_first <- function(x) x[1]
  one_tip <- group_by(a, family) %>% summarise(species = take_first(species))
  out <- ape::drop.tip(tt, tt$tip.label[!tt$tip.label %in% one_tip$species])
  out$tip.label <- one_tip$family[match(out$tip.label, one_tip$species)]
  plot(out, type = "r")
  saveRDS(out, "data/phy_plot.rds")
}
```


## Make data frame with summarised fire response, mean fire frequency and leaf traits for each family
    - Also maybe woody or herb?

**Start with resprouting and seeding data and only families with > N species/taxa with data**

### Resprouting data

Join to lookup table

```{r}
resprouting_seeding <- readRDS("data/resprouting_seeding.rds")

austraits <- read_rds("data/austraits-5.0.0.rds")

resprouting_seeding_new <- resprouting_seeding |>
  left_join(
    austraits$taxa |> select(all_of(c("taxon_name", "genus", "family"))),
    by = "taxon_name"
  )

resprouting_seeding_new$family %>% unique() %>% length() # 223 unique families
x <- resprouting_seeding_new %>% group_by(family) %>% summarise(count = n())
hist(x$count, breaks = 100)
```


Summarise resprouting by family

```{r}
resprouting_summary <-
  resprouting_seeding_new %>%
  drop_na(resprouting_binomial) %>%
  group_by(family) %>%
  summarise(
    prop_resprouting = sum(resprouting_binomial) / n(),
    data_count = n()
  ) %>%
  ungroup() |>
  arrange(-data_count) |>
  slice_head(n = 20)
```


Seeding data

```{r}
seeding_summary <-
  resprouting_seeding_new %>%
  drop_na(seeding_binomial, family) %>%
  group_by(family) %>%
  summarise(prop_seeding = sum(seeding_binomial) / n()) %>%
  ungroup()
```


Mean fire frequency data

```{r}
mean_fires <- read_csv("outputs/final_df_predicted_fires_and_fire_response.csv")
```


```{r}
mean_fires_new <-
  mean_fires %>%
  left_join(
    austraits$taxa |> select(all_of(c("taxon_name", "genus", "family"))),
    by = "taxon_name"
  )

mean_fires_summary <-
  mean_fires_new %>%
  group_by(family) %>%
  summarise(mean_fires_100 = mean(mean_fires_100)) %>%
  ungroup()
```


LMA data


```{r}
aus_native_lookup <- read_csv("data/aus_native_lookup.csv")

lma <-
  austraits$traits %>%
  filter(trait_name == "leaf_mass_per_area") %>%
  pivot_wider(names_from = "trait_name", values_from = "value") |>
  mutate(leaf_mass_per_area = as.numeric(leaf_mass_per_area))

# Filter out non-native taxa
lma <-
  lma %>%
  left_join(
    aus_native_lookup,
    by = c("taxon_name" = "canonical_name")) %>%
  filter(aus_native == TRUE) %>%
  select(-aus_native)

mean_lma <-
  lma %>%
  group_by(taxon_name) %>%
  summarise(mean_lma = mean(leaf_mass_per_area)) %>%
  ungroup()

mean_lma_new <-
  mean_lma %>%
  left_join(
    austraits$taxa |> select(all_of(c("taxon_name", "genus", "family"))),
    by = "taxon_name"
  )

mean_lma_new$family %>% unique() %>% length() # 186
x <- mean_lma_new %>% group_by(family) %>% summarise(count = n())
hist(x$count, breaks = 100)

mean_lma_summary <-
  mean_lma_new %>%
  drop_na(mean_lma, family) %>%
  group_by(family) %>%
  summarise(mean_lma = mean(mean_lma)) %>%
  ungroup()
```


Leaf N per dry mass data

```{r}
leaf_n <-
  austraits$traits %>%
  filter(trait_name == "leaf_N_per_dry_mass") %>%
  pivot_wider(names_from = "trait_name", values_from = "value") |>
  mutate(leaf_N_per_dry_mass = as.numeric(leaf_N_per_dry_mass))


# Filter out non-native taxa
leaf_n <-
  leaf_n %>%
  left_join(
    aus_native_lookup,
    by = c("taxon_name" = "canonical_name")) %>%
  filter(aus_native == TRUE) %>%
  select(-aus_native)

mean_leaf_n <-
  leaf_n %>%
  group_by(taxon_name) %>%
  summarise(mean_leaf_n = mean(leaf_N_per_dry_mass)) %>%
  ungroup()

mean_leaf_n_new <-
  mean_leaf_n %>%
  left_join(
    austraits$taxa |> select(all_of(c("taxon_name", "genus", "family"))),
    by = "taxon_name"
  )

mean_leaf_n_new$family %>% unique() %>% length() # 145
x <- mean_leaf_n_new %>% group_by(family) %>% summarise(count = n())
hist(x$count, breaks = 100)

mean_leaf_n_summary <-
  mean_leaf_n_new %>%
  drop_na(mean_leaf_n, family) %>%
  group_by(family) %>%
  summarise(mean_leaf_n = mean(mean_leaf_n)) %>%
  ungroup()
```



---

## Pruned tree with only families with resprouting and/or seeding data

```{r}
tt <- ape::read.tree("data/v0.1-big-seed-plant-trees/ALLMB.tre")
a <- taxonlookup::lookup_table(tt$tip.label, by_species = TRUE, missing_action = "NA")
a <- rownames_to_column(a, "species")
a <- a %>%
  left_join(
    austraits$taxa |>
      filter(taxonomic_status == "accepted") |>
      select(all_of(c("genus", "family"))) |>
      distinct(),
    by = "genus"
  ) %>%
  mutate(family = coalesce(family.x, family.y)) %>%
  select(-family.x, -family.y)
take_tenth <- function(x) x[10] # Not [1] because the first species of some families is taxonomically ambiguous
one_tip <- group_by(a, family) %>% summarise(species = take_tenth(species))
out <- ape::drop.tip(tt, tt$tip.label[!tt$tip.label %in% one_tip$species])
out$tip.label <- one_tip$family[match(out$tip.label, one_tip$species)]

# Align tree with APCalign #TODO

out_pruned <- ape::drop.tip(out, out$tip.label[!out$tip.label %in% resprouting_summary$family])

ggtree(out_pruned, ladderize = FALSE) +
  geom_tiplab()
```


Number of native Australian taxa

```{r}
library(tidyverse)
taxonomic_resources <- APCalign::load_taxonomic_resources()

apc <- taxonomic_resources$APC |>
  filter(
    taxonomic_status == "accepted",
    taxon_rank %in% c("species", "form", "variety", "subspecies")
  )

native_lookup <- APCalign::native_anywhere_in_australia(apc$canonical_name, taxonomic_resources)

apc_native <-
  apc %>%
  left_join(native_lookup, by = c("canonical_name" = "species"))

species_diversity <-
  apc_native %>%
  filter(
    taxon_rank == "species", # Filter for species only
    native_anywhere_in_aus == "native" # Filter for native species only
  ) %>%
  group_by(family) %>%
  summarise(n_species = n_distinct(canonical_name)) %>%
  ungroup() %>%
  arrange(-n_species)

species_diversity
```


Number of woody species #TODO

```{r}

```



New data frame to label and colour key families

```{r}
extra_data <- tibble(label = out_pruned$tip.label)

extra_data <-
  extra_data %>%
  left_join(species_diversity, by = c("label" = "family"))

new_tree <- full_join(out_pruned, extra_data, by = "label")
```



```{r}
new_tree@phylo |>
  as_tibble() |>
  filter(label %in% c("Fabaceae", "Myrtaceae", "Orchidaceae", "Proteaceae", "Asteraceae", "Poaceae"))
```

```{r}
ggtree(new_tree, layout = "rectangular", size = 0.7) +
  geom_tiplab(
    aes(subset = (!node %in% c(3, 7, 10, 14, 15, 19))),
    offset = 3,
    size = 4,
    show.legend = FALSE
  ) +
  geom_tiplab(
    aes(subset = (node %in% c(3, 7, 10, 14, 15, 19)), fontface = "bold"),
    offset = 3,
    size = 4,
    show.legend = FALSE
  ) +
  geom_text(
    aes(label = n_species),
    hjust = 1,
    nudge_x = 0,
    nudge_y = -0.25,
    size = 3.5
  ) +
  scale_colour_manual(values = c("#ECC119", "#F767C3", "#EA3138", "#B2CE6E", "#FF932C")) +
  theme(
    plot.margin = unit(c(5, 5, 5, 5), "mm"),
    legend.margin = margin(-0, 0, 0, 0, unit = "cm"),
    legend.text = element_text(size = 8.5),
    legend.justification = c("right"),
    legend.box.margin = margin(0, 0, 0, -20)
  ) +
  geom_fruit(
    data = mean_fires_summary,
    geom = geom_tile,
    mapping = aes(x = mean_fires_100, y = family, fill = mean_fires_100),
    offset = 0.7,
    pwidth = 0,
    width = 8,
    height = 1
  ) +
  scale_fill_distiller(
    palette = "YlOrRd",
    direction = 1,
    trans = "log10",
    name = "Mean fire frequency",
    guide = guide_colourbar(
      ticks.colour = "black",
      ticks.linewidth = 1,
      frame.colour = "black",
      frame.linewidth = 1,
      direction = "horizontal",
      title.position = "top",
      title.vjust = 1,
      title.theme = element_text(size = 9),
      barwidth = 5,
      barheight = 0.7,
      order = 1)
  ) +
  geom_fruit(
    data = mean_fires_summary,
    geom = geom_text,
    mapping = aes(x = mean_fires_100, y = family, label = round(mean_fires_100, digits = 1)),
    colour = "black",
    offset = 0.058,
    pwidth = 0
  ) +
  ggnewscale::new_scale_fill() +
  geom_fruit(
    data = resprouting_summary,
    geom = geom_tile,
    mapping = aes(x = prop_resprouting, y = family, fill = prop_resprouting),
    offset = 0.058,
    pwidth = 0,
    width = 8,
    height = 1
  ) +
  scale_fill_distiller(
    palette = "RdPu",
    direction = 1,
    guide = guide_colourbar(
      ticks.colour = "black",
      ticks.linewidth = 1,
      frame.colour = "black",
      frame.linewidth = 1,
      direction = "horizontal",
      title.position = "top",
      title.vjust = 1,
      title.theme = element_text(size = 9),
      barwidth = 5,
      barheight = 0.7,
      order = 2),
    name = "Proportion resprouting"
  ) +
  ggnewscale::new_scale_fill() +
  geom_fruit(
    data = seeding_summary,
    geom = geom_tile,
    mapping = aes(x = prop_seeding, y = family, fill = prop_seeding),
    offset = 0.058,
    pwidth = 0,
    width = 8,
    height = 1
  ) +
  scale_fill_distiller(
    palette = "YlOrBr",
    direction = 1,
    breaks = c(0.4, 0.6, 0.8),
    guide = guide_colourbar(
      ticks.colour = "black",
      ticks.linewidth = 1,
      frame.colour = "black",
      frame.linewidth = 1,
      direction = "horizontal",
      title.position = "top",
      title.vjust = 1,
      title.theme = element_text(size = 9),
      barwidth = 5,
      barheight = 0.7,
      order = 3),
    name = "Proportion seeding"
  ) +
  ggnewscale::new_scale_fill() +
  geom_fruit(
    data = mean_lma_summary,
    geom = geom_tile,
    mapping = aes(x = mean_lma, y = family, fill = mean_lma),
    offset = 0.058,
    pwidth = 0,
    width = 8,
    height = 1) +
  scale_fill_distiller(
    palette = "Greens",
    direction = 1,
    guide = guide_colourbar(
      ticks.colour = "black",
      ticks.linewidth = 1,
      frame.colour = "black",
      frame.linewidth = 1,
      direction = "horizontal",
      title.position = "top",
      title.vjust = 1,
      title.theme = element_text(size = 9),
      barwidth = 5,
      barheight = 0.7,
      order = 4),
    name = "Mean LMA"
  ) +
  ggnewscale::new_scale_fill() +
  geom_fruit(
    data = mean_leaf_n_summary,
    geom = geom_tile,
    mapping = aes(x = mean_leaf_n, y = family, fill = mean_leaf_n),
    offset = 0.058,
    pwidth = 0,
    width = 8,
    height = 1
  ) +
  scale_fill_distiller(
    palette = "PuBu",
    direction = 1,
    guide = guide_colourbar(
      ticks.colour = "black",
      ticks.linewidth = 1,
      frame.colour = "black",
      frame.linewidth = 1,
      direction = "horizontal",
      title.position = "top",
      title.vjust = 1,
      title.theme = element_text(size = 9),
      barwidth = 5,
      barheight = 0.7,
      order = 5),
    name = "Mean leaf N"
  )

# Myrtaceae = 7, Fabaceae = 3, Asteraceae = 10, Poaceae = 15, Proteaceae = 14
#ggsave("phylogeny-figure.png", width = 7.5, height = 6.75)
```


```{r}

```


Add leaf trait bands
Highlight significant families and label them
Can use scaleClade() to focus on important clades


---

# Make a figure for only significant families

    - Proteaceae
    - Myrtaceae
    - Fabaceae
    - Asteraceae
    - Poaceae

Subset tree into only these families

```{r}
tt <- read.tree("raw_data/v0.1-big-seed-plant-trees/ALLMB.tre")
a <- taxonlookup::lookup_table(tt$tip.label, by_species = TRUE)
a <- rownames_to_column(a, "species")
take_first <- function(x) x[1]
one_tip <- group_by(a, family) %>% summarise(species = take_first(species))
out <- drop.tip(tt, tt$tip.label[!tt$tip.label %in% one_tip$species])
out$tip.label <-
    one_tip$family[match(out$tip.label, one_tip$species)]

key_aus_families <- c("Proteaceae", "Myrtaceae", "Fabaceae", "Asteraceae", "Poaceae")
out_pruned <- drop.tip(out, out$tip.label[!out$tip.label %in% key_aus_families])
```



```{r}
ggtree(out_pruned, layout = "rectangular", size = 0.6) +
  geom_fruit(
    data = resprouting_summary,
    geom = geom_tile,
    mapping = aes(x = prop_resprouting, y = family, fill = prop_resprouting),
    pwidth = 0,
    offset = 0.5,
    width = 15,
    height = 1
  ) +
  scale_fill_distiller(palette = "Blues", direction = 1) +
  new_scale_fill() +
  geom_fruit(
    data = seeding_summary,
    geom = geom_tile,
    mapping = aes(x = prop_seeding, y = family, fill = prop_seeding),
    pwidth = 0,
    offset = 0.045,
    width = 15,
    height = 1
  ) +
  scale_fill_distiller(palette = "Reds", direction = 1) +
  new_scale_fill() +
  geom_fruit(
    data = mean_fires_summary,
    geom = geom_tile,
    mapping = aes(x = mean_fires_100, y = family, fill = mean_fires_100),
    pwidth = 0,
    offset = 0.045,
    width = 15,
    height = 1
  ) +
  scale_fill_distiller(palette = "Purples", direction = -1)
```

